### 2022-09-13 manuscript-figures.Rmd
First commit

---
title: "EIP CRE transmission network manuscript figures and data"
date: "2/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# load libraries
library(tidyverse)
library(ggtree)
library(pheatmap)
library(igraph)
library(regentrans)
library(tidygraph)
library(ggraph)
library(exact2x2)
library(cowplot)
library(readxl)
library(ape)
library(phytools)
library(ggtree)
library(ggtext)
library(lubridate)

# set plotting theme
theme_set(
  theme_bw() + 
    theme(text = element_text(size = 10), 
          axis.text.x = element_text(angle = 0),
  strip.background = element_rect(fill="white",linetype='blank')))

# load transition edge script
source('code/get_transition_edges.R')

# set seed
set.seed(0)
```

```{r}
# load in data
# facility and ST
facil_st <- read_csv('data/isolate_data.csv')
facil_st_mn <- facil_st %>% filter(State == 'MN')
write_csv(facil_st_mn, 'data/mn_isolate_data.csv')
# public isolate data
metadat_complete <- read_csv('data/public_isolate_metadata.csv')
# public isolate trees
tree_list <- read_rds('data/tree_list.rds')
# pairwise snv distances
kp_dists <- read_rds('data/kp_dists.rds')
ec_dists <- read_rds('data/eh_dists.rds')
# phylogenetic trees
tr_kp <- read.tree('data/kp_tr.tree')
tr_ec <- read.tree('data/eh_tr.tree')
# clinical factors
pt_feats_ct <- read_csv('data/pt_feats_ct.csv')
pt_feats_mn <- read_csv('data/pt_feats_mn.csv')
pt_feats_tn <- read_csv('data/pt_feats_tn.csv')
# shared facility exposures
pr_hosp_pairs_ct <- read_csv('data/pr_hosp_pairs_ct.csv')
pr_hosp_pairs_mn <- read_csv('data/pr_hosp_pairs_mn.csv')
pr_hosp_pairs_tn <- read_csv('data/pr_hosp_pairs_tn.csv')
# patient transfer networks
ct_trans_net_sub <- read_csv('data/trans_net_ct.csv')
mn_trans_net_sub <- read_csv('data/trans_net_mn.csv')
tn_trans_net_sub <- read_csv('data/trans_net_tn.csv')
# facility types
facil_types <- read_csv('data/facil_types.csv')
# facility counts
facil_type_counts <- read_csv('data/facil_type_counts.csv')
# facility w prior hospitals
facil_w <- read_csv('data/facil_w_st171_prior_hosps.csv')
# prior tn hospitals
prior_hosps_tn <- read_csv('data/pr_hosps_tn.csv')
# tn facil clutser
tn_facil_cluster <- read_csv('data/tn_facil_cluster.csv')
```

```{r}
# dominant STs and colors
st_cts <- rev(sort(table(facil_st$ST)))
dominant_sts <- names(st_cts)[st_cts > 10 & names(st_cts) != 'Other species']
st_cols <- rep(NA, length(st_cts))
names(st_cols) <- names(st_cts)
st_cols['K. pneumoniae ST258'] <- 'coral4'
st_cols['K. pneumoniae ST307'] <- 'coral3'
st_cols['E. hormaechei ST171'] <- 'aquamarine4'
st_cols['E. hormaechei ST114'] <- 'aquamarine3'
st_cols[is.na(st_cols)] <- 'grey'
st_cols['Other'] <- 'grey'
```

## Data summary

### Number of facilities with CRE per state

```{r}
facil_st %>% 
  group_by(State) %>% 
  summarize(n_facil = n_distinct(Facility_ID))
```

## Facility type information (Table S1)

```{r}
# number of each facility type per state
facil_type_counts
```

```{r}
# facility types with at least one WGS
facil_types %>% 
  group_by(state, type) %>% 
  summarize(count = n()) %>% 
  pivot_wider(names_from = state, values_from = count, values_fill = 0) 
```

### Species counts by state (Table 1)

```{r}
state_dates <- facil_st %>% mutate(Cult_Date = gsub('-.*|.*/','', Cult_Date)) %>% 
  filter(!is.na(Cult_Date)) %>% 
  group_by(State, Cult_Date) %>% summarize() %>% 
  group_by(State) %>% summarize(min_cult_date = min(Cult_Date),
                                max_cult_date = max(Cult_Date))

facil_st %>% 
  mutate(State = factor(State, levels = c('MN','TN','CT'))) %>% 
  group_by(State, Species) %>% 
  count() %>% full_join(state_dates, .) %>% 
  suppressMessages()
```

### Species counts by facility (Figure S1)

```{r}
fs1 <- facil_st %>% filter(Species != 'Other') %>% 
    mutate(ST = factor(ST, levels = rev(names(st_cols))),
           Species = factor(Species, levels = c('K. pneumoniae','E. hormaechei','Other'))) %>% 
  ggplot(aes(x = Facility_ID, fill = ST)) + geom_bar(color = 'white') + facet_grid(Species~State, scales = 'free', space = 'free') +
  scale_fill_manual(values = st_cols, 
                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
                               'K. pneumoniae ST258', 'K. pneumoniae ST307', 
                               'Other'), 
                    labels = c('*E. hormaechei* ST171', '*E. hormaechei* ST114',
                               '*K. pneumoniae* ST258', '*K. pneumoniae* ST307', 
                               'Other'),
                    drop = FALSE) +
  labs(y = 'Number of isolates', fill = '', x = 'Facility') +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.text.y = element_text(face = 'italic'), legend.text = element_markdown())

ggsave(plot = fs1, filename = 'figures/FigS1.png', width = 7, height = 3)

fs1
```

## Public data (Figure S2)

```{r}
public_loc_plot <- metadat_complete %>% 
  mutate(state = ifelse(is.na(state), 'Unknown', state),
         kpc = case_when(is.na(kpc) ~ 'Unknown',
                         kpc == 0 ~ 'No',
                         kpc == 1 ~ 'Yes'),
         kpc = factor(kpc, levels = c('Yes', 'No', 'Unknown'))) %>% 
  group_by(state) %>% 
  mutate(tot = n()) %>%
  ggplot(aes(x = reorder(state, tot), fill = kpc)) +
  geom_bar() +
  labs(x = 'State', y = 'Number of public isolates', fill = 'KPC\npresent') +
  scale_y_log10() +
  scale_fill_manual(values = c('aquamarine3', 'aquamarine4', 'darkgrey')) +
  coord_flip()

public_dates_plot <- metadat_complete %>% 
  ggplot(aes(x = year)) +
  geom_bar() +
  labs(x = 'Year', y = 'Number of public isolates')

plot_grid(public_dates_plot, public_loc_plot, ncol = 1, rel_heights = c(0.5, 1), align = 'vh', labels = 'AUTO') %>% 
  ggsave(filename = 'figures/FigS2.png', width = 7, height = 10)
```

## Importation vs. HGT

```{r}
kpc_bin <- metadat_complete %>% select(isolate_no, kpc) %>% drop_na() 
mn_bin <- metadat_complete %>% select(isolate_no, mn_bin) %>% drop_na()
tn_bin <- metadat_complete %>% select(isolate_no, tn_bin) %>% drop_na()
ct_bin <- metadat_complete %>% select(isolate_no, ct_bin) %>% drop_na()

genomes <- unique(c(kpc_bin$isolate_no, mn_bin$isolate_no, tn_bin$isolate_no, ct_bin$isolate_no))

transitions <- lapply(1:length(tree_list), function(x){

  st <- names(tree_list)[[x]]

  t <- midpoint.root(tree_list[[x]])
  t <- keep.tip(t, t$tip.label[t$tip.label %in% genomes])

  kpc_trans_edges = mn_trans_edges = tn_trans_edges = ct_trans_edges = NULL

  kpc_bin_sub <- kpc_bin %>% filter(isolate_no %in% t$tip.label)
  if(length(unique(kpc_bin_sub$kpc)) != 1){
    kpc_trans_edges <- transition_edge_pipeline(kpc_bin_sub, t, continuous = FALSE, ar_conf = 0.5)
  }
  mn_bin_sub <- mn_bin %>% filter(isolate_no %in% t$tip.label)
  if(length(unique(mn_bin_sub$mn_bin)) != 1){
    mn_trans_edges <- transition_edge_pipeline(mn_bin_sub, t, continuous = FALSE, ar_conf = 0.5)
  }
  tn_bin_sub <- tn_bin %>% filter(isolate_no %in% t$tip.label)
  if(length(unique(tn_bin_sub$tn_bin)) != 1){
    tn_trans_edges <- transition_edge_pipeline(tn_bin_sub, t, continuous = FALSE, ar_conf = 0.5)
  }
  ct_bin_sub <- ct_bin %>% filter(isolate_no %in% t$tip.label)
  if(length(unique(ct_bin_sub$ct_bin)) != 1){
    ct_trans_edges <- transition_edge_pipeline(ct_bin_sub, t, continuous = FALSE, ar_conf = 0.5)
  }

  if(is.null(kpc_trans_edges) & is.null(mn_trans_edges) & is.null(tn_trans_edges) & is.null(ct_trans_edges)) return(NULL)

  trans_sum <- paste0(
    ifelse(mn_trans_edges$mn_bin$trans_dir != 1, '', 'to_mn '), 
    ifelse(tn_trans_edges$tn_bin$trans_dir != 1, '', 'to_tn '), 
    ifelse(ct_trans_edges$ct_bin$trans_dir != 1, '', 'to_ct '), 
    ifelse(kpc_trans_edges$kpc$trans_dir != 1, '', 'gain_kpc ')
  )

  edge <- data.frame(t$edge, 
                  edge_num=1:nrow(t$edge), 
                  edge_val = trans_sum[1:(length(trans_sum)-1)]) %>% 
    mutate(edge_val = ifelse(edge_val == '', NA, edge_val))
  colnames(edge)=c("parent", "node", "edge_num","edge_val")

  trans_edge_nums <- sapply(unique(metadat_complete$isolate_no[grepl('^EIP',metadat_complete$sources) &
                                                                  metadat_complete$isolate_no %in% t$tip.label]),
                            function(i){
    trans <- NA
    child <- which(t$tip.label == i)
    edge_num <- NA
    while(is.na(trans) & child != (Ntip(t) + 1)){ # if find transition going backwards or get to root
      child_edge_info <- edge %>% filter(node == child)
      trans <- child_edge_info %>% select(edge_val) %>% deframe()
      child <- child_edge_info %>% select(parent) %>% deframe()
      if(!is.na(trans)){
        edge_num <- child_edge_info %>% select(edge_num) %>% deframe()
      }
    }
    edge_num
  })

  trans_edge_nums <- trans_edge_nums[!is.na(trans_edge_nums)] #recently added

  grps <- data.frame(id=names(trans_edge_nums),
                     class=trans_sum[trans_edge_nums],
                     cluster=trans_edge_nums) %>% 
    mutate(state = factor(gsub('_.*','',id), levels = c('MN','TN','CT')),
           ST = sapply(id, function(x){
             val <- unlist(unique(metadat_complete$ST[metadat_complete$isolate_no == x & !is.na(metadat_complete$ST)]))
             val <- ifelse(length(val) == 0, paste0('ST', unlist(unique(metadat_complete$mlst[metadat_complete$isolate_no == x & !is.na(metadat_complete$mlst)]))), val)
             val <- ifelse(length(val) == 0 | val == 'ST', st, val)
             ifelse(length(val) == 0, NA, val)
           }),
           species = sapply(id, function(x){
             val <- unlist(unique(metadat_complete$species[metadat_complete$isolate_no == x & !is.na(metadat_complete$species)]))
             ifelse(length(val) == 0, NA, val)
           }),
           ST_orig = ST,
           ST = ifelse(ST %in% c('E. hormaechei ST171','E. hormaechei ST114',
                                 'K. pneumoniae ST258','K. pneumoniae ST307', 
                                 'E. cloacae ST171', 'E. cloacae ST114', 
                                 'ST258', 'ST307', 'ST114', 'ST171'),
                       ST, 'Other'),
           tree = st
           )
  if(nrow(grps) == 0) return(NULL)
  grps
}) %>% 
  bind_rows() %>% suppressWarnings()

transitions <- transitions %>% 
  mutate(class = gsub('NA| ', '', class), 
         class = ifelse(is.na(class), '', class),
         hgt_type = ifelse(grepl('gain_kpc', class), 'Unknown', ''),
         hgt_type = ifelse(class == 'gain_kpc', 'In-state', hgt_type),
         class2 = ifelse(toupper(gsub('to_','',class)) != state & !grepl('gain_kpc', class), '', class),
         class2 = ifelse(grepl('gain_kpc', class2), 'Putative\nHGT', class2),
         class2 = ifelse(class2 == '', 'Unknown', class2),
         class2 = ifelse(grepl('to_', class2), 'Putative\nimportation', class2),
         class2 = factor(class2, levels = c('Putative\nimportation','Putative\nHGT','Unknown')),
         ST = ifelse(ST %in% c('ST258','ST307'), paste0('K. pneumoniae ', ST), ST),
         ST = ifelse(ST %in% c('ST171','ST114'), paste0('E. hormaechei ', ST), ST),
         ST = gsub('cloacae', 'hormaechei', ST),
         ST_orig = gsub('cloacae', 'hormaechei', ST_orig)
         )

keepers <- facil_st %>% 
  filter(!(duplicated(Patient_ID) & duplicated(ST))) %>% 
  pull(Sample_ID)
```

### Trees used (Figure S3)

```{r}
states <- metadat_complete %>% select(isolate_no, state) %>% deframe()
states <- unique(states[c(tree_list[[1]]$tip.label, tree_list[[4]]$tip.label, tree_list[[7]]$tip.label, tree_list[[12]]$tip.label)])
states <- states[!is.na(states)]
state_cols <- c('#e6194b', '#3cb44b', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffd8b1', '#000075', '#808080', '#ffffff', '#000000', '#ffe119')
names(state_cols) <- states

plot_tr_locs <- function(tr, og, title = NULL){
  tr <- root(tr, og)
  tr <- drop.tip(tr, og)
  states <- metadat_complete %>% select(isolate_no, state) %>% deframe()
  states <- states[tr$tip.label]
  kpc <- metadat_complete %>% select(isolate_no, kpc) %>% deframe()
  kpc <- kpc[tr$tip.label]
  no_info <- is.na(states) & is.na(kpc)
  states <- states[!no_info]
  kpc <- kpc[!no_info]
  kpc <- ifelse(kpc == 1, 'Present', 'Absent')
  tr <- keep.tip(tr, union(names(states), names(kpc)))
  states <- c(states, rep(NA, Nnode(tr)))
  kpc <- c(kpc, rep(NA, Nnode(tr)))
  ggtree(tr) + 
    geom_tippoint(aes(col = states, shape = kpc)) + 
    scale_color_manual(values = state_cols) + 
    scale_shape_manual(values = c('Present' = 16, 'Absent' = 1)) +
    guides(color = guide_legend(ncol = 2)) +
    labs(title = title, shape = 'KPC', col = 'State')
}

tr_st114 <- plot_tr_locs(tree_list[[7]], 'MN_CRE226', '*E. hormaechei* ST114') + 
  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
tr_st171 <- plot_tr_locs(tree_list[[4]], 'MN_CRE226', '*E. hormaechei* ST171') + 
  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
tr_st258 <- plot_tr_locs(tree_list[[1]], '573.1451', '*K. pneumoniae* ST258') + 
  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
tr_st307 <- plot_tr_locs(tree_list[[12]], '573.13662', '*K. pneumoniae* ST307') + 
  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
leg <- get_legend(plot_tr_locs(tree_list[[7]], 'MN_CRE226', 'E. hormaechei ST114'))

fs3 <- plot_grid(plot_grid(tr_st171, tr_st114, tr_st258, tr_st307), leg)

ggsave(plot = fs3, filename = 'figures/FigS3.png', width = 8, height = 8)

fs3
```

### Importation vs. HGT (Figure 1)

```{r}
clust_cols <- st_cols[transitions$ST]
names(clust_cols) <- paste0(transitions$state, transitions$ST, transitions$cluster)

clust_cols <- (c(clust_cols[clust_cols == "grey"],
                clust_cols[clust_cols == "coral3"],
                clust_cols[clust_cols == "coral4"],
                clust_cols[clust_cols == "aquamarine3"],
                clust_cols[clust_cols == "aquamarine4"]
                ))

transitions <- transitions %>% filter(id %in% keepers) %>% 
  mutate(ST = factor(ST, levels = (c("E. hormaechei ST171", "E. hormaechei ST114", 
                                     "K. pneumoniae ST258", "K. pneumoniae ST307", 
                                     'Other'))),
         state = factor(state, levels = c('CT', 'MN', 'TN'))) %>%
  arrange(ST)

f1 <- transitions %>% 
  ggplot(aes(x = class2, 
             fill = factor(paste0(state, ST, cluster), 
                           levels = unique(paste0(state, ST, cluster))))) + 
  geom_bar(color = 'white', position = position_stack(reverse = TRUE)) + 
  scale_fill_manual(values = clust_cols,
                    breaks = rev(names(clust_cols)[!duplicated(clust_cols)]),
                    labels = c('*E. hormaechei* ST171', '*E. hormaechei* ST114',
                               '*K. pneumoniae* ST258', '*K. pneumoniae* ST307',
                               'Other'),
                    drop = FALSE) +
  facet_grid(~state, scales = 'free', space = 'free') +
  labs(fill = 'ST cluster', x = 'Cluster type', y = 'Number of isolates') +
  theme(legend.text = ggtext::element_markdown())

ggsave(plot = f1, filename = 'figures/Fig1.png', width = 7, height = 4)

f1
```

```{r}
transitions %>% 
  group_by(ST, class2) %>% 
  summarize(count = n()) %>% 
  group_by(ST) %>% 
  mutate(tot = sum(count),
         prop = count/tot)
```

### HGT locations (Figure S4)

```{r}
fs4 <- transitions %>% 
  filter(id %in% keepers & hgt_type != '') %>%
  ggplot(aes(x = hgt_type, 
             fill = factor(paste0(state, ST, cluster), 
                           levels = unique(paste0(state, ST, cluster))))) + 
  geom_bar(color = 'white', position = position_stack(reverse = TRUE)) + 
  scale_fill_manual(values = clust_cols,
                    breaks = rev(names(clust_cols)[!duplicated(clust_cols)]),
                    labels = c('*E. hormaechei* ST171', '*E. hormaechei* ST114',
                               '*K. pneumoniae* ST258', '*K. pneumoniae* ST307',
                               'Other'),
                    drop = FALSE) +
  facet_grid(~state, scales = 'free', space = 'free') +
  labs(fill = 'ST cluster', x = 'HGT location', y = 'Number of isolates') +
  theme(legend.text = ggtext::element_markdown())

ggsave(plot = fs4, filename = 'figures/FigS4.png', width = 6, height = 3)

fs4
```
```{r}
# earliest date for ST114 and ST307 
facil_st %>% 
  filter(ST %in% c('E. hormaechei ST114', 'K. pneumoniae ST307') & State == 'TN') %>% 
  group_by(State, ST) %>% 
  mutate(Cult_Date = mdy(Cult_Date)) %>% 
  slice_min(Cult_Date) %>% 
  select(State, ST, Cult_Date)
```

## Distances

```{r}
# snv distances
kp_snv_dists <- get_pair_types(kp_dists, facil_st %>% select(Sample_ID, Facility_ID) %>% deframe(), pt = NULL) %>% 
  filter(!sapply(1:nrow(.), function(x) sort(c(as.character(isolate1[x]),as.character(isolate2[x])))) %>% t() %>% duplicated()) %>% 
  mutate(st1 = sapply(isolate1, function(x) facil_st$ST[facil_st$Sample_ID == x]),
         st2 = sapply(isolate2, function(x) facil_st$ST[facil_st$Sample_ID == x]),
         state1 = sapply(isolate1, function(x) facil_st$State[facil_st$Sample_ID == x]),
         state2 = sapply(isolate2, function(x) facil_st$State[facil_st$Sample_ID == x])) %>% 
  filter(st1 == st2 & !grepl('novel',st1)) %>% 
  mutate(state_pair = paste(state1, state2))

ec_snv_dists <- get_pair_types(ec_dists, facil_st %>% select(Sample_ID, Facility_ID) %>% deframe(), pt = NULL) %>% 
  filter(!sapply(1:nrow(.), function(x) sort(c(as.character(isolate1[x]),as.character(isolate2[x])))) %>% t() %>% duplicated()) %>% 
  mutate(st1 = sapply(isolate1, function(x) facil_st$ST[facil_st$Sample_ID == x]),
         st2 = sapply(isolate2, function(x) facil_st$ST[facil_st$Sample_ID == x]),
         state1 = sapply(isolate1, function(x) facil_st$State[facil_st$Sample_ID == x]),
         state2 = sapply(isolate2, function(x) facil_st$State[facil_st$Sample_ID == x])) %>% 
  filter(st1 == st2 & !grepl('novel',st1)) %>% 
  mutate(state_pair = paste(state1, state2))

snv_dat <- bind_rows(kp_snv_dists, ec_snv_dists) %>% filter(state1 == state2) %>% 
  mutate(state1 = factor(state1, levels = c('MN','TN','CT')),
         pair_type = ifelse(is.na(pair_type), 'Unknown', pair_type),
         under_thresh = ifelse(pairwise_dist <= 15, '(10,15]','(15,∞)'),
         under_thresh = ifelse(pairwise_dist <= 10, '(5,10]',under_thresh),
         under_thresh = ifelse(pairwise_dist <= 5, '[0,5]',under_thresh),
         under_thresh = factor(under_thresh, levels = c('[0,5]','(5,10]','(10,15]','(15,∞)')),
         ST_orig = st1,
         ST = ifelse(st1 %in% dominant_sts, st1, 'Other'),
         Species = gsub(' ST.*','',st1),
         ST = gsub('.* ','',ST))
```

```{r}
# patristic distances
kp_pat <- cophenetic.phylo(tr_kp)
ec_pat <- cophenetic.phylo(tr_ec)

kp_pat_dists <- get_pair_types(kp_pat, facil_st %>% select(Sample_ID, Facility_ID) %>% deframe(), pt = NULL) %>% 
  filter(!sapply(1:nrow(.), function(x) sort(c(as.character(isolate1[x]),as.character(isolate2[x])))) %>% t() %>% duplicated()) %>% 
  mutate(st1 = sapply(isolate1, function(x) facil_st$ST[facil_st$Sample_ID == x]),
         st2 = sapply(isolate2, function(x) facil_st$ST[facil_st$Sample_ID == x]),
         state1 = sapply(isolate1, function(x) facil_st$State[facil_st$Sample_ID == x]),
         state2 = sapply(isolate2, function(x) facil_st$State[facil_st$Sample_ID == x])) %>% 
  filter(st1 == st2 & !grepl('novel',st1)) %>% 
  mutate(state_pair = paste(state1, state2))

ec_pat_dists <- get_pair_types(ec_pat, facil_st %>% select(Sample_ID, Facility_ID) %>% deframe(), pt = NULL) %>% 
  filter(!sapply(1:nrow(.), function(x) sort(c(as.character(isolate1[x]),as.character(isolate2[x])))) %>% t() %>% duplicated()) %>% 
  mutate(st1 = sapply(isolate1, function(x) facil_st$ST[facil_st$Sample_ID == x]),
         st2 = sapply(isolate2, function(x) facil_st$ST[facil_st$Sample_ID == x]),
         state1 = sapply(isolate1, function(x) facil_st$State[facil_st$Sample_ID == x]),
         state2 = sapply(isolate2, function(x) facil_st$State[facil_st$Sample_ID == x])) %>% 
  filter(st1 == st2 & !grepl('novel',st1)) %>% 
  mutate(state_pair = paste(state1, state2))

pat_dat <- bind_rows(kp_pat_dists, ec_pat_dists) %>% filter(state1 == state2) %>% 
  mutate(state1 = factor(state1, levels = c('MN','TN','CT')),
         pair_type = ifelse(is.na(pair_type), 'Unknown', pair_type),
         ST_orig = st1,
         ST = ifelse(st1 %in% dominant_sts, st1, 'Other'),
         Species = gsub(' ST.*','',st1),
         ST = gsub('.* ','',ST))
```

## Intra- and inter-facility pairwise SNV distances

### Fraction of intra-facility isolates over time (Figure S5)

```{r}
snvdat_orig <- snv_dat[,1:6] %>% mutate(pairwise_dist = as.numeric(pairwise_dist))

frac_intras <- lapply(unique(c(snv_dat$st1, snv_dat$st2)), function(st){
  lapply(c('MN MN','TN TN','CT CT'), function(s){
      snvdat_orig_sub <- snvdat_orig[snv_dat$st1 == st & snv_dat$st2 == st & snv_dat$state_pair == s,]
      frac_intra <- NULL
      if(nrow(snvdat_orig_sub) != 0){
        frac_intra <- get_frac_intra(snvdat_orig_sub) %>% mutate(ST = st, state = gsub(' .*','', s))
      }
      frac_intra
  }) %>% bind_rows()
}) %>% bind_rows()

fs5 <- frac_intras %>% 
    mutate(ST = ifelse(ST %in% c('E. hormaechei ST114','E. hormaechei ST171', 
                                 'K. pneumoniae ST258', 'K. pneumoniae ST307'), ST, 
                       'Other'),
           ST = factor(ST, levels = c('E. hormaechei ST171', 'E. hormaechei ST114', 
                                      'K. pneumoniae ST258', 'K. pneumoniae ST307', 
                                      'Other'),
                       labels = c('*E. hormaechei* ST171', '*E. hormaechei* ST114', 
                                      '*K. pneumoniae* ST258', '*K. pneumoniae* ST307', 
                                      'Other'))) %>% 
  filter(frac_intra != 0 & ST != 'Other') %>% 
  ggplot(aes(x = pairwise_dist, y = frac_intra)) + 
  geom_bar(stat = "identity", fill = "blue", alpha = 0.5) + 
  labs(x = "Pairwise SNV distance", y = "Fraction of intra-facility pairs") + xlim(0, 50) + 
  facet_wrap(state~ST, scales = 'free') +
  theme(strip.text = element_markdown())

ggsave(plot = fs5, filename = 'figures/FigS5.png', width = 6, height = 5)

fs5
```

### Pairwise SNV distances by pair type, state, and dominant STs (Figure 2)

```{r}
f2 <- snv_dat %>% 
  mutate(pairwise_dist = ifelse(pairwise_dist == 0, 1, pairwise_dist),
         pair_type = factor(gsub(' pair', '', pair_type), levels = c('Intra-facility','Inter-facility')),
         Species = gsub('^', '\\*', Species),
         Species = gsub('$', '\\*', Species),
         state1 = factor(state1, levels = c('CT', 'MN', 'TN'))) %>% 
  filter(pair_type != 'Unknown' & ST != 'Other') %>% 
  ggplot(aes(x = pair_type, y = pairwise_dist)) + 
  geom_jitter(aes(fill = under_thresh, color = under_thresh),alpha = 0.5) +
  scale_color_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
  geom_violin(position = 'identity',alpha=0, color = 'azure4') +
  facet_grid(Species+ST~state1, scales = 'free') + scale_y_log10() +
  labs(x = 'Pair type',y = 'Pairwise SNV distance', color = 'SNV bin', fill = 'SNV bin') +
  theme(strip.text.y = element_markdown())

ggsave(plot = f2, filename = 'figures/Fig2.png', width = 6.5, height = 5)

f2
```

### Pairwise SNV distances by pair type, state, and all STs (Figure S6)

```{r}
fs6 <- snv_dat %>% filter(state1 == state2 & pair_type != 'Unknown') %>% 
  mutate(state1 = factor(state1, levels = c('CT','MN','TN')),
         pair_type = ifelse(is.na(pair_type), 'Unknown', pair_type),
         st1 = gsub('^', '\\*', st1),
         st1 = gsub('e ', 'e\\* ', st1),
         st1 = gsub('i ', 'i\\* ', st1)) %>% 
  ggplot(aes(x = pairwise_dist, y = st1, color = pair_type)) + 
  geom_vline(xintercept = 10, color = 'darkgrey') +
  geom_jitter(alpha=0.25) + facet_grid(~state1, scales = 'free') +
  scale_color_manual(values = c('lightslateblue','black')) +
  labs(y = '', x = 'Pairwise SNV distance', color = '') +
  theme(axis.text.y = element_markdown())

ggsave(plot = fs6, filename = 'figures/FigS6.png', width = 7, height = 5)

fs6
```

### Patristic distance sensitivity analysis (Figure S7)

```{r}
fs7 <- left_join(snv_dat %>% rename(pd_snv = pairwise_dist), pat_dat %>% rename(pd_pat = pairwise_dist)) %>% 
  mutate(state1 = factor(state1, levels = c('CT', 'MN', 'TN')),
         pair_type = factor(gsub(' pair', '', pair_type), levels = c('Intra-facility','Inter-facility')),
         Species = gsub('^', '\\*', Species),
         Species = gsub('$', '\\*', Species)) %>% 
  filter(pair_type != 'Unknown' & ST != 'Other') %>% 
  ggplot(aes(x = pair_type, y = pd_pat)) + geom_jitter(aes(fill = under_thresh, color = under_thresh),alpha = 0.5) +
  scale_color_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
  geom_violin(position = 'identity',alpha=0, color = 'azure4') +
  facet_grid(Species+ST~state1, scales = 'free') + scale_y_log10() +
  labs(x = 'Pair type',y = 'Pairwise patristic distance', color = 'SNV bin', fill = 'SNV bin') +
  theme(strip.text.y = element_markdown())

ggsave(plot = fs7, filename = 'figures/FigS7.png', width = 6.5, height = 5)

fs7
```

## Closely related facility type pairs (Table S2)

```{r}
pr_hosp_pairs_ct <- left_join(pr_hosp_pairs_ct, snv_dat)
pr_hosp_pairs_mn <- left_join(pr_hosp_pairs_mn, snv_dat)
pr_hosp_pairs_tn <- left_join(pr_hosp_pairs_tn, snv_dat)

nodes <- bind_rows(pr_hosp_pairs_mn %>% select(loc1, state1, st1) %>% 
                     rename(Facility_ID = loc1, state = state1, st = st1),
                   pr_hosp_pairs_mn %>% select(loc2, state2, st2) %>% 
                     rename(Facility_ID = loc2, state = state2, st = st2),
                   pr_hosp_pairs_tn %>% select(loc1, state1, st1) %>% 
                     rename(Facility_ID = loc1, state = state1, st = st1),
                   pr_hosp_pairs_tn %>% select(loc2, state2, st2) %>% 
                     rename(Facility_ID = loc2, state = state2, st = st2),
                   pr_hosp_pairs_ct %>% select(loc1, state1, st1) %>% 
                     rename(Facility_ID = loc1, state = state1, st = st1),
                   pr_hosp_pairs_ct %>% select(loc2, state2, st2) %>% 
                     rename(Facility_ID = loc2, state = state2, st = st2)) %>% 
  filter(!duplicated(.)) %>% filter(!is.na(Facility_ID)) %>%
  select(-st) 
  # mutate(st = gsub('cloacae', 'hormaechei', st),
  #        st = ifelse(st %in% c('E. hormaechei ST114','E. hormaechei ST171', 'K. pneumoniae ST258', 'K. pneumoniae ST307'), st, 'Other'))

edges <- bind_rows(pr_hosp_pairs_mn %>% rename(from = loc1, to = loc2), 
          pr_hosp_pairs_tn %>% rename(from = loc1, to = loc2), 
          pr_hosp_pairs_ct %>% rename(from = loc1, to = loc2)) %>% 
  filter(pairwise_dist < 10) %>% 
  select(from, to, same_pr_hosp, st1) %>%
  filter(!is.na(to) & !is.na(from)) %>% 
  mutate(st1 = gsub('cloacae', 'hormaechei', st1),
         st1 = ifelse(st1 %in% c('E. hormaechei ST114','E. hormaechei ST171', 'K. pneumoniae ST258', 'K. pneumoniae ST307'), st1, 'Other'),
         st1 = factor(st1, levels = c('E. hormaechei ST114','E. hormaechei ST171', 
                                      'K. pneumoniae ST258', 'K. pneumoniae ST307', 'Other'),
                      labels = c('*E. hormaechei* ST114','*E. hormaechei* ST171', 
                                      '*K. pneumoniae* ST258', '*K. pneumoniae* ST307', 'Other')),
         same_pr_hosp = ifelse(same_pr_hosp, 'Yes','No'))

graph <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE) %>% 
  induced.subgraph(igraph::degree(.) > 0)
```

```{r}
left_join(edges, facil_types %>% rename(from = Facility_ID, type_from = type)) %>% 
  left_join(facil_types %>% rename(to = Facility_ID, type_to = type)) %>% 
  mutate(type_pair = sapply(1:nrow(.), function(x) paste0(sort(c(type_from[x], type_to[x])), collapse = '-')),
         type_pair = ifelse(type_pair %in% c('ACH-ACH', 'ACH-LTACH', 'ACH-SNF'), type_pair, 'Other')) %>% 
  group_by(state, st1, type_pair) %>% 
  summarize(count = n()) %>% 
  arrange(state, st1, type_pair) %>% 
  pivot_wider(names_from = type_pair, values_from = count, values_fill = 0) %>% 
  select(state, st1, `ACH-ACH`, `ACH-LTACH`, `ACH-SNF`, Other)
```

## Intra-facility analysis (Figure S8)

```{r}
min_dat <- lapply(unique(c(as.character(snv_dat$isolate1), as.character(snv_dat$isolate2))), function(i){
  snv_dat %>% filter(pair_type == 'Intra-facility pair' & (isolate1 == i | isolate2 == i)) %>% 
  group_by(loc1, ST_orig, state1) %>% 
  summarize(dist_min = min(pairwise_dist))
}) %>% bind_rows() %>% suppressMessages() %>% suppressWarnings()
```

```{r}
fs8 <- min_dat %>% 
    mutate(under_thresh = ifelse(dist_min <= 15, '(10,15]','(15,∞)'),
         under_thresh = ifelse(dist_min <= 10, '(5,10]',under_thresh),
         under_thresh = ifelse(dist_min <= 5, '[0,5]',under_thresh),
         under_thresh = factor(under_thresh, levels = rev(c('[0,5]','(5,10]','(10,15]','(15,∞)'))),
         ST = ifelse(ST_orig %in% c('K. pneumoniae ST258','E. hormaechei ST171', 'K. pneumoniae ST307','E. hormaechei ST114'), ST_orig, 'Other'),
         ST = factor(ST, levels = c('E. hormaechei ST171', 'E. hormaechei ST114', 'K. pneumoniae ST258', 'K. pneumoniae ST307', 'Other'),
                     labels = c('*E. hormaechei*<br>ST171', '*E. hormaechei*<br>ST114', 
                                '*K. pneumoniae*<br>ST258', '*K. pneumoniae*<br>ST307', 'Other')),
         state1 = factor(state1, levels = c('CT', 'MN', 'TN'))) %>% 
  ggplot(aes(x = reorder(loc1, loc1, function(x)-length(x)), fill = under_thresh)) + 
  geom_bar() +
       scale_fill_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
         labs(y = "Number of isolates", x = "Facility", fill = 'SNV bin of closest\nintra-facility isolate') +
  facet_grid(ST~state1, space = 'free', scales = 'free_x') +
  theme(axis.text.x = element_blank(), strip.text.y = element_markdown()) + 
  scale_y_continuous(breaks = scales::pretty_breaks())

ggsave(plot = fs8, filename = 'figures/FigS8.png', width = 7, height = 5)

fs8
```

## Clinical factor analysis (Table S3)

```{r}
fisher <- function(a,b,c,d){
  data <- matrix(c(a,b,c,d),ncol=2)
  c(p = fisher.test(data)$p.value,
    OR = fisher.test(data)$estimate)
}
```

```{r}
intra_pairs_ct <- snv_dat %>% filter(loc1 == loc2 & st1 == st2 & state1 == 'CT')
intra_pairs_mn <- snv_dat %>% filter(loc1 == loc2 & st1 == st2 & state1 == 'MN')
intra_pairs_tn <- snv_dat %>% filter(loc1 == loc2 & st1 == st2 & state1 == 'TN')

pt_info_ct <- suppressMessages(lapply(1:nrow(intra_pairs_ct), function(x){
  sub <- pt_feats_ct %>% filter(Sample_ID %in% unlist(c(intra_pairs_ct[x,c('isolate1','isolate2')]))) %>% 
  select_if(colSums(is.na(.)) == 0)
  info <- bind_cols(intra_pairs_ct[x,])%>% mutate(shared_clin = FALSE)
  if(ncol(sub) > 2 & nrow(sub) > 1){
    info <- info %>% mutate(shared_clin = TRUE)
  }
  info
})) %>% bind_rows()

pt_info_mn <- suppressMessages(lapply(1:nrow(intra_pairs_mn), function(x){
  sub <- pt_feats_mn %>% filter(Sample_ID %in% unlist(c(intra_pairs_mn[x,c('isolate1','isolate2')]))) %>% 
  select_if(colSums(is.na(.)) == 0)
  info <- bind_cols(intra_pairs_mn[x,])%>% mutate(shared_clin = FALSE)
  if(ncol(sub) > 2 & nrow(sub) > 1){
    info <- info %>% mutate(shared_clin = TRUE, shared_comorb = paste0(names(sub)[!names(sub) == 'Sample_ID'], collapse = ','))
  }
  info
})) %>% bind_rows()

pt_info_tn <- suppressMessages(lapply(1:nrow(intra_pairs_tn), function(x){
  sub <- pt_feats_tn %>% filter(Sample_ID %in% unlist(c(intra_pairs_tn[x,c('isolate1','isolate2')]))) %>% 
    unique() %>% 
    select_if(c(TRUE, colSums(.[,2:ncol(.)]) != 0))
  info <- bind_cols(intra_pairs_tn[x,]) %>% mutate(shared_clin = FALSE)
  if(ncol(sub) > 2 & nrow(sub) > 1){
    info <- info %>% mutate(shared_clin = TRUE, shared_comorb = paste0(names(sub)[!names(sub) == 'Sample_ID'], collapse = ','))
  }
  info
})) %>% bind_rows()

pt_info_comb <- bind_rows(pt_info_mn, pt_info_tn, pt_info_ct) %>%
  mutate(`pairwise_dist <= 10` = pairwise_dist <= 10) %>% 
  select(state1, st1, shared_clin, `pairwise_dist <= 10`)

pt_comorb_compare <- pt_info_comb %>% 
  group_by(state1, st1, shared_clin, `pairwise_dist <= 10`) %>% summarize(count = n()) %>% 
  mutate(sharedclin_pdlt10 = factor(paste0(ifelse(shared_clin, 'shared', 'not shared'), ', ', 
                                           ifelse(`pairwise_dist <= 10`, '≤10', '>10')),
                                    levels = c('shared, ≤10', 'shared, >10', 'not shared, ≤10', 'not shared, >10'))) %>% 
  ungroup() %>% 
  arrange(sharedclin_pdlt10) %>% 
  select(-c(shared_clin, `pairwise_dist <= 10`)) %>% 
  pivot_wider(names_from = sharedclin_pdlt10, values_from = count, values_fill = 0) %>% 
  group_by(state1, st1) %>% 
  rowwise() %>% 
  mutate(fe_p = fisher(`shared, ≤10`, `shared, >10`, `not shared, ≤10`, `not shared, >10`)[[1]]) %>% 
  arrange(state1, st1) %>% 
  mutate(frac_shared_lt10 = `shared, ≤10`/(`shared, ≤10` + `not shared, ≤10`),
         frac_shared_gt10 = `shared, >10`/(`shared, >10` + `not shared, >10`))

pt_comorb_compare
```

## Inter-facility analysis

### Aggregate patient transfer (Figure S9)

```{r}
mn_pt_trans <- get_patient_flow(mn_trans_net_sub, 
                                unique(facil_st$Facility_ID)[unique(facil_st$Facility_ID) %in% 
                                                        c(mn_trans_net_sub$source_facil,
                                                          mn_trans_net_sub$dest_facil)]) %>% 
  filter(!duplicated(.))
tn_pt_trans <- get_patient_flow(tn_trans_net_sub, unique(facil_st$Facility_ID)[unique(facil_st$Facility_ID) %in% 
                                                        c(tn_trans_net_sub$source_facil,
                                                          tn_trans_net_sub$dest_facil)]) %>% 
  filter(!duplicated(.))
ct_pt_trans <- get_patient_flow(ct_trans_net_sub, unique(facil_st$Facility_ID)[unique(facil_st$Facility_ID) %in% 
                                                        c(ct_trans_net_sub$source_facil,
                                                          ct_trans_net_sub$dest_facil)]) %>% 
  filter(!duplicated(.))

sts <- c('ST171','ST258','ST114','ST307')

snv_by_st <- suppressWarnings(lapply(sts, function(x){
  snvdat_sub <- snv_dat %>% filter(ST == x) %>% mutate(pairwise_dist = as.numeric(pairwise_dist))
  snvdat_sub <- snvdat_sub[,1:6] 
    mn_snv_by_st_sub <- summarize_pairs(snvdat_sub) %>% mutate(ST = x)
})) %>% bind_rows()

pt_snv <- bind_rows(bind_cols(state = 'MN', mn_pt_trans), 
                    bind_cols(state = 'TN', tn_pt_trans),
                    bind_cols(state = 'CT', ct_pt_trans)) %>% inner_join(snv_by_st) %>%
  mutate(under_thresh = ifelse(dist_min <= 15, '(10,15]','(15,∞)'),
         under_thresh = ifelse(dist_min <= 10, '(5,10]',under_thresh),
         under_thresh = ifelse(dist_min <= 5, '[0,5]',under_thresh),
         under_thresh = factor(under_thresh, levels = c('[0,5]','(5,10]','(10,15]','(15,∞)')),
         species = ifelse(ST %in% c('ST258','ST307'), 'K. pneumoniae','E. hormaechei'))
```

```{r}
pvals <- pt_snv %>% 
  mutate(leq_10_bool = ifelse(dist_min <= 10, 'close','not_close')) %>% 
  select(state, ST, leq_10_bool, sum_pt_trans_metric) %>% group_by(state, ST, leq_10_bool) %>% 
  summarize(sum_pt_trans_metric = list(sum_pt_trans_metric)) %>% 
  spread(leq_10_bool, sum_pt_trans_metric) %>% 
  group_by(state, ST) %>% 
  filter(!(is.null(unlist(close)) | is.null(unlist(not_close)))) %>% 
  mutate(close_median=median(unlist(close)),
         not_close_median=median(unlist(not_close)),
         p=wilcox.test(unlist(close),unlist(not_close))$p.value) %>% 
  mutate(species = ifelse(ST %in% c('ST258','ST307'), '*K. pneumoniae*','*E. hormaechei*'))

fs9 <- pt_snv %>% 
  mutate(species = gsub('^', '\\*',species),
         species = gsub('$', '\\*',species),
         sum_pt_trans_metric = sum_pt_trans_metric + min(sum_pt_trans_metric[sum_pt_trans_metric != 0]/2), # no infinite values in transformation
         dist_min = dist_min + min(dist_min[dist_min != 0]/2)) %>%  # no infinite values in transformation
  ggplot(aes(x = sum_pt_trans_metric, y = dist_min)) + 
  geom_point(aes(color = under_thresh)) + scale_y_log10() + scale_x_log10() +
  scale_color_manual(
    values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")
    ) +
  geom_text(data = pvals, aes(x = 5e-5, y = 1, label = paste0('p=',signif(p, 1))), hjust = 0) +
  facet_grid(species+ST~state) + 
  labs(x = 'Sum(patient flow between facility pair)', y = 'Minimum pairwise SNV distance', color = 'SNV bin') +
  theme(strip.text.y = element_markdown())

ggsave(plot = fs9, filename = 'figures/FigS9.png', width = 7, height = 4)

fs9
```

### Prior hospital

#### Individual healthcare exposures (Table 2)

```{r}
pr_summary <- bind_rows(pr_hosp_pairs_ct, pr_hosp_pairs_mn, pr_hosp_pairs_tn) %>%
  left_join(snv_dat %>% filter(loc1 != loc2)) %>% 
  mutate(st1 = ifelse(st1 %in% c("E. hormaechei ST114", "E. hormaechei ST171", "K. pneumoniae ST258", 
"K. pneumoniae ST307"), st1, 'Other')) %>% 
  group_by(state1, st1, same_pr_hosp, pairwise_dist <= 10) %>% 
  summarize(total_l10 = n()) %>% 
  mutate(same_pr_hosp = ifelse(same_pr_hosp, 'pr_hosp','no_pr_hosp'),
         `pairwise_dist <= 10` = ifelse(`pairwise_dist <= 10`, 'lt10','gt10')) %>% 
  pivot_wider(names_from = c(same_pr_hosp, `pairwise_dist <= 10`), values_from = c(total_l10)) %>% 
  rename(state = state1, st = st1) %>% 
  select(state, st, pr_hosp_lt10, no_pr_hosp_lt10, pr_hosp_gt10, no_pr_hosp_gt10) %>% 
  replace(is.na(.), 0)

pr_summary$fisher_p <- sapply(1:nrow(pr_summary), function(x){
  exact2x2(matrix(unlist(c(pr_summary[x,3:6])), nrow = 2))$p.value %>% signif(digits = 1)
})

pr_summary <- pr_summary %>% 
  mutate(pr_hosp_lt10_pct = round(pr_hosp_lt10/(no_pr_hosp_lt10 + pr_hosp_lt10), 1)*100, .after = no_pr_hosp_lt10) %>% 
  mutate(pr_hosp_gt10_pct = round(pr_hosp_gt10/(no_pr_hosp_gt10 + pr_hosp_gt10), 1)*100, .after = no_pr_hosp_gt10) %>% 
  mutate('Pairwise SNV distance ≤ 10\nwith shared facility exposure (%)' =
           paste0(pr_hosp_lt10, '/', (pr_hosp_lt10+no_pr_hosp_lt10), ' (',round(pr_hosp_lt10/(pr_hosp_lt10+no_pr_hosp_lt10),2)*100,')'),
         'Pairwise SNV distance > 10\nwith shared facility exposure (%)' =
           paste0(pr_hosp_gt10, '/', (pr_hosp_gt10+no_pr_hosp_gt10), ' (',round(pr_hosp_gt10/(pr_hosp_gt10+no_pr_hosp_gt10),2)*100,')')) %>% 
  select(c("state", "st", 
           "Pairwise SNV distance ≤ 10\nwith shared facility exposure (%)", 
           "Pairwise SNV distance > 10\nwith shared facility exposure (%)",
           "fisher_p"
)) %>% 
  rename(State = state, ST = st, "Fisher's exact p" = fisher_p)

pr_summary
```

### Individual graph (Figure S10)

```{r}
fs10 <- graph %>% ggraph(layout = 'kk') + 
  geom_edge_parallel(aes(color = st1, linetype = same_pr_hosp, alpha = same_pr_hosp),sep = unit(0.4, "mm")) + 
  geom_node_point(size = 0.5) + 
  scale_edge_linetype_manual(values = c('dotted','solid')) +
  facet_nodes(~state, nrow = 3) +
  scale_edge_alpha_manual(values = c(0.5, 1)) + 
  scale_edge_color_manual(values = c('*E. hormaechei* ST171'="aquamarine4", '*E. hormaechei* ST114'="aquamarine3",
                                     '*K. pneumoniae* ST258'="coral4", '*K. pneumoniae* ST307'="coral3",
                                     'Other'='grey'), 
                          breaks = c('*E. hormaechei* ST171', '*E. hormaechei* ST114',
                                      '*K. pneumoniae* ST258', '*K. pneumoniae* ST307', 'Other'), drop = FALSE) +
  labs(edge_color = 'Isolate pair ≤ 10 SNVs', edge_linetype = 'Patients have\nshared facility exposure',
       edge_alpha = 'Patients have\nshared facility exposure') +
  theme_graph() + theme(legend.text = element_markdown())

ggsave(plot = fs10, filename = 'figures/FigS10.png', width = 5, height = 6)

fs10
```

### Comparison of individual to aggregate (Figure S11)

```{r}
fs11 <- full_join(pr_summary %>% mutate(ST_orig = ST, ST = gsub('.* ','',ST)) %>% rename('ind_p' = "Fisher's exact p"), 
          pvals %>% rename(State = state, 'agg_p' = 'p')) %>% 
  mutate(ST_orig = gsub('^', '\\*', ST_orig),
         ST_orig = gsub('e ', 'e\\* ', ST_orig),
         ST_orig = gsub('i ', 'i\\* ', ST_orig)) %>% 
  filter(ST != 'Other') %>% 
  ggplot(aes(x = ind_p, y = agg_p, col = State, shape = ST_orig)) +
  geom_point(size = 3) +
  scale_color_manual(values = c(CT = "#46f0f0", MN = "#3cb44b", TN = "#4363d8")) +
  geom_hline(yintercept = 0.05) +
  geom_vline(xintercept = 0.05) +
  labs(x = 'Patient transfer p-value from\nindividual-level analysis', 
       y = 'Patient transfer p-value from\naggregate-level analysis',
       shape = 'ST') +
    theme(legend.text = element_markdown())

ggsave(plot = fs11, filename = 'figures/FigS11.png', width = 5, height = 3)

fs11
```

## MN maps (Figure 3)

```{r}
# connecting each patient to their closest isolate with a subsequent collection date
edge_dat <- lapply(unique(c(as.character(snv_dat$isolate1), as.character(snv_dat$isolate2))), function(i){
  cultdate <- facil_st$Cult_Date[facil_st$Sample_ID == i]
  subset_dat <- snv_dat %>% filter(pair_type == 'Inter-facility pair' & (isolate1 == i | isolate2 == i) & state1 == 'MN')
  dates <- sapply(unique(c(as.character(subset_dat$isolate1), as.character(subset_dat$isolate2))), function(j){
    as.character(facil_st$Cult_Date[facil_st$Sample_ID == j])
  }) 
  dat <- NULL
  if(length(dates) > 0){
    keepers <- names(dates)[as.Date(dates) > cultdate]
    dat <- subset_dat %>% 
      filter(isolate1 %in% keepers | isolate2 %in% keepers) %>% 
      filter(pairwise_dist == min(pairwise_dist)) %>% 
      mutate(earlier = as.character(ifelse(isolate2 == i, 'loc2','loc1')),
             date1 = dates[as.character(isolate1)],
             date2 = dates[as.character(isolate2)],
             datediff = abs(as.numeric(as.Date(dates[as.character(isolate1)]) - as.Date(dates[as.character(isolate2)]))))
  }
}) %>% bind_rows()

edges_for_plot <- edge_dat %>% 
  filter(pairwise_dist <= 15) %>% 
  left_join(facil_st %>% select(Sample_ID, Facility_ID, longitude, latitude), 
            by = c('isolate1' = 'Sample_ID', 'loc1' = 'Facility_ID')) %>%
  rename(x = longitude, y = latitude) %>%
  left_join(facil_st %>% select(Sample_ID, Facility_ID, longitude, latitude), 
            by = c('isolate2' = 'Sample_ID', 'loc2' = 'Facility_ID')) %>%
  rename(xend = longitude, yend = latitude) %>% 
  filter(!(is.na(x) | is.na(xend)) & ST %in% c('ST171','ST258')) %>% 
  mutate(dist = abs(x - xend)^2 + abs(y - yend)^2)

facilities <- facil_st %>% filter(State == 'MN' & ST %in% c('E. hormaechei ST171', 'K. pneumoniae ST258')) %>% 
  select(Facility_ID, latitude, longitude) %>% 
  filter(!duplicated(.) & !is.na(latitude)) 
```

```{r}
edges_for_plot_ec <- edges_for_plot %>% filter(ST == 'ST171') %>% 
  mutate(under_thresh = factor(under_thresh, levels = c("[0,5]", "(5,10]", "(10,15]")))

whole_state_ec <- facil_st %>% 
  mutate(ST = gsub('cloacae', 'hormaechei', ST)) %>% 
  filter(ST %in% c('E. hormaechei ST171')) %>%
  mutate(ST = gsub('.* ST', 'ST', ST)) %>% 
  group_by(Facility_ID, latitude, longitude, ST)  %>% tally() %>%
  ggplot(aes(x=longitude, y=latitude), color = "grey99") +
    geom_curve(aes(x = xend, y = yend, xend = x, yend = y, color = under_thresh),     # draw edges as arcs
                 alpha = 1,
             data = edges_for_plot_ec, curvature = 0.33) + 
  scale_color_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4")) +
  geom_point(data = facilities, aes(x = longitude, y = latitude), alpha = 0.25, size = 1) +
  geom_point(aes(size=n), alpha = 1, color = 'white') +
  geom_point(aes(size=n), fill = adjustcolor(st_cols[c('E. hormaechei ST171')], alpha.f = 0.5), pch = 21) + 
  scale_size_continuous(breaks = seq(0, 20, 5), limits = c(0,25),
                        labels = c(0,5,10,15,20),
                        name = 'Number of\nisolates') +
  coord_quickmap() +
  # geom_rect(aes(xmin = -986875.33, xmax = -986875.165, ymin = 896963.81, ymax = 896964.1),
  geom_rect(aes(xmin = -986875.33+0.58, xmax = -986875.165+0.58, ymin = 896963.81+0.145, ymax = 896964.1+0.145),
  fill = "transparent", color = "lightgrey", size = 1) +
  labs(color = 'SNV bin', size = 'Number of\nisolates') +
  guides(size = guide_legend(override.aes = list(alpha = c(0.25, 1, 1, 1, 1), fill = 'black'))) +
  theme_void() 
whole_state_ec

zoom_state_ec <- whole_state_ec + 
  # coord_quickmap(xlim=c(-986875.33+0.007, -986875.165-0.007), ylim=c(896963.81+0.012, 896964.1-0.012)) +
  coord_quickmap(xlim=c(-986875.33+0.58+0.007, -986875.165+0.58-0.007), ylim=c(896963.81+0.145+0.013, 896964.1+0.145-0.013)) +
  theme(legend.position = "none") + 
  geom_text(aes(label = ifelse(Facility_ID == 'F38', 'W', NA)),hjust=1.5, vjust=2)
zoom_state_ec
```

```{r}
edges_for_plot_kp <- edges_for_plot %>% filter(ST == 'ST258')
whole_state_kp <- facil_st %>% 
  filter(ST %in% c('K. pneumoniae ST258')) %>%
  mutate(ST = gsub('.* ST', 'ST', ST)) %>% 
  group_by(Facility_ID, latitude, longitude, ST)  %>% tally() %>%
  ggplot(aes(x=longitude, y=latitude), color = "grey99") +
    geom_curve(aes(x = xend, y = yend, xend = x, yend = y, color = under_thresh),     # draw edges as arcs
                 alpha = 1,
             data = edges_for_plot_kp, curvature = 0.33) + 
  scale_color_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
  geom_point(data = facilities, aes(x = longitude, y = latitude), size = 1, alpha = 0.25) +
  geom_point(aes(size=n), alpha = 1, color = 'white') +
  geom_point(aes(size=n), fill = adjustcolor(st_cols['K. pneumoniae ST258'], alpha.f = 0.5), pch = 21) + 
  scale_size_continuous(breaks = seq(5, 20, 5), limits = c(1,25)) +
  coord_quickmap() +
  # geom_rect(aes(xmin = -986875.33, xmax = -986875.165, ymin = 896963.81, ymax = 896964.1),
  geom_rect(aes(xmin = -986875.33+0.58, xmax = -986875.165+0.58, ymin = 896963.81+0.145, ymax = 896964.1+0.145),
               fill = "transparent", color = "lightgrey", size = 1) +
  labs(size = 'Number of\nisolates', color = 'SNV bin') +
  theme_void() 
whole_state_kp

zoom_state_kp <- whole_state_kp + 
  # coord_quickmap(xlim=c(-986875.33+0.007, -986875.165-0.007), ylim=c(896963.81+0.012, 896964.1-0.012)) +
  coord_quickmap(xlim=c(-986875.33+0.58+0.007, -986875.165+0.58-0.007), ylim=c(896963.81+0.145+0.013, 896964.1+0.145-0.013)) +
  theme(legend.position = "none") +
   geom_text(aes(label = ifelse(Facility_ID == 'F38', 'W', NA)),hjust=1.5, vjust=2)
zoom_state_kp
```

```{r}
# grob plotting function to plot panel colors
plot_panel_cols <- function(p, fills){
  gt <- ggplot_gtable(ggplot_build(p))
  strip_both <- which(grepl('strip-', gt$layout$name))
  k <- 1
  for (i in strip_both) {
    j <- which(grepl('rect', gt$grobs[[i]]$grobs[[1]]$childrenOrder))
    gt$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
    k <- k+1
  }
  ggplotify::as.ggplot(gt)
}
```

```{r}
bp <- edge_dat %>% pivot_longer(cols = c('loc1','loc2'), names_to = 'fname', values_to = 'loc') %>% 
  filter(ST %in% c('ST171','ST258') & under_thresh != '(15,∞)') %>%
  mutate(under_thresh = factor(under_thresh, levels = rev(levels(under_thresh))),
         ST = factor(ifelse(ST == 'ST171', '*E. hormaechei* ST171', '*K. pneumoniae* ST258'),
                     levels = c('*E. hormaechei* ST171', '*K. pneumoniae* ST258'))) %>% 
  filter(fname == earlier) %>% 
  select(loc, ST, under_thresh) %>%
  ggplot(aes(x = loc)) + 
    geom_bar(aes(col = ifelse(loc == 'F38', 'yes','no'))) +
  geom_bar(aes(fill = under_thresh)) + 
  facet_grid(ST~.) +
    scale_fill_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
  scale_color_manual(values = c('yes' = 'black','no' = NA)) +
  labs(x = 'Facility', y = 'Number of isolates closely related\nto an isolate from another facility', fill = 'SNV bin')

bp_labeled <- bp + scale_x_discrete(labels = ifelse(ggplot_build(bp)$layout$panel_params[[1]]$x$breaks == 'F38', 'W','')) +
  theme_classic() +
  theme(legend.position = 'none', #axis.ticks.x = element_blank(), 
        text = element_text(size = 10), 
        axis.text.x = element_text(angle = 0),
        strip.background = element_rect(fill="white",linetype='blank'),
        strip.text.y = element_markdown())
bp_col <- plot_panel_cols(bp_labeled, adjustcolor(st_cols[c('E. hormaechei ST171','K. pneumoniae ST258')], alpha.f = 0.5)) 
```

```{r}
leg <- ggpubr::get_legend(whole_state_ec +
                            theme(legend.box.margin = margin(0, 0, 0, 50))
)
f3 <- plot_grid(leg, NULL,
          plot_grid(whole_state_ec + theme(legend.position = 'none'), zoom_state_ec,
                    NULL, NULL,
                    whole_state_kp + theme(legend.position = 'none'), zoom_state_kp,
                    NULL, NULL,
                    nrow = 4, 
                    rel_heights = c(1,0.05,1,0.1,1,0.05,1,0.1)),
          plot_grid(NULL, bp_col, NULL, ncol = 1, rel_heights = c(0.1,0.8,0.05)), nrow = 1,
          rel_widths = c(0.2, 0.05, 1, 0.8) 
)

ggsave(plot = f3, filename = 'figures/Fig3.png', width = 10, height = 4)

f3
```

```{r}
# prior facilities for facility w/ st171 isolates 
facil_w %>% 
  mutate(n_patients_tot = n_distinct(Sample_ID)) %>% 
  filter(!is.na(prior_hosp_alias)) %>% 
  unique() %>% 
  group_by(prior_hosp_alias, n_patients_tot) %>% 
  summarize(n_patients = n()) %>% 
  ungroup() %>% 
  mutate(prop_patients = n_patients/n_patients_tot) %>% 
  arrange(-n_patients) 
```

### Comparison of ST171 and ST258 inter-facility transmission patterns (Figure S12)

```{r}
## for each facility, how many times it's involved in a link

frac_ec <- lapply(unique(edges_for_plot_ec$loc1, edges_for_plot_ec$loc2), function(x){
  edges_for_plot_ec %>% filter(pairwise_dist <= 10) %>% 
  mutate(conn = ifelse(loc1 == !!x | loc2 == !!x, 'connected', 'not')) %>% 
  group_by(conn, ST) %>% summarize(count = n()) %>% mutate(facil = x)
}) %>% bind_rows() %>% suppressMessages() %>% ungroup() %>% filter(!duplicated(.)) %>% 
  group_by(facil) %>% 
  mutate(n_facil = sum(count), frac = count/n_facil) %>% 
  ungroup() %>% 
  filter(conn == 'connected')

frac_kp <- lapply(unique(edges_for_plot_kp$loc1, edges_for_plot_kp$loc2), function(x){
  edges_for_plot_kp %>% filter(pairwise_dist <= 10) %>% 
  mutate(conn = ifelse(loc1 == !!x | loc2 == !!x, 'connected', 'not')) %>% 
  group_by(conn, ST) %>% summarize(count = n()) %>% mutate(facil = x)
}) %>% bind_rows() %>% suppressMessages() %>% ungroup() %>% filter(!duplicated(.)) %>% 
  group_by(facil) %>% 
  mutate(n_facil = sum(count), frac = count/n_facil) %>% 
  ungroup() %>% 
  filter(conn == 'connected')

max(frac_ec$frac)
facils_ec <- frac_ec %>% pull(facil)
ec_perm_cts <- lapply(1:1000, function(x){
  sample(facils_ec, sum(frac_ec$count), replace = TRUE) %>% table() %>% data.frame() %>% rename(facil = '.') %>% 
    mutate(tot = sum(Freq), perm = x, frac = Freq/tot) %>% slice_max(Freq) %>% slice(1)
}) %>% bind_rows()

max(frac_kp$frac)
facils_kp <- frac_kp %>% pull(facil)
kp_perm_cts <- lapply(1:1000, function(x){
  sample(facils_kp, sum(frac_kp$count), replace = TRUE) %>% table() %>% data.frame() %>% rename(facil = '.') %>% 
    mutate(tot = sum(Freq), perm = x, frac = Freq/tot) %>% slice_max(Freq) %>% slice(1)
}) %>% bind_rows()

fs12 <- bind_rows(ec_perm_cts %>% select(frac) %>% mutate(ST = 'ST171'),
          kp_perm_cts %>% select(frac) %>% mutate(ST = 'ST258')) %>% 
  mutate(ST = factor(ifelse(ST == 'ST171', '*E. hormaechei* ST171', '*K. pneumoniae* ST258'),
                     levels = c('*E. hormaechei* ST171', '*K. pneumoniae* ST258'))) %>% 
  ggplot(aes(x = frac)) +
  geom_histogram(binwidth = 0.05) +
  geom_point(data = tibble(frac = c(max(frac_ec$frac), max(frac_kp$frac)), 
                           st = c('*E. hormaechei* ST171', '*K. pneumoniae* ST258')), 
             aes(y = 0), col = 'red', shape = 18, size = 5) +
  facet_grid(~st, scales = 'free') +
  labs(x = 'Maximum proportion of edges connected to a given facility', y = 'Number of edges') +
  theme(strip.text.x = element_markdown())

sum(ec_perm_cts$frac > max(frac_ec$frac))/length(ec_perm_cts$frac)
sum(kp_perm_cts$frac > max(frac_kp$frac))/length(kp_perm_cts$frac)

fs12 <- plot_panel_cols(fs12, adjustcolor(st_cols[c('E. hormaechei ST171','K. pneumoniae ST258')], alpha.f = 0.5))

# ggsave(plot = fs12, filename = 'figures/FigS12.png', width = 7, height = 4)

fs12
```

### Facility W (Figure S13)

```{r}
g_mn <- mn_trans_net_sub %>% 
  select(source_facil, dest_facil) %>%
  as.matrix() %>% 
  graph_from_edgelist(directed = TRUE)

E(g_mn)$weight = mn_trans_net_sub %>% select(n_transfers) %>% deframe()

deg <- igraph::degree(g_mn)
stre <- strength(g_mn)
betw <- betweenness(g_mn)
clos <- closeness(g_mn, mode = 'all')

facil_w_ind <- which(names(V(g_mn)) == 'F38')

fs13 <- bind_cols(inds=names(deg), 'Degree'=deg, 'Strength'=stre, 'Betweenness'=betw, 'Closeness'=clos) %>% 
  pivot_longer(cols = -inds) %>% 
  ggplot(aes(x = value)) + geom_histogram(bins = 70) +
  geom_point(data = data.frame(
    value=c(deg[facil_w_ind], stre[facil_w_ind], betw[facil_w_ind],clos[facil_w_ind]),
    name=c('Degree','Strength','Betweenness','Closeness')),
    aes(x = value, y = 0), col = 'red', shape = 18, size = 3) +
  ggrepel::geom_text_repel(data = data.frame(
    value=c(deg[facil_w_ind], stre[facil_w_ind], betw[facil_w_ind],clos[facil_w_ind]), 
    name=c('Degree','Strength','Betweenness','Closeness')), 
    aes(label = 'W', y = 0), col = 'red', size = 3) +
  labs(x = 'Centrality metric', y = 'Number of facilities') +
  facet_wrap(~name, scales = 'free') +
  theme(plot.margin = margin(0,0.5,0,0, "cm"))

ggsave(plot = fs13, filename = 'figures/FigS13.png', width = 5, height = 3.5)

fs13
```

## TN facility cluster (Figure 4)

```{r}
prior_hosps_tn_counts <- prior_hosps_tn %>% 
  select(Sample_ID, Fac_ID) %>% 
  unique() %>% 
  rename(Facility = Fac_ID) %>% 
  bind_rows(prior_hosps_tn %>% 
              select(Sample_ID, Prior_facil) %>% 
              rename(Facility = Prior_facil)) %>% 
  left_join(facil_st %>% 
              mutate(ST = ifelse(ST %in% c('E. hormaechei ST171','E. hormaechei ST114',
                                 'K. pneumoniae ST258','K. pneumoniae ST307', 
                                 'E. cloacae ST171', 'E. cloacae ST114', 
                                 'ST258', 'ST307', 'ST114', 'ST171'),
                       ST, 'Other'))) %>% 
  filter(!is.na(ST)) %>% 
  group_by(ST, Facility) %>% 
  summarize(count = n()) %>% 
  arrange(-count)

tn_hm <- prior_hosps_tn_counts %>% 
  pivot_wider(names_from = Facility, values_from = count, values_fill = 0) %>% 
  data.frame(row.names = .$ST) %>% 
  select(-ST) %>% 
  pheatmap(show_colnames = FALSE, fontsize_col = 8)

tn_facil_hm <- prior_hosps_tn_counts %>% 
  mutate(Facility = factor(Facility, levels = rev(tn_hm$tree_col$labels[tn_hm$tree_col$order]) %>% gsub('TDH.', 'TDH-', .) %>% gsub('LONG.', 'LONG-', .) %>% gsub('VA.', 'VA-', .) %>% gsub('\\.', ' ', .)),
         ST = factor(ST, levels = tn_hm$tree_row$labels),
         ST = gsub('E. hormaechei', '*E. hormaechei*', ST),
         ST = gsub('K. pneumoniae', '*K. pneumoniae*', ST)) %>% 
  ggplot(aes(x = Facility, y = ST, fill = count)) +
  geom_tile() +
  geom_rect(aes(xmin = 0+0.5, xmax = 7+0.5, ymin = 0+0.5, ymax = 3+0.5), fill = NA, col = 'black', size = 1, linetype = 'dotted') +
  scale_fill_viridis_c(direction = -1, breaks = scales::pretty_breaks(), option = 'magma') +
  theme_classic() +
  theme(axis.text.x = element_blank(), axis.text.y = element_markdown()) +
  labs(fill = 'Count')

tn_facil_hm
```

```{r}
tn_facils_st <- facil_st %>% filter(State == 'TN') %>% 
  mutate(ST = ifelse(ST %in% c('E. hormaechei ST171','E. hormaechei ST114',
                                 'K. pneumoniae ST258','K. pneumoniae ST307', 
                                 'E. cloacae ST171', 'E. cloacae ST114', 
                                 'ST258', 'ST307', 'ST114', 'ST171'),
                       ST, 'Other')) %>% 
  filter(!is.na(Facility_ID)) %>% 
  group_by(Facility_ID, ST) %>% 
  summarize(count = n()) %>% 
  arrange(Facility_ID) %>% 
  group_by(Facility_ID) %>% 
  mutate(max_count = count[which.max(count)],
         n_isolates = sum(count)) %>% 
  group_by(Facility_ID, n_isolates) %>% 
  summarize(max_sts = str_c(ST[count == max_count], collapse = ', ')) %>% 
  left_join(tn_facil_cluster) %>% 
  mutate(facil_st_type = case_when(max_sts == 'K. pneumoniae ST258' ~ 'K. pneumoniae ST258',
                                   grepl('K. pneumoniae ST258', max_sts) & max_sts != 'K. pneumoniae ST258' ~ 'K. pneumoniae ST258 & Other',
                                   of_interest ~ 'ST307/ST114/Other facility cluster'),
         facil_st_type = ifelse(is.na(facil_st_type), 'Other', facil_st_type)) 
```

```{r}
graph_tn <- tbl_graph(nodes = tn_facils_st, edges = tn_trans_net_sub %>% filter(!grepl('I', source_facil) & !grepl('I', dest_facil) & n_transfers > 1), directed = TRUE) %>%
  induced.subgraph(igraph::degree(.) > 0)

tn_network <- graph_tn %>% ggraph(layout = 'kk') + 
  geom_edge_link(aes(color = n_transfers, alpha = n_transfers), arrow = arrow(length = unit(3, 'mm'))) + 
  geom_node_point(aes(color = facil_st_type, size = n_isolates)) + 
  scale_edge_color_viridis(direction = -1, option = 'E') +
  scale_color_manual(values = c(unname(st_cols[1]), 'coral3', 'lightsalmon', 'grey')) +
  guides(edge_alpha = 'none') +
  theme_graph() +
  labs(color = 'Most common ST', shape = '', size = 'Number of isolates',
       edge_color = 'Number of transfers (in thousands)')  #+
  # theme(legend.text = element_markdown())
tn_network
```

```{r}
tn_facil_transfers <- tn_trans_net_sub %>% 
  left_join(tn_facils_st %>% rename(source_of_interest = of_interest) %>% 
              select(Facility_ID, source_of_interest), 
            by = c(source_facil = 'Facility_ID')) %>% 
  left_join(tn_facils_st %>% rename(dest_of_interest = of_interest) %>% 
              select(Facility_ID, dest_of_interest), 
            by = c(dest_facil = 'Facility_ID')) %>% 
  replace(is.na(.), 0) %>% 
  mutate(of_interest = source_of_interest + dest_of_interest) %>% 
  filter(!grepl('I', source_facil) & !grepl('I', dest_facil) & n_transfers > 1) %>%
  ggplot(aes(x = n_transfers, linetype = factor(of_interest, levels = c(2, 1, 0)))) +
  geom_density() +
  labs(x = 'Number of patient transfers', y = 'Density', linetype = 'Number of isolates\nof pair in facility\ncluster') +
  theme(legend.position = 'left')

tn_facil_transfers
```

```{r}
plot_grid(plot_grid(tn_facil_hm + theme(legend.position = 'left'), tn_facil_transfers, nrow = 2, align = 'hv', labels = c('A', 'C')), tn_network, nrow = 1, rel_widths = c(0.8, 1), labels = c('', 'B')) %>% 
  ggsave(filename = 'figures/Fig4.png', width = 12, height = 5)
```

### 2022-09-13 code/get_transition_edges.R

@@ -1,0 +1,160 @@
+# from hogwash
+
+is_tip <- function(node_num, tr){
+  return(node_num <= ape::Ntip(tr))
+}
+
+identify_transition_edges <- function(tr, mat, num, node_recon, disc_cont, same_thresh = 1, disc_bin = TRUE){
+
+  # FUNCTION -------------------------------------------------------------------
+  transition <- transition_direction <-
+    parent_node <- child_node <- integer(ape::Nedge(tr))
+  older <- 1 # older node is 1st column in tr$edge
+  younger <- 2 # younger node is 2nd column in tr$edge
+  parent_0_child_1 <- 1
+  parent_1_child_0 <- -1
+  parent_equals_child <- 0
+  both_parent_and_child_are_one <- 2
+
+  for (i in 1:ape::Nedge(tr)) {
+    if (is_tip(tr$edge[i, older], tr)) {
+      stop("tree invalid")
+    }
+    parent_node[i] <- node_recon[tr$edge[i, older] - ape::Ntip(tr)]
+    if (is_tip(tr$edge[i, younger], tr)) {
+      # child is a tip
+      child_node[i]  <- mat[, num][tr$edge[i, younger]]
+    } else {
+      # child is internal nodes
+      child_node[i]  <- node_recon[tr$edge[i, younger] - ape::Ntip(tr)]
+    }
+    if(!disc_bin){
+      nodes_diff = parent_node[i] != child_node[i]
+      if(is.na(parent_node[i]) | is.na(child_node[i])){
+        transition[i] <- NA
+      }else if(!nodes_diff){
+        transition[i] <- parent_node[i]
+      }else{
+        transition[i] <- 'trans'#paste0(parent_node[i],'-',child_node[i])
+
+      }
+      transition_direction[i] <- NA
+      next
+    }
+    transition[i] <- sum(parent_node[i] + child_node[i])
+    # transition[i] is either 0, 1, or 2 for discrete traits
+    # transition[i] is not to be used when the trait is continuous because all,
+    # or very nearly all edges are transition edges.
+
+    if (parent_node[i] > child_node[i]) {
+      transition_direction[i] <- parent_1_child_0
+    } else if (parent_node[i] < child_node[i]) {
+      transition_direction[i] <- parent_0_child_1
+    }
+    if(disc_cont == "continuous"){
[…]

### 2022-09-13 manuscript-figures.Rmd
Add supplementary file

@@
 ```
+
+## Supplemental file with isolate IDs and information
+
+```{r}
+sample_metadat <- read_csv('data/eip-metadat-short.csv') %>% 
+  rename(Pat_ID = Patient_ID) %>% 
+  mutate(Species = gsub('cloacae', 'hormaechei', Species),
+         ST = gsub('cloacae', 'hormaechei', ST)) %>% 
+  left_join(facil_st) %>% 
+  select(-c(latitude, longitude))
+
+all_sample_info_not_tn <- sample_metadat %>% 
+  select(Fac_ID, Facility_ID) %>% 
+  filter(!is.na(Facility_ID)) %>% 
+  unique() %>% 
+  right_join(sample_metadat %>% select(-Facility_ID)) %>% 
+  select(-Patient_ID) %>% 
+  left_join(sample_metadat %>% 
+              select(Pat_ID, Patient_ID) %>% 
+              filter(!is.na(Patient_ID)) %>% 
+              unique()) %>% 
+  mutate(Cult_Year = gsub('-.*', '', Cult_Date)) %>% 
+  filter(State != 'TN') %>% 
+  arrange(Patient_ID) %>% 
+  mutate(Pat_ID = factor(Pat_ID, levels = unique(Pat_ID)),
+         Patient_ID = ifelse(is.na(Patient_ID), paste0('P', as.numeric(factor(Pat_ID))), Patient_ID),
+         Patient_ID = ifelse(Patient_ID == 'PNA', NA, Patient_ID)) %>% 
+  select(Sample_ID, Patient_ID, Facility_ID, Cult_Year, State, Species, ST) %>% 
+  arrange(Sample_ID)
+
+all_sample_info_not_tn %>% 
+  write_csv('data/supp_file_1.csv')
+```

### 2022-09-13 manuscript-figures.Rmd
Fix order of fig 4 y axis

@@
          ST = gsub('E. hormaechei', '*E. hormaechei*', ST),
-         ST = gsub('K. pneumoniae', '*K. pneumoniae*', ST)) %>% 
+         ST = gsub('K. pneumoniae', '*K. pneumoniae*', ST), 
+         ST = factor(ST, levels = rev(c('*K. pneumoniae* ST258', '*E. hormaechei* ST171', 'Other', '*E. hormaechei* ST114', '*K. pneumoniae* ST307'))),
+
   ggplot(aes(x = Facility, y = ST, fill = count)) +

### 2023-01-21 manuscript-figures.Rmd
Update manuscript figures

--- 2022-09-13 manuscript-figures.Rmd
@@
 library(lubridate)
+library(scales)
+library(ggrepel)
 
@@
 # facility and ST
-facil_st <- read_csv('data/isolate_data.csv')
-facil_st_mn <- facil_st %>% filter(State == 'MN')
-write_csv(facil_st_mn, 'data/mn_isolate_data.csv')
+facil_st_orig <- read_csv('data/isolate_data.csv') %>% 
+  mutate(cult_year = gsub('-.*|.*/', '', Cult_Date))
+# to make analyses more consistent across states
+facil_st <- facil_st_orig %>% 
+  filter(cult_year >= 2016)
 # public isolate data
@@
 pr_hosp_pairs_ct <- read_csv('data/pr_hosp_pairs_ct.csv')
-pr_hosp_pairs_mn <- read_csv('data/pr_hosp_pairs_mn.csv')
+pr_hosp_pairs_mn_orig <- read_csv('data/pr_hosp_pairs_mn.csv')
+pr_hosp_pairs_mn <- pr_hosp_pairs_mn_orig
 pr_hosp_pairs_tn <- read_csv('data/pr_hosp_pairs_tn.csv')
@@
 facil_type_counts <- read_csv('data/facil_type_counts.csv')
-# facility w prior hospitals
-facil_w <- read_csv('data/facil_w_st171_prior_hosps.csv')
 # prior tn hospitals
@@
 
-### Species counts by state (Table 1)
+### Isolate counts by state
 
@@
   mutate(State = factor(State, levels = c('MN','TN','CT'))) %>% 
-  group_by(State, Species) %>% 
+  group_by(State) %>% 
   count() %>% full_join(state_dates, .) %>% 
@@
 
-### Species counts by facility (Figure S1)
-
-```{r}
-fs1 <- facil_st %>% filter(Species != 'Other') %>% 
-    mutate(ST = factor(ST, levels = rev(names(st_cols))),
-           Species = factor(Species, levels = c('K. pneumoniae','E. hormaechei','Other'))) %>% 
-  ggplot(aes(x = Facility_ID, fill = ST)) + geom_bar(color = 'white') + facet_grid(Species~State, scales = 'free', space = 'free') +
+### Isolate overview (Figure S1A)
+
+```{r}
+fs1a <- facil_st %>% 
+  mutate(ST = ifelse(ST %in% c('E. hormaechei ST114','E. hormaechei ST171', 
+                                 'K. pneumoniae ST258', 'K. pneumoniae ST307'), ST, 
+                       'Other'),
+           ST = factor(ST, levels = rev(c('E. hormaechei ST171', 'E. hormaechei ST114', 
+                                      'K. pneumoniae ST258', 'K. pneumoniae ST307', 
+                                      'Other')))) %>% 
+  group_by(State, ST, cult_year) %>% 
+  filter(!is.na(cult_year)) %>% 
+  summarize(count = n()) %>% 
+  ggplot(aes(x = cult_year, y = count, fill = ST)) +
+  facet_grid(~State, scales = 'free', space = 'free') +
+  geom_col() +
   scale_fill_manual(values = st_cols, 
@@
                                'Other'), 
-                    labels = c('*E. hormaechei* ST171', '*E. hormaechei* ST114',
-                               '*K. pneumoniae* ST258', '*K. pneumoniae* ST307', 
+                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
+                               'KPC-Kp ST258', 'KPC-Kp ST307',
+                               'Other')) +
+  theme(legend.text = element_markdown()) +
+  labs(x = 'Culture year', y = 'Number of isolates')
+```
+
+### Species counts by facility (Figure S1B)
+
+```{r}
+fs1b <- facil_st %>% filter(Species != 'Other') %>% 
+    mutate(ST = factor(ST, levels = rev(names(st_cols))),
+           Species = factor(Species, levels = c('K. pneumoniae','E. hormaechei','Other'),
+                            labels = c('KPC-Kp', 'KPC-Eh', 'Other'))) %>% 
+  ggplot(aes(x = Facility_ID, fill = ST)) + geom_bar(color = 'white') + 
+  facet_grid(Species~State, scales = 'free', space = 'free') +
+  scale_fill_manual(values = st_cols, 
+                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
+                               'K. pneumoniae ST258', 'K. pneumoniae ST307', 
+                               'Other'), 
+                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
+                               'KPC-Kp ST258', 'KPC-Kp ST307', 
                                'Other'),
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        strip.text.y = element_text(face = 'italic'), legend.text = element_markdown())
-
-ggsave(plot = fs1, filename = 'figures/FigS1.png', width = 7, height = 3)
+        legend.text = element_markdown())
+```
+
+```{r, fig.width=5, fig.height=5}
+fs1 <- plot_grid(fs1a, fs1b + theme(legend.position = 'none'), ncol = 1, labels = 'AUTO')
+
+ggsave(plot = fs1, filename = 'figures/FigS1.png', width = 5, height = 5)
 
@@
 
-## Public data (Figure S2)
+## Public data & trees (Figure S2)
 
@@
 
-plot_grid(public_dates_plot, public_loc_plot, ncol = 1, rel_heights = c(0.5, 1), align = 'vh', labels = 'AUTO') %>% 
-  ggsave(filename = 'figures/FigS2.png', width = 7, height = 10)
-```
-
-## Importation vs. HGT
+fs2ab <- plot_grid(public_dates_plot, public_loc_plot, ncol = 1, rel_heights = c(0.5, 1), align = 'vh', axis = 'trbl', labels = 'AUTO') 
+```
+
+```{r}
+states <- metadat_complete %>% select(isolate_no, state) %>% deframe()
+states <- unique(states[c(tree_list[[1]]$tip.label, tree_list[[4]]$tip.label, tree_list[[7]]$tip.label, tree_list[[12]]$tip.label)])
+states <- states[!is.na(states)]
+state_cols <- c('#e6194b', '#3cb44b', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffd8b1', '#000075', '#808080', '#ffffff', '#000000', '#ffe119')
+names(state_cols) <- states
+
+plot_tr_locs <- function(tr, og, title = NULL){
+  tr <- root(tr, og)
+  tr <- drop.tip(tr, og)
+  states <- metadat_complete %>% select(isolate_no, state) %>% deframe()
+  states <- states[tr$tip.label]
+  kpc <- metadat_complete %>% select(isolate_no, kpc) %>% deframe()
+  kpc <- kpc[tr$tip.label]
+  no_info <- is.na(states) & is.na(kpc)
+  states <- states[!no_info]
+  kpc <- kpc[!no_info]
+  kpc <- ifelse(kpc == 1, 'Present', 'Absent')
+  tr <- keep.tip(tr, union(names(states), names(kpc)))
+  states <- c(states, rep(NA, Nnode(tr)))
+  states <- factor(states, levels = sort(names(state_cols[!is.na(names(state_cols))])))
+  kpc <- c(kpc, rep(NA, Nnode(tr)))
+  ggtree(tr) + 
+    geom_tippoint(aes(col = states, shape = kpc)) + 
+    scale_color_manual(values = state_cols, drop=FALSE) + 
+    scale_shape_manual(values = c('Present' = 16, 'Absent' = 1)) +
+    guides(color = guide_legend(ncol = 3)) +
+    labs(title = title, shape = 'KPC', col = 'State')
+}
+
+tr_st114 <- plot_tr_locs(tree_list[[7]], 'MN_CRE226', '*E. hormaechei* ST114') + 
+  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
+tr_st171 <- plot_tr_locs(tree_list[[4]], 'MN_CRE226', '*E. hormaechei* ST171') + 
+  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
+tr_st258 <- plot_tr_locs(tree_list[[1]], '573.1451', '*K. pneumoniae* ST258') + 
+  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
+tr_st307 <- plot_tr_locs(tree_list[[12]], '573.13662', '*K. pneumoniae* ST307') + 
+  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
+leg <- get_legend(plot_tr_locs(tree_list[[1]], '573.1451', '*K. pneumoniae* ST258'))
+
+fs2c <- plot_grid(plot_grid(tr_st171, tr_st114, tr_st258, tr_st307), 
+                  leg, ncol = 1, rel_heights = c(1, 0.6))
+```
+
+```{r, fig.width=10, fig.height=8}
+fs2 <- plot_grid(fs2ab, fs2c, labels = c('', 'C'), label_x = -0.1)
+ggsave('figures/FigS2.png', width = 10, height = 8)
+fs2
+```
+
+## Evidence of importation
 
@@
 genomes <- unique(c(kpc_bin$isolate_no, mn_bin$isolate_no, tn_bin$isolate_no, ct_bin$isolate_no))
+
+keepers <- facil_st %>% 
+  filter(!(duplicated(Patient_ID) & duplicated(ST))) %>% 
+  pull(Sample_ID)
 
@@
 }) %>% 
-  bind_rows() %>% suppressWarnings()
-
-transitions <- transitions %>% 
+  bind_rows() %>% suppressWarnings() %>% 
   mutate(class = gsub('NA| ', '', class), 
          class = ifelse(is.na(class), '', class),
-         hgt_type = ifelse(grepl('gain_kpc', class), 'Unknown', ''),
-         hgt_type = ifelse(class == 'gain_kpc', 'In-state', hgt_type),
-         class2 = ifelse(toupper(gsub('to_','',class)) != state & !grepl('gain_kpc', class), '', class),
-         class2 = ifelse(grepl('gain_kpc', class2), 'Putative\nHGT', class2),
-         class2 = ifelse(class2 == '', 'Unknown', class2),
-         class2 = ifelse(grepl('to_', class2), 'Putative\nimportation', class2),
-         class2 = factor(class2, levels = c('Putative\nimportation','Putative\nHGT','Unknown')),
+         class2 = ifelse(grepl('gain_kpc', class), 'No', class),
+         class2 = ifelse(class2 == '', 'No', class2),
+         class2 = ifelse(grepl('to_', class2), 'Yes', class2),
+         class2 = factor(class2, levels = c('Yes','No')),
          ST = ifelse(ST %in% c('ST258','ST307'), paste0('K. pneumoniae ', ST), ST),
@@
          ST = gsub('cloacae', 'hormaechei', ST),
-         ST_orig = gsub('cloacae', 'hormaechei', ST_orig)
-         )
-
-keepers <- facil_st %>% 
-  filter(!(duplicated(Patient_ID) & duplicated(ST))) %>% 
-  pull(Sample_ID)
-```
-
-### Trees used (Figure S3)
-
-```{r}
-states <- metadat_complete %>% select(isolate_no, state) %>% deframe()
-states <- unique(states[c(tree_list[[1]]$tip.label, tree_list[[4]]$tip.label, tree_list[[7]]$tip.label, tree_list[[12]]$tip.label)])
-states <- states[!is.na(states)]
-state_cols <- c('#e6194b', '#3cb44b', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffd8b1', '#000075', '#808080', '#ffffff', '#000000', '#ffe119')
-names(state_cols) <- states
-
-plot_tr_locs <- function(tr, og, title = NULL){
-  tr <- root(tr, og)
-  tr <- drop.tip(tr, og)
-  states <- metadat_complete %>% select(isolate_no, state) %>% deframe()
-  states <- states[tr$tip.label]
-  kpc <- metadat_complete %>% select(isolate_no, kpc) %>% deframe()
-  kpc <- kpc[tr$tip.label]
-  no_info <- is.na(states) & is.na(kpc)
-  states <- states[!no_info]
-  kpc <- kpc[!no_info]
-  kpc <- ifelse(kpc == 1, 'Present', 'Absent')
-  tr <- keep.tip(tr, union(names(states), names(kpc)))
-  states <- c(states, rep(NA, Nnode(tr)))
-  kpc <- c(kpc, rep(NA, Nnode(tr)))
-  ggtree(tr) + 
-    geom_tippoint(aes(col = states, shape = kpc)) + 
-    scale_color_manual(values = state_cols) + 
-    scale_shape_manual(values = c('Present' = 16, 'Absent' = 1)) +
-    guides(color = guide_legend(ncol = 2)) +
-    labs(title = title, shape = 'KPC', col = 'State')
-}
-
-tr_st114 <- plot_tr_locs(tree_list[[7]], 'MN_CRE226', '*E. hormaechei* ST114') + 
-  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
-tr_st171 <- plot_tr_locs(tree_list[[4]], 'MN_CRE226', '*E. hormaechei* ST171') + 
-  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
-tr_st258 <- plot_tr_locs(tree_list[[1]], '573.1451', '*K. pneumoniae* ST258') + 
-  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
-tr_st307 <- plot_tr_locs(tree_list[[12]], '573.13662', '*K. pneumoniae* ST307') + 
-  theme(legend.position = 'none', plot.title = ggtext::element_markdown())
-leg <- get_legend(plot_tr_locs(tree_list[[7]], 'MN_CRE226', 'E. hormaechei ST114'))
-
-fs3 <- plot_grid(plot_grid(tr_st171, tr_st114, tr_st258, tr_st307), leg)
-
-ggsave(plot = fs3, filename = 'figures/FigS3.png', width = 8, height = 8)
-
-fs3
-```
-
-### Importation vs. HGT (Figure 1)
-
-```{r}
-clust_cols <- st_cols[transitions$ST]
-names(clust_cols) <- paste0(transitions$state, transitions$ST, transitions$cluster)
-
-clust_cols <- (c(clust_cols[clust_cols == "grey"],
-                clust_cols[clust_cols == "coral3"],
-                clust_cols[clust_cols == "coral4"],
-                clust_cols[clust_cols == "aquamarine3"],
-                clust_cols[clust_cols == "aquamarine4"]
-                ))
-
-transitions <- transitions %>% filter(id %in% keepers) %>% 
-  mutate(ST = factor(ST, levels = (c("E. hormaechei ST171", "E. hormaechei ST114", 
+         ST_orig = gsub('cloacae', 'hormaechei', ST_orig),
+         ST = factor(ST, levels = (c("E. hormaechei ST171", "E. hormaechei ST114", 
                                      "K. pneumoniae ST258", "K. pneumoniae ST307", 
@@
          state = factor(state, levels = c('CT', 'MN', 'TN'))) %>% 
+  filter(id %in% keepers) %>% 
   arrange(ST)
-
+```
+
+### Evidence of importation (Figure 1)
+
+```{r}
+transitions <- transitions %>% 
+  mutate(ST = ifelse(grepl('ST258', ST_orig), 'K. pneumoniae ST258', as.character(ST))) 
+clust_cols <- st_cols[as.character(transitions$ST)]
+names(clust_cols) <- paste0(transitions$state, transitions$ST, transitions$cluster)
+
+clust_cols <- (c(clust_cols[clust_cols == "grey"][sort(names(clust_cols[clust_cols == "grey"]), decreasing = TRUE)],
+                clust_cols[clust_cols == "coral3"][sort(names(clust_cols[clust_cols == "coral3"]), decreasing = TRUE)],
+                clust_cols[clust_cols == "coral4"][sort(names(clust_cols[clust_cols == "coral4"]), decreasing = TRUE)],
+                clust_cols[clust_cols == "aquamarine3"][sort(names(clust_cols[clust_cols == "aquamarine3"]), decreasing = TRUE)],
+                clust_cols[clust_cols == "aquamarine4"][sort(names(clust_cols[clust_cols == "aquamarine4"]), decreasing = TRUE)]
+                ))
+```
+
+```{r, fig.width=7, fig.height=4}
 f1 <- transitions %>% 
@@
                     breaks = rev(names(clust_cols)[!duplicated(clust_cols)]),
-                    labels = c('*E. hormaechei* ST171', '*E. hormaechei* ST114',
-                               '*K. pneumoniae* ST258', '*K. pneumoniae* ST307',
+                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
+                               'KPC-Kp ST258', 'KPC-Kp ST307',
                                'Other'),
@@
   facet_grid(~state, scales = 'free', space = 'free') +
-  labs(fill = 'ST cluster', x = 'Cluster type', y = 'Number of isolates') +
+  labs(fill = 'ST cluster', x = 'Evidence that KPC-harboring strain was imported', y = 'Number of isolates') +
   theme(legend.text = ggtext::element_markdown())
@@
 
-### HGT locations (Figure S4)
-
-```{r}
-fs4 <- transitions %>% 
-  filter(id %in% keepers & hgt_type != '') %>%
-  ggplot(aes(x = hgt_type, 
-             fill = factor(paste0(state, ST, cluster), 
-                           levels = unique(paste0(state, ST, cluster))))) + 
-  geom_bar(color = 'white', position = position_stack(reverse = TRUE)) + 
-  scale_fill_manual(values = clust_cols,
-                    breaks = rev(names(clust_cols)[!duplicated(clust_cols)]),
-                    labels = c('*E. hormaechei* ST171', '*E. hormaechei* ST114',
-                               '*K. pneumoniae* ST258', '*K. pneumoniae* ST307',
-                               'Other'),
-                    drop = FALSE) +
-  facet_grid(~state, scales = 'free', space = 'free') +
-  labs(fill = 'ST cluster', x = 'HGT location', y = 'Number of isolates') +
-  theme(legend.text = ggtext::element_markdown())
-
-ggsave(plot = fs4, filename = 'figures/FigS4.png', width = 6, height = 3)
-
-fs4
-```
 ```{r}
@@
          Species = gsub(' ST.*','',st1),
-         ST = gsub('.* ','',ST))
+         ST = gsub('.* ','',ST)) %>% 
+  left_join(facil_st %>% select(Sample_ID, cult_year) %>% 
+              rename(isolate1 = Sample_ID, year1 = cult_year)) %>% 
+  left_join(facil_st %>% select(Sample_ID, cult_year) %>% 
+              rename(isolate2 = Sample_ID, year2 = cult_year))
 ```
@@
 
-### Fraction of intra-facility isolates over time (Figure S5)
+### Fraction of intra-facility isolates over time (Figure S3A)
 
@@
 
-fs5 <- frac_intras %>% 
+fs3a <- frac_intras %>% 
     mutate(ST = ifelse(ST %in% c('E. hormaechei ST114','E. hormaechei ST171', 
@@
                                       'Other'),
-                       labels = c('*E. hormaechei* ST171', '*E. hormaechei* ST114', 
-                                      '*K. pneumoniae* ST258', '*K. pneumoniae* ST307', 
+                       labels = c('KPC-Eh<br>ST171', 'KPC-Eh<br>ST114', 
+                                      'KPC-Kp<br>ST258', 'KPC-Kp<br>ST307', 
                                       'Other'))) %>% 
   filter(frac_intra != 0 & ST != 'Other') %>% 
-  ggplot(aes(x = pairwise_dist, y = frac_intra)) + 
-  geom_bar(stat = "identity", fill = "blue", alpha = 0.5) + 
-  labs(x = "Pairwise SNV distance", y = "Fraction of intra-facility pairs") + xlim(0, 50) + 
-  facet_wrap(state~ST, scales = 'free') +
-  theme(strip.text = element_markdown())
-
-ggsave(plot = fs5, filename = 'figures/FigS5.png', width = 6, height = 5)
-
-fs5
-```
-
-### Pairwise SNV distances by pair type, state, and dominant STs (Figure 2)
-
-```{r}
-f2 <- snv_dat %>% 
-  mutate(pairwise_dist = ifelse(pairwise_dist == 0, 1, pairwise_dist),
+  filter(pairwise_dist <= 50) %>%
+  ggplot(aes(x = pairwise_dist, y = frac_intra, fill = state)) + 
+  geom_bar(stat = "identity") + # fill = "blue", 
+  geom_vline(xintercept = c(5, 10, 15), alpha = 0.2) +
+  labs(x = "Pairwise SNV distance", y = "Fraction of intra-facility pairs", fill = 'State') + 
+  facet_grid(ST~., scales = 'free') +
+  scale_fill_brewer(palette = 'Set1') +
+  scale_x_continuous(breaks = seq(0, 50, by = 5)) +
+  theme(strip.text.y = element_markdown(angle = 0), panel.grid = element_blank())
+```
+
+### Pairwise SNV distances by pair type, state, and other STs (Figure S3B)
+
+```{r}
+fs3b <- snv_dat %>% 
+  filter(state1 == state2 & pair_type != 'Unknown' & !is.na(pair_type)) %>%
+  mutate(pairwise_dist = ifelse(pairwise_dist == 0, 0.1, pairwise_dist),
          pair_type = factor(gsub(' pair', '', pair_type), levels = c('Intra-facility','Inter-facility')),
-         Species = gsub('^', '\\*', Species),
-         Species = gsub('$', '\\*', Species),
-         state1 = factor(state1, levels = c('CT', 'MN', 'TN'))) %>% 
-  filter(pair_type != 'Unknown' & ST != 'Other') %>% 
-  ggplot(aes(x = pair_type, y = pairwise_dist)) + 
-  geom_jitter(aes(fill = under_thresh, color = under_thresh),alpha = 0.5) +
+         state1 = factor(state1, levels = c('CT','MN','TN')),
+         # pair_type = ifelse(is.na(pair_type), 'Unknown', pair_type),
+         # pair_type = gsub(' pair', '', pair_type),
+         st1 = gsub('K. pneumoniae', 'KPC-Kp', st1),
+         st1 = gsub('E. hormaechei', 'KPC-Eh', st1)) %>% 
+  ggplot(aes(x = pairwise_dist, y = st1, color = under_thresh)) + 
+  # geom_vline(xintercept = c(5, 10, 15), color = 'darkgrey', linetype = 'dashed') +
+  geom_jitter(alpha=0.5) + 
+  facet_grid(pair_type~state1, scales = 'free', space = 'free') +
   scale_color_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
-  geom_violin(position = 'identity',alpha=0, color = 'azure4') +
-  facet_grid(Species+ST~state1, scales = 'free') + scale_y_log10() +
-  labs(x = 'Pair type',y = 'Pairwise SNV distance', color = 'SNV bin', fill = 'SNV bin') +
-  theme(strip.text.y = element_markdown())
-
-ggsave(plot = f2, filename = 'figures/Fig2.png', width = 6.5, height = 5)
-
-f2
-```
-
-### Pairwise SNV distances by pair type, state, and all STs (Figure S6)
-
-```{r}
-fs6 <- snv_dat %>% filter(state1 == state2 & pair_type != 'Unknown') %>% 
-  mutate(state1 = factor(state1, levels = c('CT','MN','TN')),
-         pair_type = ifelse(is.na(pair_type), 'Unknown', pair_type),
-         st1 = gsub('^', '\\*', st1),
-         st1 = gsub('e ', 'e\\* ', st1),
-         st1 = gsub('i ', 'i\\* ', st1)) %>% 
-  ggplot(aes(x = pairwise_dist, y = st1, color = pair_type)) + 
-  geom_vline(xintercept = 10, color = 'darkgrey') +
-  geom_jitter(alpha=0.25) + facet_grid(~state1, scales = 'free') +
-  scale_color_manual(values = c('lightslateblue','black')) +
-  labs(y = '', x = 'Pairwise SNV distance', color = '') +
-  theme(axis.text.y = element_markdown())
-
-ggsave(plot = fs6, filename = 'figures/FigS6.png', width = 7, height = 5)
-
-fs6
-```
-
-### Patristic distance sensitivity analysis (Figure S7)
-
-```{r}
-fs7 <- left_join(snv_dat %>% rename(pd_snv = pairwise_dist), pat_dat %>% rename(pd_pat = pairwise_dist)) %>% 
-  mutate(state1 = factor(state1, levels = c('CT', 'MN', 'TN')),
-         pair_type = factor(gsub(' pair', '', pair_type), levels = c('Intra-facility','Inter-facility')),
-         Species = gsub('^', '\\*', Species),
-         Species = gsub('$', '\\*', Species)) %>% 
-  filter(pair_type != 'Unknown' & ST != 'Other') %>% 
-  ggplot(aes(x = pair_type, y = pd_pat)) + geom_jitter(aes(fill = under_thresh, color = under_thresh),alpha = 0.5) +
+  scale_x_continuous(breaks = seq(0, 600, by = 50)) +
+  # scale_color_manual(values = c('lightslateblue','black')) +
+  labs(y = '', x = 'Pairwise SNV distance', color = 'SNV bin') 
+```
+
+### Patristic distance sensitivity analysis (Figure S3C)
+
+```{r}
+fs3c <- left_join(snv_dat %>% rename(pd_snv = pairwise_dist), pat_dat %>% rename(pd_pat = pairwise_dist)) %>% 
+  filter(state1 == state2 & pair_type != 'Unknown' & !is.na(pair_type)) %>%
+  mutate(pair_type = factor(gsub(' pair', '', pair_type), levels = c('Intra-facility','Inter-facility')),
+         state1 = factor(state1, levels = c('CT','MN','TN')),
+         st1 = gsub('K. pneumoniae', 'KPC-Kp', st1),
+         st1 = gsub('E. hormaechei', 'KPC-Eh', st1)) %>% 
+  ggplot(aes(x = pd_pat, y = st1, color = under_thresh)) + 
+  geom_jitter(alpha=0.5) + 
+  facet_grid(pair_type~state1, scales = 'free', space = 'free') +
   scale_color_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
-  geom_violin(position = 'identity',alpha=0, color = 'azure4') +
-  facet_grid(Species+ST~state1, scales = 'free') + scale_y_log10() +
-  labs(x = 'Pair type',y = 'Pairwise patristic distance', color = 'SNV bin', fill = 'SNV bin') +
-  theme(strip.text.y = element_markdown())
-
-ggsave(plot = fs7, filename = 'figures/FigS7.png', width = 6.5, height = 5)
-
-fs7
-```
-
-## Closely related facility type pairs (Table S2)
-
-```{r}
-pr_hosp_pairs_ct <- left_join(pr_hosp_pairs_ct, snv_dat)
-pr_hosp_pairs_mn <- left_join(pr_hosp_pairs_mn, snv_dat)
-pr_hosp_pairs_tn <- left_join(pr_hosp_pairs_tn, snv_dat)
-
-nodes <- bind_rows(pr_hosp_pairs_mn %>% select(loc1, state1, st1) %>% 
-                     rename(Facility_ID = loc1, state = state1, st = st1),
-                   pr_hosp_pairs_mn %>% select(loc2, state2, st2) %>% 
-                     rename(Facility_ID = loc2, state = state2, st = st2),
-                   pr_hosp_pairs_tn %>% select(loc1, state1, st1) %>% 
-                     rename(Facility_ID = loc1, state = state1, st = st1),
-                   pr_hosp_pairs_tn %>% select(loc2, state2, st2) %>% 
-                     rename(Facility_ID = loc2, state = state2, st = st2),
-                   pr_hosp_pairs_ct %>% select(loc1, state1, st1) %>% 
-                     rename(Facility_ID = loc1, state = state1, st = st1),
-                   pr_hosp_pairs_ct %>% select(loc2, state2, st2) %>% 
-                     rename(Facility_ID = loc2, state = state2, st = st2)) %>% 
-  filter(!duplicated(.)) %>% filter(!is.na(Facility_ID)) %>%
-  select(-st) 
-  # mutate(st = gsub('cloacae', 'hormaechei', st),
-  #        st = ifelse(st %in% c('E. hormaechei ST114','E. hormaechei ST171', 'K. pneumoniae ST258', 'K. pneumoniae ST307'), st, 'Other'))
-
-edges <- bind_rows(pr_hosp_pairs_mn %>% rename(from = loc1, to = loc2), 
-          pr_hosp_pairs_tn %>% rename(from = loc1, to = loc2), 
-          pr_hosp_pairs_ct %>% rename(from = loc1, to = loc2)) %>% 
-  filter(pairwise_dist < 10) %>% 
-  select(from, to, same_pr_hosp, st1) %>%
-  filter(!is.na(to) & !is.na(from)) %>% 
-  mutate(st1 = gsub('cloacae', 'hormaechei', st1),
-         st1 = ifelse(st1 %in% c('E. hormaechei ST114','E. hormaechei ST171', 'K. pneumoniae ST258', 'K. pneumoniae ST307'), st1, 'Other'),
-         st1 = factor(st1, levels = c('E. hormaechei ST114','E. hormaechei ST171', 
-                                      'K. pneumoniae ST258', 'K. pneumoniae ST307', 'Other'),
-                      labels = c('*E. hormaechei* ST114','*E. hormaechei* ST171', 
-                                      '*K. pneumoniae* ST258', '*K. pneumoniae* ST307', 'Other')),
-         same_pr_hosp = ifelse(same_pr_hosp, 'Yes','No'))
-
-graph <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE) %>% 
-  induced.subgraph(igraph::degree(.) > 0)
-```
-
-```{r}
-left_join(edges, facil_types %>% rename(from = Facility_ID, type_from = type)) %>% 
-  left_join(facil_types %>% rename(to = Facility_ID, type_to = type)) %>% 
-  mutate(type_pair = sapply(1:nrow(.), function(x) paste0(sort(c(type_from[x], type_to[x])), collapse = '-')),
-         type_pair = ifelse(type_pair %in% c('ACH-ACH', 'ACH-LTACH', 'ACH-SNF'), type_pair, 'Other')) %>% 
-  group_by(state, st1, type_pair) %>% 
-  summarize(count = n()) %>% 
-  arrange(state, st1, type_pair) %>% 
-  pivot_wider(names_from = type_pair, values_from = count, values_fill = 0) %>% 
-  select(state, st1, `ACH-ACH`, `ACH-LTACH`, `ACH-SNF`, Other)
-```
-
-## Intra-facility analysis (Figure S8)
+  scale_x_continuous(breaks = seq(0, 0.01, by = 0.00003)) +
+  labs(y = '', x = 'Pairwise patristic distance', color = 'SNV bin') 
+```
+
+```{r, fig.width=7.5, fig.height=10.5}
+fs3 <- plot_grid(plotlist = list(fs3a, fs3b, fs3c), ncol = 1, labels = 'AUTO',
+                 rel_heights = c(0.75, 1.1, 1.1),
+                 align = 'hv', axis = 'trbl')
+
+ggsave(plot = fs3, filename = 'figures/FigS3.png', width = 7.5, height = 10.5)
+
+fs3
+```
+
+## Intra-facility analysis (Figure 2A)
 
@@
 min_dat <- lapply(unique(c(as.character(snv_dat$isolate1), as.character(snv_dat$isolate2))), function(i){
-  snv_dat %>% filter(pair_type == 'Intra-facility pair' & (isolate1 == i | isolate2 == i)) %>% 
-  group_by(loc1, ST_orig, state1) %>% 
-  summarize(dist_min = min(pairwise_dist))
+  snv_dat %>% filter((isolate1 == i | isolate2 == i) & pair_type != 'Unknown') %>%
+    group_by(pair_type, ST_orig, state1) %>% 
+    summarize(dist_min = min(pairwise_dist)) %>% 
+    pivot_wider(names_from = pair_type, values_from = dist_min)
 }) %>% bind_rows() %>% suppressMessages() %>% suppressWarnings()
@@
 ```{r}
-fs8 <- min_dat %>% 
-    mutate(under_thresh = ifelse(dist_min <= 15, '(10,15]','(15,∞)'),
-         under_thresh = ifelse(dist_min <= 10, '(5,10]',under_thresh),
-         under_thresh = ifelse(dist_min <= 5, '[0,5]',under_thresh),
-         under_thresh = factor(under_thresh, levels = rev(c('[0,5]','(5,10]','(10,15]','(15,∞)'))),
-         ST = ifelse(ST_orig %in% c('K. pneumoniae ST258','E. hormaechei ST171', 'K. pneumoniae ST307','E. hormaechei ST114'), ST_orig, 'Other'),
-         ST = factor(ST, levels = c('E. hormaechei ST171', 'E. hormaechei ST114', 'K. pneumoniae ST258', 'K. pneumoniae ST307', 'Other'),
-                     labels = c('*E. hormaechei*<br>ST171', '*E. hormaechei*<br>ST114', 
-                                '*K. pneumoniae*<br>ST258', '*K. pneumoniae*<br>ST307', 'Other')),
-         state1 = factor(state1, levels = c('CT', 'MN', 'TN'))) %>% 
-  ggplot(aes(x = reorder(loc1, loc1, function(x)-length(x)), fill = under_thresh)) + 
-  geom_bar() +
-       scale_fill_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
-         labs(y = "Number of isolates", x = "Facility", fill = 'SNV bin of closest\nintra-facility isolate') +
-  facet_grid(ST~state1, space = 'free', scales = 'free_x') +
-  theme(axis.text.x = element_blank(), strip.text.y = element_markdown()) + 
-  scale_y_continuous(breaks = scales::pretty_breaks())
-
-ggsave(plot = fs8, filename = 'figures/FigS8.png', width = 7, height = 5)
-
-fs8
-```
-
-## Clinical factor analysis (Table S3)
-
-```{r}
-fisher <- function(a,b,c,d){
-  data <- matrix(c(a,b,c,d),ncol=2)
-  c(p = fisher.test(data)$p.value,
-    OR = fisher.test(data)$estimate)
-}
-```
+min_dat %>% 
+  filter(!is.na(`Inter-facility pair`) & !is.na(`Intra-facility pair`)) %>% 
+  group_by(ST_orig, state1) %>% 
+  summarize(n = n(), 
+            median_intra = median(`Intra-facility pair`),
+            median_inter = median(`Inter-facility pair`),
+            wilcox_p = wilcox.test(`Intra-facility pair`, `Inter-facility pair`, 
+                                   paired = TRUE)$p.value) %>% 
+  arrange(-n)
+```
+
+```{r}
+f2a <- min_dat %>% 
+  filter(!is.na(`Inter-facility pair`) & !is.na(`Intra-facility pair`)) %>% 
+  group_by(ST_orig, state1) %>% 
+  filter(!`Intra-facility pair` > 400) %>% 
+  mutate(ST = ifelse(ST_orig %in% c('E. hormaechei ST171', 'E. hormaechei ST114', 
+                                        'K. pneumoniae ST258', 'K. pneumoniae ST307'),
+                     ST_orig, 'Other')) %>% 
+  ggplot(aes(x = `Intra-facility pair`, y = `Inter-facility pair`)) +
+  facet_grid(~state1, scales = 'free') +
+  geom_abline(intercept = 0, linetype = 'dotted') +
+  geom_point(aes( color = ST), alpha = 0.5) +
+  geom_text(aes(label = 'y=x', x = 65, y = 80), check_overlap = TRUE) +
+  scale_color_manual(values = st_cols, 
+                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
+                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
+                               'Other'), 
+                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
+                               'KPC-Kp ST258', 'KPC-Kp ST307',
+                               'Other')) +
+  scale_x_continuous(breaks = seq(0, 160, by = 20)) +
+  theme(panel.grid = element_blank()) +
+  labs(x = 'Minimum pairwise SNV distance of\nan isolate from the same facility', y = 'Minimum pairwise SNV distance of\nan isolate from a different facility')
+```
+
+## Clinical factor analysis
 
@@
 })) %>% bind_rows() 
-
-pt_info_comb <- bind_rows(pt_info_mn, pt_info_tn, pt_info_ct) %>%
-  mutate(`pairwise_dist <= 10` = pairwise_dist <= 10) %>% 
-  select(state1, st1, shared_clin, `pairwise_dist <= 10`)
-
-pt_comorb_compare <- pt_info_comb %>% 
-  group_by(state1, st1, shared_clin, `pairwise_dist <= 10`) %>% summarize(count = n()) %>% 
-  mutate(sharedclin_pdlt10 = factor(paste0(ifelse(shared_clin, 'shared', 'not shared'), ', ', 
-                                           ifelse(`pairwise_dist <= 10`, '≤10', '>10')),
-                                    levels = c('shared, ≤10', 'shared, >10', 'not shared, ≤10', 'not shared, >10'))) %>% 
+```
+
+```{r}
+shared_clin_pvals <- bind_rows(pt_info_mn, pt_info_tn, pt_info_ct) %>%
+  left_join(snv_dat %>% filter(loc1 != loc2)) %>% 
+  group_by(state1, st1) %>% 
+  filter(sum(shared_clin) != 0 & sum(!shared_clin) != 0) %>%
+  summarize(wilcox_p = paste0('p = ', signif(wilcox.test(pairwise_dist[shared_clin], pairwise_dist[!shared_clin], alternative = 'less')$p.value, 5)))
+shared_clin_pvals
+```
+
+## Inter-facility analysis
+
+### Prior hospital
+
+#### Individual healthcare exposures
+
+```{r}
+pr_hosp_pairs_ct <- inner_join(pr_hosp_pairs_ct, snv_dat)
+pr_hosp_pairs_mn <- inner_join(pr_hosp_pairs_mn_orig, snv_dat) 
+pr_hosp_pairs_tn <- inner_join(pr_hosp_pairs_tn, snv_dat)
+# pr_hosp_pairs_mn_all <- inner_join(pr_hosp_pairs_mn_orig, mn_snv_dat_orig)
+
+pr_hops_pvals <- bind_rows(pr_hosp_pairs_ct, pr_hosp_pairs_mn, pr_hosp_pairs_tn,
+                           # pr_hosp_pairs_mn_all %>% mutate(state1 = 'MN_all')
+                           ) %>%
+  left_join(snv_dat %>% filter(loc1 != loc2)) %>% 
+  group_by(state1, st1) %>% 
+  filter(sum(same_pr_hosp) != 0 & sum(!same_pr_hosp) != 0) %>% 
+  summarize(n_shared = sum(same_pr_hosp),
+            n_not_shared = sum(!same_pr_hosp),
+            median_shared = median(pairwise_dist[same_pr_hosp]),
+            median_not_shared = median(pairwise_dist[!same_pr_hosp]),
+            wilcox_p = paste0('p = ', signif(wilcox.test(pairwise_dist[same_pr_hosp],
+                                                         pairwise_dist[!same_pr_hosp], 
+                                                         alternative = 'less')$p.value, 1)))
+```
+
+```{r}
+f2b <- bind_rows(pr_hosp_pairs_ct, pr_hosp_pairs_mn %>% filter(!is.na(state1)), pr_hosp_pairs_tn) %>%
+  left_join(snv_dat %>% filter(loc1 != loc2)) %>% 
+  mutate(st1 = ifelse(st1 %in% c("E. hormaechei ST114", "E. hormaechei ST171", "K. pneumoniae ST258", "K. pneumoniae ST307"), st1, 'Other')) %>% 
+  left_join(pr_hops_pvals) %>% 
+  filter(state1 == 'MN' & st1 == 'E. hormaechei ST171' |
+           state1 == 'TN' & st1 %in% c('K. pneumoniae ST258', 'K. pneumoniae ST307', 'E. hormaechei ST114')) %>% 
+  mutate(same_pr_hosp = factor(ifelse(same_pr_hosp, 'Yes', 'No'), levels = c('Yes', 'No')),
+         # st1 = gsub('E', '\\*E', st1),
+         # st1 = gsub('K', '\\*K', st1),
+         # st1 = gsub('i ', 'i\\* ', st1),
+         # st1 = gsub('e ', 'e\\* ', st1),
+         wilcox_p = ifelse(wilcox_p == 'p = 9e-07', 'p < 0.001', wilcox_p),
+         st1 = gsub('K. pneumoniae', 'KPC-Kp', st1),
+         st1 = gsub('E. hormaechei', 'KPC-Eh', st1),
+         st1 = paste0(st1, ': ', wilcox_p)) %>% 
+  ggplot() +
+  facet_grid(~state1+st1, scales = 'free') +
+  # geom_text(aes(x = 0, y = 50, label = wilcox_p), check_overlap = TRUE, vjust = 0, hjust = 0) +
+  # geom_histogram(aes(x = pairwise_dist), fill = 'white') +
+  # geom_histogram(aes(x = pairwise_dist, fill = same_pr_hosp), alpha = 0.5) +
+  geom_density(aes(x = pairwise_dist, fill = same_pr_hosp), alpha = 0.5) +
+  scale_fill_grey(start = 0.3, end = 0.7) +
+  labs(x = 'Pairwise SNV distance', y = 'Density of pairs', fill = 'Shared facility\nexposure') +
+  theme(strip.text.x = element_markdown())
+```
+
+```{r}
+bind_rows(pr_hosp_pairs_ct, pr_hosp_pairs_mn %>% filter(!is.na(state1)), pr_hosp_pairs_tn) %>%
+  left_join(snv_dat %>% filter(loc1 != loc2)) %>% 
+  mutate(st1 = ifelse(st1 %in% c("E. hormaechei ST114", "E. hormaechei ST171", "K. pneumoniae ST258", "K. pneumoniae ST307"), st1, 'Other')) %>% 
+  left_join(pr_hops_pvals) %>% 
+  filter(state1 == 'MN' & st1 == 'E. hormaechei ST171' |
+           state1 == 'TN' & st1 %in% c('K. pneumoniae ST258', 'K. pneumoniae ST307', 'E. hormaechei ST114')) %>% 
+  select(state1, ST, n_shared, n_not_shared) %>% 
+  unique()
+```
+
+```{r, fig.width=7, fig.height=7}
+f2 <- plot_grid(f2a + theme(legend.position = 'top'), 
+                          f2b + theme(legend.position = 'bottom'), 
+                          ncol = 1, labels = 'AUTO', align = 'h', axis = 'tr')
+
+ggsave(plot = f2, filename = 'figures/Fig2.png', width = 7, height = 7)
+
+f2
+```
+
+### MN over time (Figure S4)
+
+```{r}
+fs4a <- facil_st_orig %>% 
+  filter(State == 'MN' & ST %in% c('K. pneumoniae ST258', 'E. hormaechei ST171')) %>% 
+  mutate(facil = ifelse(Facility_ID == 'F38', 'F38', 'Other'),
+         facil = ifelse(is.na(facil), 'Other', facil),
+         ST = gsub('K. pneumoniae', 'KPC-Kp', ST),
+         ST = gsub('E. hormaechei', 'KPC-Eh', ST)) %>% 
+  ggplot(aes(x = cult_year, fill = facil)) +
+  facet_wrap(~ST) +
+  geom_bar() +
+  scale_fill_manual(values = rev(c('grey80', 'mistyrose3'))) + #'mistyrose1', 'grey80')) +
+  labs(x = 'Year', y = 'Number of isolates', fill = 'Facility')
+```
+
+```{r}
+# snv distances
+mn_kp_snv_dists <- get_pair_types(kp_dists, 
+                                  facil_st_orig %>% 
+                                    filter(State == 'MN') %>% 
+                                    select(Sample_ID, Facility_ID) %>% deframe(), 
+                                  pt = NULL) %>% 
+  filter(!sapply(1:nrow(.), function(x) sort(c(as.character(isolate1[x]),as.character(isolate2[x])))) %>% t() %>% duplicated()) %>% 
+  mutate(st1 = sapply(isolate1, function(x) facil_st_orig$ST[facil_st_orig$Sample_ID == x]),
+         st2 = sapply(isolate2, function(x) facil_st_orig$ST[facil_st_orig$Sample_ID == x]),
+         state1 = sapply(isolate1, function(x) facil_st_orig$State[facil_st_orig$Sample_ID == x]),
+         state2 = sapply(isolate2, function(x) facil_st_orig$State[facil_st_orig$Sample_ID == x])) %>% 
+  filter(st1 == st2 & !grepl('novel',st1)) %>% 
+  mutate(state_pair = paste(state1, state2))
+
+mn_ec_snv_dists <- get_pair_types(ec_dists, facil_st_orig %>% 
+                                    filter(State == 'MN') %>% 
+                                    select(Sample_ID, Facility_ID) %>% 
+                                    deframe(), 
+                                  pt = NULL) %>% 
+  filter(!sapply(1:nrow(.), function(x) sort(c(as.character(isolate1[x]),as.character(isolate2[x])))) %>% t() %>% duplicated()) %>% 
+  mutate(st1 = sapply(isolate1, function(x) facil_st_orig$ST[facil_st_orig$Sample_ID == x]),
+         st2 = sapply(isolate2, function(x) facil_st_orig$ST[facil_st_orig$Sample_ID == x]),
+         state1 = sapply(isolate1, function(x) facil_st_orig$State[facil_st_orig$Sample_ID == x]),
+         state2 = sapply(isolate2, function(x) facil_st_orig$State[facil_st_orig$Sample_ID == x])) %>% 
+  filter(st1 == st2 & !grepl('novel',st1)) %>% 
+  mutate(state_pair = paste(state1, state2))
+
+mn_snv_dat_orig <- bind_rows(mn_kp_snv_dists, mn_ec_snv_dists) %>% filter(state1 == state2) %>% 
+  mutate(state1 = factor(state1, levels = c('MN','TN','CT')),
+         pair_type = ifelse(is.na(pair_type), 'Unknown', pair_type),
+         under_thresh = ifelse(pairwise_dist <= 15, '(10,15]','(15,∞)'),
+         under_thresh = ifelse(pairwise_dist <= 10, '(5,10]',under_thresh),
+         under_thresh = ifelse(pairwise_dist <= 5, '[0,5]',under_thresh),
+         under_thresh = factor(under_thresh, levels = c('[0,5]','(5,10]','(10,15]','(15,∞)')),
+         ST_orig = st1,
+         ST = ifelse(st1 %in% dominant_sts, st1, 'Other'),
+         Species = gsub(' ST.*','',st1),
+         ST = gsub('.* ','',ST)) %>% 
+  left_join(facil_st_orig %>% select(Sample_ID, Cult_Date, cult_year) %>% 
+              rename(isolate1 = Sample_ID, date1 = Cult_Date, year1 = cult_year)) %>% 
+  left_join(facil_st_orig %>% select(Sample_ID, Cult_Date, cult_year) %>% 
+              rename(isolate2 = Sample_ID, date2 = Cult_Date, year2 = cult_year)) %>% 
+  mutate(datediff = abs(ymd(date1) - ymd(date2)))
+
+mn_snv_dat <- mn_snv_dat_orig %>% 
+  filter(ST %in% c('ST258', 'ST171') & year1 == year2)
+```
+
+```{r}
+fs4b <- mn_snv_dat %>% 
+  filter(pair_type != 'Unknown') %>% 
+  mutate(pair_type = gsub(' pair', '', pair_type),
+         ST = ifelse(ST == 'ST171', 'KPC-Eh ST171', 'KPC-Kp ST258')) %>% 
+  ggplot(aes(x = year1, y = pairwise_dist, col = under_thresh, shape = pair_type)) +
+  facet_grid(~ST) +
+  geom_jitter(alpha = 0.7) +
+  scale_color_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
+  scale_shape_manual(values = c(1, 19)) +
+  labs(x = 'Year', y = 'Pairwise SNV distance', col = 'SNV bin', shape = 'Isolate\npair type')
+```
+
+```{r, fig.width=7, fig.height=7}
+fs4 <- plot_grid(fs4a, fs4b, ncol = 1, labels = 'AUTO')
+ggsave(plot = fs4, filename = 'figures/FigS4.png', width = 7, height = 7)
+fs4
+```
+
+## MN ST171 transmission (Figure 3)
+
+```{r}
+f3a <- mn_snv_dat_orig %>% 
+  filter(ST %in% c('ST258', 'ST171') & pair_type != 'Unknown') %>% 
+  mutate(datediff_bin = cut(as.numeric(datediff), breaks = c(-1, 365*1:7),
+                            labels = c('0-1', paste0(1:6, '-', 2:7))),
+         under_thresh = factor(under_thresh, 
+                               levels = rev(unique(levels(under_thresh)))),
+         st1 = gsub('E. hormaechei', 'KPC-Eh', st1),
+         st1 = gsub('K. pneumoniae', 'KPC-Kp', st1)) %>% 
+  ggplot(aes(x = datediff_bin, fill = under_thresh)) +
+  facet_grid(pair_type~st1, scales = 'free') + 
+  geom_bar() +
+  scale_fill_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
+  labs(x = 'Number of years separating isolates', y = 'Number of isolate pairs',
+       fill = 'SNV bin')
+```
+
+```{r}
+f3b <- mn_snv_dat_orig %>% 
+  filter(ST %in% c('ST171') & loc1 == loc2 & pairwise_dist <= 15) %>% 
+  mutate(datediff_bin = cut(as.numeric(datediff), breaks = c(-1, 365*1:7),
+                            labels = c('0-1', paste0(1:6, '-', 2:7))),
+         under_thresh = factor(under_thresh,
+                               levels = rev(unique(levels(under_thresh))))
+         ) %>% 
+  group_by(loc1, ST, under_thresh, datediff_bin) %>% 
+  summarize(count = n()) %>%
+  mutate(f38 = ifelse(count > 50, loc1, NA),
+         ST = paste0('KPC-Eh ', ST)) %>%
+  ggplot(aes(x = loc1, y = count, fill = under_thresh)) +
+  facet_grid(~ST) +
+  geom_col() +
+  scale_y_continuous(breaks = pretty_breaks())  +
+  scale_fill_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
+  labs(y = 'Number of pairs', x = 'Facility of intra-facility pair', fill = 'SNV bin')
+```
+
+```{r}
+f3c <- mn_snv_dat_orig %>% 
+  filter(ST %in% c('ST171') & pairwise_dist <= 15 & loc1 != loc2) %>% 
+  mutate(datediff_bin = cut(as.numeric(datediff), breaks = c(-1, 365*1:7),
+                            labels = c('0-1', paste0(1:6, '-', 2:7))),
+         under_thresh = factor(under_thresh, 
+                               levels = rev(unique(levels(under_thresh))))) %>% 
+  pivot_longer(c(loc1, loc2)) %>% 
+  group_by(value, ST, under_thresh) %>% 
+  summarize(count = n()) %>%
+  mutate(f38 = ifelse(count > 50, value, NA),
+         ST = paste0('KPC-Eh ', ST)) %>% 
+  ggplot(aes(x = value, y = count, fill = under_thresh)) +
+  facet_grid(~ST, scales = 'free') +
+  # geom_histogram(binwidth = 5) +
+  geom_col() +
+  # geom_text(aes(y = 1.5, label = f38)) +
+  scale_y_continuous(breaks = pretty_breaks())  +
+  scale_fill_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
+  labs(y = 'Number of isolates', x = 'Facility of inter-facility pair', fill = 'SNV bin')
+```
+
+```{r}
+mn_snv_dat_orig %>% 
+  filter(ST %in% c('ST171') & pairwise_dist <= 10 &
+           !is.na(loc1) & !is.na(loc2) & loc1 != loc2) %>% 
+  mutate(datediff_bin = cut(as.numeric(datediff), breaks = c(-1, 365*1:7),
+                            labels = c('0-1 year', paste0(1:6, '-', 2:7, ' years'))),
+         under_thresh = factor(under_thresh, 
+                               levels = rev(unique(levels(under_thresh)))),
+         loc1 = ifelse(loc1 %in% c('F38'), loc1, 'Other'),
+         loc2 = ifelse(loc2 %in% c('F38'), loc2, 'Other')
+         # loc1 = ifelse(loc1 %in% c('F38', 'F55', 'F34', 'F46'), loc1, 'Other'),
+         # loc2 = ifelse(loc2 %in% c('F38', 'F55', 'F34', 'F46'), loc2, 'Other')
+         ) %>% 
+  rowwise() %>% 
+  mutate(loc1_new = paste(sort(c(loc1, loc2)))[1],
+         loc2_new = paste(sort(c(loc1, loc2)))[2]) %>% 
   ungroup() %>%
-  arrange(sharedclin_pdlt10) %>% 
-  select(-c(shared_clin, `pairwise_dist <= 10`)) %>% 
-  pivot_wider(names_from = sharedclin_pdlt10, values_from = count, values_fill = 0) %>% 
-  group_by(state1, st1) %>% 
+  mutate(f38 = loc1_new == 'F38' | loc2_new == 'F38') %>% 
+  group_by(f38) %>% 
+  summarize(count = n()) %>% 
+  ungroup() %>% 
+  mutate(tot = sum(count),
+         pct = paste0(round(count/tot*100), '%')) 
+```
+
+```{r}
+mn_snv_dat_orig %>% 
+  filter(ST %in% c('ST171') & pairwise_dist <= 10 &
+           !is.na(loc1) & !is.na(loc2) & loc1 == loc2) %>% 
+  mutate(f38 = loc1 == 'F38' | loc2 == 'F38') %>% 
+  group_by(f38) %>% 
+  summarize(count = n()) %>% 
+  ungroup() %>% 
+  mutate(tot = sum(count),
+         pct = paste0(round(count/tot*100), '%')) 
+```
+
+```{r}
+f3d <- mn_snv_dat_orig %>% 
+  filter(ST %in% c('ST171') & pairwise_dist <= 10 &
+           !is.na(loc1) & !is.na(loc2) & loc1 != loc2) %>% 
+  mutate(datediff_bin = cut(as.numeric(datediff), breaks = c(-1, 365*1:7),
+                            labels = c('0-1 year', paste0(1:6, '-', 2:7, ' years'))),
+         under_thresh = factor(under_thresh, 
+                               levels = rev(unique(levels(under_thresh)))),
+         loc1 = ifelse(loc1 %in% c('F38', 'F55', 'F34', 'F46'), loc1, 'Other'),
+         loc2 = ifelse(loc2 %in% c('F38', 'F55', 'F34', 'F46'), loc2, 'Other')
+         ) %>% 
   rowwise() %>% 
-  mutate(fe_p = fisher(`shared, ≤10`, `shared, >10`, `not shared, ≤10`, `not shared, >10`)[[1]]) %>% 
-  arrange(state1, st1) %>% 
-  mutate(frac_shared_lt10 = `shared, ≤10`/(`shared, ≤10` + `not shared, ≤10`),
-         frac_shared_gt10 = `shared, >10`/(`shared, >10` + `not shared, >10`))
-
-pt_comorb_compare
-```
-
-## Inter-facility analysis
-
-### Aggregate patient transfer (Figure S9)
-
-```{r}
-mn_pt_trans <- get_patient_flow(mn_trans_net_sub, 
-                                unique(facil_st$Facility_ID)[unique(facil_st$Facility_ID) %in% 
-                                                        c(mn_trans_net_sub$source_facil,
-                                                          mn_trans_net_sub$dest_facil)]) %>% 
-  filter(!duplicated(.))
-tn_pt_trans <- get_patient_flow(tn_trans_net_sub, unique(facil_st$Facility_ID)[unique(facil_st$Facility_ID) %in% 
-                                                        c(tn_trans_net_sub$source_facil,
-                                                          tn_trans_net_sub$dest_facil)]) %>% 
-  filter(!duplicated(.))
-ct_pt_trans <- get_patient_flow(ct_trans_net_sub, unique(facil_st$Facility_ID)[unique(facil_st$Facility_ID) %in% 
-                                                        c(ct_trans_net_sub$source_facil,
-                                                          ct_trans_net_sub$dest_facil)]) %>% 
-  filter(!duplicated(.))
-
-sts <- c('ST171','ST258','ST114','ST307')
-
-snv_by_st <- suppressWarnings(lapply(sts, function(x){
-  snvdat_sub <- snv_dat %>% filter(ST == x) %>% mutate(pairwise_dist = as.numeric(pairwise_dist))
-  snvdat_sub <- snvdat_sub[,1:6] 
-    mn_snv_by_st_sub <- summarize_pairs(snvdat_sub) %>% mutate(ST = x)
-})) %>% bind_rows()
-
-pt_snv <- bind_rows(bind_cols(state = 'MN', mn_pt_trans), 
-                    bind_cols(state = 'TN', tn_pt_trans),
-                    bind_cols(state = 'CT', ct_pt_trans)) %>% inner_join(snv_by_st) %>%
-  mutate(under_thresh = ifelse(dist_min <= 15, '(10,15]','(15,∞)'),
-         under_thresh = ifelse(dist_min <= 10, '(5,10]',under_thresh),
-         under_thresh = ifelse(dist_min <= 5, '[0,5]',under_thresh),
-         under_thresh = factor(under_thresh, levels = c('[0,5]','(5,10]','(10,15]','(15,∞)')),
-         species = ifelse(ST %in% c('ST258','ST307'), 'K. pneumoniae','E. hormaechei'))
-```
-
-```{r}
-pvals <- pt_snv %>% 
-  mutate(leq_10_bool = ifelse(dist_min <= 10, 'close','not_close')) %>% 
-  select(state, ST, leq_10_bool, sum_pt_trans_metric) %>% group_by(state, ST, leq_10_bool) %>% 
-  summarize(sum_pt_trans_metric = list(sum_pt_trans_metric)) %>% 
-  spread(leq_10_bool, sum_pt_trans_metric) %>% 
-  group_by(state, ST) %>% 
-  filter(!(is.null(unlist(close)) | is.null(unlist(not_close)))) %>% 
-  mutate(close_median=median(unlist(close)),
-         not_close_median=median(unlist(not_close)),
-         p=wilcox.test(unlist(close),unlist(not_close))$p.value) %>% 
-  mutate(species = ifelse(ST %in% c('ST258','ST307'), '*K. pneumoniae*','*E. hormaechei*'))
-
-fs9 <- pt_snv %>% 
-  mutate(species = gsub('^', '\\*',species),
-         species = gsub('$', '\\*',species),
-         sum_pt_trans_metric = sum_pt_trans_metric + min(sum_pt_trans_metric[sum_pt_trans_metric != 0]/2), # no infinite values in transformation
-         dist_min = dist_min + min(dist_min[dist_min != 0]/2)) %>%  # no infinite values in transformation
-  ggplot(aes(x = sum_pt_trans_metric, y = dist_min)) + 
-  geom_point(aes(color = under_thresh)) + scale_y_log10() + scale_x_log10() +
-  scale_color_manual(
-    values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")
-    ) +
-  geom_text(data = pvals, aes(x = 5e-5, y = 1, label = paste0('p=',signif(p, 1))), hjust = 0) +
-  facet_grid(species+ST~state) + 
-  labs(x = 'Sum(patient flow between facility pair)', y = 'Minimum pairwise SNV distance', color = 'SNV bin') +
-  theme(strip.text.y = element_markdown())
-
-ggsave(plot = fs9, filename = 'figures/FigS9.png', width = 7, height = 4)
-
-fs9
-```
-
-### Prior hospital
-
-#### Individual healthcare exposures (Table 2)
-
-```{r}
-pr_summary <- bind_rows(pr_hosp_pairs_ct, pr_hosp_pairs_mn, pr_hosp_pairs_tn) %>%
-  left_join(snv_dat %>% filter(loc1 != loc2)) %>% 
-  mutate(st1 = ifelse(st1 %in% c("E. hormaechei ST114", "E. hormaechei ST171", "K. pneumoniae ST258", 
-"K. pneumoniae ST307"), st1, 'Other')) %>% 
-  group_by(state1, st1, same_pr_hosp, pairwise_dist <= 10) %>% 
-  summarize(total_l10 = n()) %>% 
-  mutate(same_pr_hosp = ifelse(same_pr_hosp, 'pr_hosp','no_pr_hosp'),
-         `pairwise_dist <= 10` = ifelse(`pairwise_dist <= 10`, 'lt10','gt10')) %>% 
-  pivot_wider(names_from = c(same_pr_hosp, `pairwise_dist <= 10`), values_from = c(total_l10)) %>% 
-  rename(state = state1, st = st1) %>% 
-  select(state, st, pr_hosp_lt10, no_pr_hosp_lt10, pr_hosp_gt10, no_pr_hosp_gt10) %>% 
-  replace(is.na(.), 0)
-
-pr_summary$fisher_p <- sapply(1:nrow(pr_summary), function(x){
-  exact2x2(matrix(unlist(c(pr_summary[x,3:6])), nrow = 2))$p.value %>% signif(digits = 1)
-})
-
-pr_summary <- pr_summary %>% 
-  mutate(pr_hosp_lt10_pct = round(pr_hosp_lt10/(no_pr_hosp_lt10 + pr_hosp_lt10), 1)*100, .after = no_pr_hosp_lt10) %>% 
-  mutate(pr_hosp_gt10_pct = round(pr_hosp_gt10/(no_pr_hosp_gt10 + pr_hosp_gt10), 1)*100, .after = no_pr_hosp_gt10) %>% 
-  mutate('Pairwise SNV distance ≤ 10\nwith shared facility exposure (%)' =
-           paste0(pr_hosp_lt10, '/', (pr_hosp_lt10+no_pr_hosp_lt10), ' (',round(pr_hosp_lt10/(pr_hosp_lt10+no_pr_hosp_lt10),2)*100,')'),
-         'Pairwise SNV distance > 10\nwith shared facility exposure (%)' =
-           paste0(pr_hosp_gt10, '/', (pr_hosp_gt10+no_pr_hosp_gt10), ' (',round(pr_hosp_gt10/(pr_hosp_gt10+no_pr_hosp_gt10),2)*100,')')) %>% 
-  select(c("state", "st", 
-           "Pairwise SNV distance ≤ 10\nwith shared facility exposure (%)", 
-           "Pairwise SNV distance > 10\nwith shared facility exposure (%)",
-           "fisher_p"
-)) %>% 
-  rename(State = state, ST = st, "Fisher's exact p" = fisher_p)
-
-pr_summary
-```
-
-### Individual graph (Figure S10)
-
-```{r}
-fs10 <- graph %>% ggraph(layout = 'kk') + 
-  geom_edge_parallel(aes(color = st1, linetype = same_pr_hosp, alpha = same_pr_hosp),sep = unit(0.4, "mm")) + 
-  geom_node_point(size = 0.5) + 
-  scale_edge_linetype_manual(values = c('dotted','solid')) +
-  facet_nodes(~state, nrow = 3) +
-  scale_edge_alpha_manual(values = c(0.5, 1)) + 
-  scale_edge_color_manual(values = c('*E. hormaechei* ST171'="aquamarine4", '*E. hormaechei* ST114'="aquamarine3",
-                                     '*K. pneumoniae* ST258'="coral4", '*K. pneumoniae* ST307'="coral3",
-                                     'Other'='grey'), 
-                          breaks = c('*E. hormaechei* ST171', '*E. hormaechei* ST114',
-                                      '*K. pneumoniae* ST258', '*K. pneumoniae* ST307', 'Other'), drop = FALSE) +
-  labs(edge_color = 'Isolate pair ≤ 10 SNVs', edge_linetype = 'Patients have\nshared facility exposure',
-       edge_alpha = 'Patients have\nshared facility exposure') +
-  theme_graph() + theme(legend.text = element_markdown())
-
-ggsave(plot = fs10, filename = 'figures/FigS10.png', width = 5, height = 6)
-
-fs10
-```
-
-### Comparison of individual to aggregate (Figure S11)
-
-```{r}
-fs11 <- full_join(pr_summary %>% mutate(ST_orig = ST, ST = gsub('.* ','',ST)) %>% rename('ind_p' = "Fisher's exact p"), 
-          pvals %>% rename(State = state, 'agg_p' = 'p')) %>% 
-  mutate(ST_orig = gsub('^', '\\*', ST_orig),
-         ST_orig = gsub('e ', 'e\\* ', ST_orig),
-         ST_orig = gsub('i ', 'i\\* ', ST_orig)) %>% 
-  filter(ST != 'Other') %>% 
-  ggplot(aes(x = ind_p, y = agg_p, col = State, shape = ST_orig)) +
-  geom_point(size = 3) +
-  scale_color_manual(values = c(CT = "#46f0f0", MN = "#3cb44b", TN = "#4363d8")) +
-  geom_hline(yintercept = 0.05) +
-  geom_vline(xintercept = 0.05) +
-  labs(x = 'Patient transfer p-value from\nindividual-level analysis', 
-       y = 'Patient transfer p-value from\naggregate-level analysis',
-       shape = 'ST') +
-    theme(legend.text = element_markdown())
-
-ggsave(plot = fs11, filename = 'figures/FigS11.png', width = 5, height = 3)
-
-fs11
-```
-
-## MN maps (Figure 3)
-
-```{r}
-# connecting each patient to their closest isolate with a subsequent collection date
-edge_dat <- lapply(unique(c(as.character(snv_dat$isolate1), as.character(snv_dat$isolate2))), function(i){
-  cultdate <- facil_st$Cult_Date[facil_st$Sample_ID == i]
-  subset_dat <- snv_dat %>% filter(pair_type == 'Inter-facility pair' & (isolate1 == i | isolate2 == i) & state1 == 'MN')
-  dates <- sapply(unique(c(as.character(subset_dat$isolate1), as.character(subset_dat$isolate2))), function(j){
-    as.character(facil_st$Cult_Date[facil_st$Sample_ID == j])
-  }) 
-  dat <- NULL
-  if(length(dates) > 0){
-    keepers <- names(dates)[as.Date(dates) > cultdate]
-    dat <- subset_dat %>% 
-      filter(isolate1 %in% keepers | isolate2 %in% keepers) %>% 
-      filter(pairwise_dist == min(pairwise_dist)) %>% 
-      mutate(earlier = as.character(ifelse(isolate2 == i, 'loc2','loc1')),
-             date1 = dates[as.character(isolate1)],
-             date2 = dates[as.character(isolate2)],
-             datediff = abs(as.numeric(as.Date(dates[as.character(isolate1)]) - as.Date(dates[as.character(isolate2)]))))
-  }
-}) %>% bind_rows()
-
-edges_for_plot <- edge_dat %>% 
-  filter(pairwise_dist <= 15) %>% 
-  left_join(facil_st %>% select(Sample_ID, Facility_ID, longitude, latitude), 
-            by = c('isolate1' = 'Sample_ID', 'loc1' = 'Facility_ID')) %>%
-  rename(x = longitude, y = latitude) %>%
-  left_join(facil_st %>% select(Sample_ID, Facility_ID, longitude, latitude), 
-            by = c('isolate2' = 'Sample_ID', 'loc2' = 'Facility_ID')) %>%
-  rename(xend = longitude, yend = latitude) %>% 
-  filter(!(is.na(x) | is.na(xend)) & ST %in% c('ST171','ST258')) %>% 
-  mutate(dist = abs(x - xend)^2 + abs(y - yend)^2)
-
-facilities <- facil_st %>% filter(State == 'MN' & ST %in% c('E. hormaechei ST171', 'K. pneumoniae ST258')) %>% 
-  select(Facility_ID, latitude, longitude) %>% 
-  filter(!duplicated(.) & !is.na(latitude)) 
-```
-
-```{r}
-edges_for_plot_ec <- edges_for_plot %>% filter(ST == 'ST171') %>% 
-  mutate(under_thresh = factor(under_thresh, levels = c("[0,5]", "(5,10]", "(10,15]")))
-
-whole_state_ec <- facil_st %>% 
-  mutate(ST = gsub('cloacae', 'hormaechei', ST)) %>% 
-  filter(ST %in% c('E. hormaechei ST171')) %>%
-  mutate(ST = gsub('.* ST', 'ST', ST)) %>% 
-  group_by(Facility_ID, latitude, longitude, ST)  %>% tally() %>%
-  ggplot(aes(x=longitude, y=latitude), color = "grey99") +
-    geom_curve(aes(x = xend, y = yend, xend = x, yend = y, color = under_thresh),     # draw edges as arcs
-                 alpha = 1,
-             data = edges_for_plot_ec, curvature = 0.33) + 
-  scale_color_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4")) +
-  geom_point(data = facilities, aes(x = longitude, y = latitude), alpha = 0.25, size = 1) +
-  geom_point(aes(size=n), alpha = 1, color = 'white') +
-  geom_point(aes(size=n), fill = adjustcolor(st_cols[c('E. hormaechei ST171')], alpha.f = 0.5), pch = 21) + 
-  scale_size_continuous(breaks = seq(0, 20, 5), limits = c(0,25),
-                        labels = c(0,5,10,15,20),
-                        name = 'Number of\nisolates') +
-  coord_quickmap() +
-  # geom_rect(aes(xmin = -986875.33, xmax = -986875.165, ymin = 896963.81, ymax = 896964.1),
-  geom_rect(aes(xmin = -986875.33+0.58, xmax = -986875.165+0.58, ymin = 896963.81+0.145, ymax = 896964.1+0.145),
-  fill = "transparent", color = "lightgrey", size = 1) +
-  labs(color = 'SNV bin', size = 'Number of\nisolates') +
-  guides(size = guide_legend(override.aes = list(alpha = c(0.25, 1, 1, 1, 1), fill = 'black'))) +
-  theme_void() 
-whole_state_ec
-
-zoom_state_ec <- whole_state_ec + 
-  # coord_quickmap(xlim=c(-986875.33+0.007, -986875.165-0.007), ylim=c(896963.81+0.012, 896964.1-0.012)) +
-  coord_quickmap(xlim=c(-986875.33+0.58+0.007, -986875.165+0.58-0.007), ylim=c(896963.81+0.145+0.013, 896964.1+0.145-0.013)) +
-  theme(legend.position = "none") + 
-  geom_text(aes(label = ifelse(Facility_ID == 'F38', 'W', NA)),hjust=1.5, vjust=2)
-zoom_state_ec
-```
-
-```{r}
-edges_for_plot_kp <- edges_for_plot %>% filter(ST == 'ST258')
-whole_state_kp <- facil_st %>% 
-  filter(ST %in% c('K. pneumoniae ST258')) %>%
-  mutate(ST = gsub('.* ST', 'ST', ST)) %>% 
-  group_by(Facility_ID, latitude, longitude, ST)  %>% tally() %>%
-  ggplot(aes(x=longitude, y=latitude), color = "grey99") +
-    geom_curve(aes(x = xend, y = yend, xend = x, yend = y, color = under_thresh),     # draw edges as arcs
-                 alpha = 1,
-             data = edges_for_plot_kp, curvature = 0.33) + 
-  scale_color_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
-  geom_point(data = facilities, aes(x = longitude, y = latitude), size = 1, alpha = 0.25) +
-  geom_point(aes(size=n), alpha = 1, color = 'white') +
-  geom_point(aes(size=n), fill = adjustcolor(st_cols['K. pneumoniae ST258'], alpha.f = 0.5), pch = 21) + 
-  scale_size_continuous(breaks = seq(5, 20, 5), limits = c(1,25)) +
-  coord_quickmap() +
-  # geom_rect(aes(xmin = -986875.33, xmax = -986875.165, ymin = 896963.81, ymax = 896964.1),
-  geom_rect(aes(xmin = -986875.33+0.58, xmax = -986875.165+0.58, ymin = 896963.81+0.145, ymax = 896964.1+0.145),
-               fill = "transparent", color = "lightgrey", size = 1) +
-  labs(size = 'Number of\nisolates', color = 'SNV bin') +
-  theme_void() 
-whole_state_kp
-
-zoom_state_kp <- whole_state_kp + 
-  # coord_quickmap(xlim=c(-986875.33+0.007, -986875.165-0.007), ylim=c(896963.81+0.012, 896964.1-0.012)) +
-  coord_quickmap(xlim=c(-986875.33+0.58+0.007, -986875.165+0.58-0.007), ylim=c(896963.81+0.145+0.013, 896964.1+0.145-0.013)) +
-  theme(legend.position = "none") +
-   geom_text(aes(label = ifelse(Facility_ID == 'F38', 'W', NA)),hjust=1.5, vjust=2)
-zoom_state_kp
-```
-
-```{r}
-# grob plotting function to plot panel colors
-plot_panel_cols <- function(p, fills){
-  gt <- ggplot_gtable(ggplot_build(p))
-  strip_both <- which(grepl('strip-', gt$layout$name))
-  k <- 1
-  for (i in strip_both) {
-    j <- which(grepl('rect', gt$grobs[[i]]$grobs[[1]]$childrenOrder))
-    gt$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-    k <- k+1
-  }
-  ggplotify::as.ggplot(gt)
-}
-```
-
-```{r}
-bp <- edge_dat %>% pivot_longer(cols = c('loc1','loc2'), names_to = 'fname', values_to = 'loc') %>% 
-  filter(ST %in% c('ST171','ST258') & under_thresh != '(15,∞)') %>%
-  mutate(under_thresh = factor(under_thresh, levels = rev(levels(under_thresh))),
-         ST = factor(ifelse(ST == 'ST171', '*E. hormaechei* ST171', '*K. pneumoniae* ST258'),
-                     levels = c('*E. hormaechei* ST171', '*K. pneumoniae* ST258'))) %>% 
-  filter(fname == earlier) %>% 
-  select(loc, ST, under_thresh) %>%
-  ggplot(aes(x = loc)) + 
-    geom_bar(aes(col = ifelse(loc == 'F38', 'yes','no'))) +
-  geom_bar(aes(fill = under_thresh)) + 
-  facet_grid(ST~.) +
-    scale_fill_manual(values = c('[0,5]' = "#9EBCDA",'(5,10]' = "#BFD3E6", '(10,15]' = "#E0ECF4",'(15,∞)' = "#FBB4AE")) +
-  scale_color_manual(values = c('yes' = 'black','no' = NA)) +
-  labs(x = 'Facility', y = 'Number of isolates closely related\nto an isolate from another facility', fill = 'SNV bin')
-
-bp_labeled <- bp + scale_x_discrete(labels = ifelse(ggplot_build(bp)$layout$panel_params[[1]]$x$breaks == 'F38', 'W','')) +
-  theme_classic() +
-  theme(legend.position = 'none', #axis.ticks.x = element_blank(), 
-        text = element_text(size = 10), 
-        axis.text.x = element_text(angle = 0),
-        strip.background = element_rect(fill="white",linetype='blank'),
-        strip.text.y = element_markdown())
-bp_col <- plot_panel_cols(bp_labeled, adjustcolor(st_cols[c('E. hormaechei ST171','K. pneumoniae ST258')], alpha.f = 0.5)) 
-```
-
-```{r}
-leg <- ggpubr::get_legend(whole_state_ec +
-                            theme(legend.box.margin = margin(0, 0, 0, 50))
-)
-f3 <- plot_grid(leg, NULL,
-          plot_grid(whole_state_ec + theme(legend.position = 'none'), zoom_state_ec,
-                    NULL, NULL,
-                    whole_state_kp + theme(legend.position = 'none'), zoom_state_kp,
-                    NULL, NULL,
-                    nrow = 4, 
-                    rel_heights = c(1,0.05,1,0.1,1,0.05,1,0.1)),
-          plot_grid(NULL, bp_col, NULL, ncol = 1, rel_heights = c(0.1,0.8,0.05)), nrow = 1,
-          rel_widths = c(0.2, 0.05, 1, 0.8) 
-)
-
-ggsave(plot = f3, filename = 'figures/Fig3.png', width = 10, height = 4)
-
-f3
-```
-
-```{r}
-# prior facilities for facility w/ st171 isolates 
-facil_w %>% 
-  mutate(n_patients_tot = n_distinct(Sample_ID)) %>% 
-  filter(!is.na(prior_hosp_alias)) %>% 
-  unique() %>% 
-  group_by(prior_hosp_alias, n_patients_tot) %>% 
-  summarize(n_patients = n()) %>% 
+  mutate(loc1_new = paste(sort(c(loc1, loc2)))[1],
+         loc2_new = paste(sort(c(loc1, loc2)))[2]) %>% 
   ungroup() %>%
-  mutate(prop_patients = n_patients/n_patients_tot) %>% 
-  arrange(-n_patients) 
-```
-
-### Comparison of ST171 and ST258 inter-facility transmission patterns (Figure S12)
-
-```{r}
-## for each facility, how many times it's involved in a link
-
-frac_ec <- lapply(unique(edges_for_plot_ec$loc1, edges_for_plot_ec$loc2), function(x){
-  edges_for_plot_ec %>% filter(pairwise_dist <= 10) %>% 
-  mutate(conn = ifelse(loc1 == !!x | loc2 == !!x, 'connected', 'not')) %>% 
-  group_by(conn, ST) %>% summarize(count = n()) %>% mutate(facil = x)
-}) %>% bind_rows() %>% suppressMessages() %>% ungroup() %>% filter(!duplicated(.)) %>% 
-  group_by(facil) %>% 
-  mutate(n_facil = sum(count), frac = count/n_facil) %>% 
+  group_by(loc1_new, loc2_new) %>% 
+  summarize(count = n()) %>% 
   ungroup() %>% 
-  filter(conn == 'connected')
-
-frac_kp <- lapply(unique(edges_for_plot_kp$loc1, edges_for_plot_kp$loc2), function(x){
-  edges_for_plot_kp %>% filter(pairwise_dist <= 10) %>% 
-  mutate(conn = ifelse(loc1 == !!x | loc2 == !!x, 'connected', 'not')) %>% 
-  group_by(conn, ST) %>% summarize(count = n()) %>% mutate(facil = x)
-}) %>% bind_rows() %>% suppressMessages() %>% ungroup() %>% filter(!duplicated(.)) %>% 
-  group_by(facil) %>% 
-  mutate(n_facil = sum(count), frac = count/n_facil) %>% 
-  ungroup() %>% 
-  filter(conn == 'connected')
-
-max(frac_ec$frac)
-facils_ec <- frac_ec %>% pull(facil)
-ec_perm_cts <- lapply(1:1000, function(x){
-  sample(facils_ec, sum(frac_ec$count), replace = TRUE) %>% table() %>% data.frame() %>% rename(facil = '.') %>% 
-    mutate(tot = sum(Freq), perm = x, frac = Freq/tot) %>% slice_max(Freq) %>% slice(1)
-}) %>% bind_rows()
-
-max(frac_kp$frac)
-facils_kp <- frac_kp %>% pull(facil)
-kp_perm_cts <- lapply(1:1000, function(x){
-  sample(facils_kp, sum(frac_kp$count), replace = TRUE) %>% table() %>% data.frame() %>% rename(facil = '.') %>% 
-    mutate(tot = sum(Freq), perm = x, frac = Freq/tot) %>% slice_max(Freq) %>% slice(1)
-}) %>% bind_rows()
-
-fs12 <- bind_rows(ec_perm_cts %>% select(frac) %>% mutate(ST = 'ST171'),
-          kp_perm_cts %>% select(frac) %>% mutate(ST = 'ST258')) %>% 
-  mutate(ST = factor(ifelse(ST == 'ST171', '*E. hormaechei* ST171', '*K. pneumoniae* ST258'),
-                     levels = c('*E. hormaechei* ST171', '*K. pneumoniae* ST258'))) %>% 
-  ggplot(aes(x = frac)) +
-  geom_histogram(binwidth = 0.05) +
-  geom_point(data = tibble(frac = c(max(frac_ec$frac), max(frac_kp$frac)), 
-                           st = c('*E. hormaechei* ST171', '*K. pneumoniae* ST258')), 
-             aes(y = 0), col = 'red', shape = 18, size = 5) +
-  facet_grid(~st, scales = 'free') +
-  labs(x = 'Maximum proportion of edges connected to a given facility', y = 'Number of edges') +
-  theme(strip.text.x = element_markdown())
-
-sum(ec_perm_cts$frac > max(frac_ec$frac))/length(ec_perm_cts$frac)
-sum(kp_perm_cts$frac > max(frac_kp$frac))/length(kp_perm_cts$frac)
-
-fs12 <- plot_panel_cols(fs12, adjustcolor(st_cols[c('E. hormaechei ST171','K. pneumoniae ST258')], alpha.f = 0.5))
-
-# ggsave(plot = fs12, filename = 'figures/FigS12.png', width = 7, height = 4)
-
-fs12
-```
-
-### Facility W (Figure S13)
+  mutate(tot = sum(count),
+         pct = paste0(round(count/tot*100), '%')) %>% 
+  ggplot(aes(x = loc1_new, y = loc2_new, fill = count, label = pct)) + 
+  geom_tile() +
+  geom_text(col = 'white') +
+  labs(x = 'Location of isolate 1', y = 'Location of isolate 2', 
+       fill = 'Number of pairs\nwithin 10 SNVs')
+```
 
@@
 
-facil_w_ind <- which(names(V(g_mn)) == 'F38')
-
-fs13 <- bind_cols(inds=names(deg), 'Degree'=deg, 'Strength'=stre, 'Betweenness'=betw, 'Closeness'=clos) %>% 
+facil_w_ind <- which(names(V(g_mn)) == 'F38') # F55
+
+f3e <- bind_cols(inds=names(deg), 'Degree'=deg, 'Strength'=stre, 'Betweenness'=betw, 'Closeness'=clos) %>% 
   pivot_longer(cols = -inds) %>% 
+  filter(name == 'Degree') %>% 
   ggplot(aes(x = value)) + geom_histogram(bins = 70) +
@@
     value=c(deg[facil_w_ind], stre[facil_w_ind], betw[facil_w_ind],clos[facil_w_ind]),
-    name=c('Degree','Strength','Betweenness','Closeness')),
+    name=c('Degree','Strength','Betweenness','Closeness')) %>% 
+      filter(name == 'Degree'),
     aes(x = value, y = 0), col = 'red', shape = 18, size = 3) +
@@
     value=c(deg[facil_w_ind], stre[facil_w_ind], betw[facil_w_ind],clos[facil_w_ind]), 
-    name=c('Degree','Strength','Betweenness','Closeness')), 
-    aes(label = 'W', y = 0), col = 'red', size = 3) +
-  labs(x = 'Centrality metric', y = 'Number of facilities') +
-  facet_wrap(~name, scales = 'free') +
+    name=c('Degree','Strength','Betweenness','Closeness')) %>% 
+      filter(name == 'Degree'), 
+    aes(label = 'F38', y = 0), col = 'red', size = 3) +
+  labs(x = 'Degree centrality', y = 'Number of facilities') +
+  # facet_wrap(~name, scales = 'free') +
   theme(plot.margin = margin(0,0.5,0,0, "cm"))
-
-ggsave(plot = fs13, filename = 'figures/FigS13.png', width = 5, height = 3.5)
-
-fs13
+```
+
+```{r}
+mn_pt_trans_all <- get_patient_flow(mn_trans_net_sub,
+                                    unique(facil_st_orig$Facility_ID)[unique(facil_st_orig$Facility_ID) %in% 
+                                                        c(mn_trans_net_sub$source_facil,
+                                                          mn_trans_net_sub$dest_facil)]) %>% 
+  filter(!duplicated(.))
+```
+
+```{r}
+f3f <- mn_pt_trans_all %>% 
+  mutate(of_interest = loc1 == 'F38' & loc2 %in% c('F55', 'F46', 'F34') |
+           loc2 == 'F38' & loc1 %in% c('F55', 'F46', 'F34')) %>% 
+  ggplot(aes(x = sum_transfers)) +
+  geom_histogram(bins = 100) +
+  geom_point(data = mn_pt_trans_all %>% 
+               filter(loc1 == 'F38' & loc2 %in% c('F55', 'F46', 'F34') |
+                        loc2 == 'F38' & loc1 %in% c('F55', 'F46', 'F34')) %>% 
+               mutate(loc = ifelse(loc1 == 'F38', paste0('F38-', loc2), 
+                                   paste0('F38-', loc1))),
+             aes(y = 0), col = 'red') +
+  geom_label_repel(data = mn_pt_trans_all %>% 
+               filter(loc1 == 'F38' & loc2 %in% c('F55', 'F46', 'F34') |
+                        loc2 == 'F38' & loc1 %in% c('F55', 'F46', 'F34')) %>% 
+               mutate(loc = ifelse(loc1 == 'F38', paste0('F38-', loc2), 
+                                   paste0('F38-', loc1))),
+             aes(y = 0, label = loc), col = 'red', 
+             box.padding = 0.5, max.overlaps = Inf) +
+  labs(y = 'Number of facilities', x = 'Number of patient transfers between F38 and another facility')
+
+```
+
+```{r, fig.width=15, fig.height=7}
+f3 <- plot_grid(f3a + theme(legend.position = 'left'), 
+          f3b + theme(legend.position = 'none'), 
+          f3c + theme(legend.position = 'none',
+                      axis.text.x = element_text(size = 5)), 
+          f3d + facet_grid(''~.) + theme(legend.position = 'left'),
+          f3e, 
+          f3f, 
+          labels = 'AUTO', rel_widths = c(1, 0.8, 1))
+
+ggsave(plot = f3, filename = 'figures/Fig3.png', width = 15, height = 7)
+f3
+```
+
+```{r}
+mn_pt_trans_all %>%
+  mutate(n_transfers_f12 = ifelse(is.na(n_transfers_f12), 0, n_transfers_f12),
+         n_transfers_f21 = ifelse(is.na(n_transfers_f21), 0, n_transfers_f21),
+         sum_transfers = n_transfers_f12 + n_transfers_f21) %>% 
+  arrange(sum_transfers) %>% 
+  mutate(rank = row_number(), 
+         tot = nrow(.),
+         percentile = rank/tot) %>% 
+  filter(loc1 == 'F38' & loc2 %in% c('F55', 'F46', 'F34') |
+           loc2 == 'F38' & loc1 %in% c('F55', 'F46', 'F34')) %>% 
+  select(loc1, loc2, rank, tot, percentile)
 ```
@@
 
-```{r}
+```{r, results='hide'}
 prior_hosps_tn_counts <- prior_hosps_tn %>% 
@@
          ST = factor(ST, levels = tn_hm$tree_row$labels),
-         ST = gsub('E. hormaechei', '*E. hormaechei*', ST),
-         ST = gsub('K. pneumoniae', '*K. pneumoniae*', ST)) %>% 
+         ST = gsub('E. hormaechei', 'KPC-Eh', ST),
+         ST = gsub('K. pneumoniae', 'KPC-Kp', ST),
+         ST = factor(ST, levels = rev(c('KPC-Kp ST258', 'KPC-Eh ST171', 'Other', 'KPC-Eh ST114', 'KPC-Kp ST307'))),
+) %>% 
   ggplot(aes(x = Facility, y = ST, fill = count)) +
@@
   labs(fill = 'Count')
-
-tn_facil_hm
 ```
@@
 ```{r}
-graph_tn <- tbl_graph(nodes = tn_facils_st, edges = tn_trans_net_sub %>% filter(!grepl('I', source_facil) & !grepl('I', dest_facil) & n_transfers > 1), directed = TRUE) %>%
+graph_tn <- tbl_graph(nodes = tn_facils_st %>% 
+                        mutate(facil_st_type = gsub('K. pneumoniae', 'KPC-Kp', facil_st_type),
+                               facil_st_type = gsub('E. hormaechei', 'KPC-Eh', facil_st_type),
+                               facil_st_type = gsub(' facility cluster', 
+                                                    '\nfacility cluster', facil_st_type)
+                               ), 
+                      edges = tn_trans_net_sub %>% 
+                        filter(!grepl('I', source_facil) & !grepl('I', dest_facil) & n_transfers > 1),
+                      directed = TRUE) %>%
   induced.subgraph(igraph::degree(.) > 0)
 
-tn_network <- graph_tn %>% ggraph(layout = 'kk') + 
+tn_network <- graph_tn %>% 
+  ggraph(layout = 'kk') + 
   geom_edge_link(aes(color = n_transfers, alpha = n_transfers), arrow = arrow(length = unit(3, 'mm'))) + 
   geom_node_point(aes(color = facil_st_type, size = n_isolates)) + 
-  scale_edge_color_viridis(direction = -1, option = 'E') +
+  # scale_edge_color_viridis(direction = -1, option = 'E') +
+  scale_edge_color_continuous(high = '#00204c', low = '#ffe945') +
   scale_color_manual(values = c(unname(st_cols[1]), 'coral3', 'lightsalmon', 'grey')) +
@@
   labs(color = 'Most common ST', shape = '', size = 'Number of isolates',
-       edge_color = 'Number of transfers (in thousands)')  #+
-  # theme(legend.text = element_markdown())
-tn_network
+       edge_color = 'Number of transfers\n(in thousands)')
 ```
@@
   theme(legend.position = 'left')
-
-tn_facil_transfers
-```
-
-```{r}
-plot_grid(plot_grid(tn_facil_hm + theme(legend.position = 'left'), tn_facil_transfers, nrow = 2, align = 'hv', labels = c('A', 'C')), tn_network, nrow = 1, rel_widths = c(0.8, 1), labels = c('', 'B')) %>%
-  ggsave(filename = 'figures/Fig4.png', width = 12, height = 5)
-```
+```
+
+```{r}
+tn_trans_net_sub %>% 
+  left_join(tn_facils_st %>% rename(source_of_interest = of_interest) %>% 
+              select(Facility_ID, source_of_interest), 
+            by = c(source_facil = 'Facility_ID')) %>% 
+  left_join(tn_facils_st %>% rename(dest_of_interest = of_interest) %>% 
+              select(Facility_ID, dest_of_interest), 
+            by = c(dest_facil = 'Facility_ID')) %>% 
+  replace(is.na(.), 0) %>% 
+  mutate(of_interest = source_of_interest + dest_of_interest) %>% 
+  filter(!grepl('I', source_facil) & !grepl('I', dest_facil) & n_transfers > 1) %>% 
+  filter(source_of_interest | dest_of_interest) %>% 
+  ungroup() %>% 
+  summarize(wilcox_p = wilcox.test(n_transfers[of_interest == 1], n_transfers[of_interest == 2])$p.value)
+```
+
+```{r, fig.width=12, fig.height=5}
+f4 <- plot_grid(plot_grid(tn_facil_hm + theme(legend.position = 'left'), tn_facil_transfers, nrow = 2, align = 'hv', labels = c('A', 'C')), tn_network, nrow = 1, rel_widths = c(0.8, 1), labels = c('', 'B'))
+
+ggsave(filename = 'figures/Fig4.png', plot = f4, width = 12, height = 5)
+f4
+```
+
+## Supplemental file with isolate IDs and information
+
+```{r}
+sample_metadat <- read_csv('data/eip-metadat-short.csv') %>% 
+  rename(Pat_ID = Patient_ID) %>% 
+  mutate(Species = gsub('cloacae', 'hormaechei', Species),
+         ST = gsub('cloacae', 'hormaechei', ST)) %>% 
+  left_join(facil_st) %>% 
+  select(-c(latitude, longitude))
+
+all_sample_info <- sample_metadat %>% 
+  select(Fac_ID, Facility_ID) %>% 
+  filter(!is.na(Facility_ID)) %>% 
+  unique() %>% 
+  right_join(sample_metadat %>% select(-Facility_ID)) %>% 
+  select(-Patient_ID) %>% 
+  left_join(sample_metadat %>% 
+              select(Pat_ID, Patient_ID) %>% 
+              filter(!is.na(Patient_ID)) %>% 
+              unique()) %>% 
+  mutate(Cult_Year = gsub('-.*', '', Cult_Date))
+
+all_sample_info_not_tn <- all_sample_info %>% 
+  filter(State != 'TN') %>% 
+  arrange(Patient_ID) %>% 
+  mutate(Pat_ID = factor(Pat_ID, levels = unique(Pat_ID)),
+         Patient_ID = ifelse(is.na(Patient_ID), paste0('P', as.numeric(factor(Pat_ID))), Patient_ID),
+         Patient_ID = ifelse(Patient_ID == 'PNA', NA, Patient_ID)) %>% 
+  select(Sample_ID, Patient_ID, Facility_ID, Cult_Year, State, Species, ST) %>% 
+  arrange(Sample_ID)
+
+all_sample_info_not_tn %>% 
+  write_csv('data/supp_file_1.csv')
+
+# to not submit to ncbi (because previously submitted)
+dont_submit <- all_sample_info_not_tn %>% 
+  filter(State == 'MN' & Species == 'E. hormaechei' & Cult_Year < 2013) %>% 
+  pull(Sample_ID)
+
+dont_submit %>% 
+  cat(sep = ',')
+```
+
+```{r}
+all_sample_info %>% 
+  select(Sample_ID, State, Species, ST) %>% 
+  left_join(all_sample_info_not_tn) %>% 
+  filter(!Sample_ID %in% dont_submit) %>% 
+  rename(sample_name = Sample_ID,
+         organism = Species,
+         collection_date = Cult_Year) %>% 
+  mutate(bioproject_accession = 'PRJNA873034',
+         isolate = paste0(sample_name, '_', Patient_ID, '_', Facility_ID),
+         strain = ifelse(grepl('ST', ST), gsub('.*ST', 'ST', ST), NA),
+         collected_by = case_when(State == 'MN' ~ 'Minnesota Department of Health',
+                                  State == 'TN' ~ 'Tennessee Department of Health',
+                                  State == 'CT' ~ 'Connecticut Department of Public Health'),
+         geo_loc_name = paste0('USA: ', State),
+         host = 'Homo sapiens',
+         host_disease = 'not applicable', 
+         isolation_source = 'not collected',
+         lat_lon = 'not collected') %>% 
+  select(sample_name,	bioproject_accession,	organism,	strain,	isolate,	collected_by,	collection_date,
+         geo_loc_name,	host,	host_disease,	isolation_source,	lat_lon) %>% 
+  write_csv('data/sra_submission_info.csv')
+```

### 2023-06-15 manuscript-figures.Rmd
Updates based on review responses

@@
 tr_ec <- read.tree('data/eh_tr.tree')
+# plasmid clusters
+mob_plasmids <- read_rds('data/mobtyper/CRE_EIP_mobtyper_mash_nn_plasmid_matrix_050923.RDS')
+mob_clusters <- read_rds('data/mobtyper/CRE_EIP_mobtyper_plasmid_cluster_matrix_050923.RDS')
+mob_results <- read_rds('data/mobtyper/CRE_EIP_mobtyper_results_clean_KPC_050923.RDS')
+# mob_contigs <- read_rds('data/mobtyper/CRE_EIP_mobtyper_contigs_clean_KPC_050923.RDS') 
 # clinical factors
@@
 
+### Number of isolates
+
+```{r}
+nrow(facil_st)
+```
+### Number of patients
+
+```{r}
+n_distinct(facil_st$Patient_ID)
+```
+
+### Number of facilities
+
+```{r}
+n_distinct(facil_st$Facility_ID)
+```
+
 ### Number of facilities with CRE per state
@@
   group_by(State) %>% 
-  summarize(n_facil = n_distinct(Facility_ID))
+  summarize(n_isolates = n(),
+            n_facil = n_distinct(Facility_ID))
 ```
@@
 
-### Isolate counts by state
-
-```{r}
-state_dates <- facil_st %>% mutate(Cult_Date = gsub('-.*|.*/','', Cult_Date)) %>% 
-  filter(!is.na(Cult_Date)) %>% 
-  group_by(State, Cult_Date) %>% summarize() %>% 
-  group_by(State) %>% summarize(min_cult_date = min(Cult_Date),
-                                max_cult_date = max(Cult_Date))
-
-facil_st %>% 
-  mutate(State = factor(State, levels = c('MN','TN','CT'))) %>% 
-  group_by(State) %>% 
-  count() %>% full_join(state_dates, .) %>% 
-  suppressMessages()
-```
-
 ### Isolate overview (Figure S1A)
@@
 ```{r}
-fs1a <- facil_st %>% 
+fs1a <- facil_st_orig %>% 
   mutate(ST = ifelse(ST %in% c('E. hormaechei ST114','E. hormaechei ST171', 
@@
 
-```{r, fig.width=5, fig.height=5}
-fs1 <- plot_grid(fs1a, fs1b + theme(legend.position = 'none'), ncol = 1, labels = 'AUTO')
-
-ggsave(plot = fs1, filename = 'figures/FigS1.png', width = 5, height = 5)
+### Number of distinct STs by state and year
+
+```{r}
+fs1c <- facil_st_orig %>% 
+  group_by(State, cult_year, Species, ST) %>% 
+  filter(!is.na(cult_year)) %>% 
+  summarize() %>% 
+  group_by(State, cult_year, Species) %>% 
+  count() %>% 
+  mutate(Species = ifelse(grepl('hormaechei', Species), 'KPC-Eh', 'KPC-Kp')) %>% 
+  ggplot(aes(x = cult_year, y = n, fill = Species)) +
+  facet_grid(~State, scales = 'free', space = 'free') +
+  geom_col() +
+  scale_fill_manual(values = rev(c('coral2', 'aquamarine2'))) +
+  labs(x = 'Year', y = 'Number of distinct STs')
+```
+
+```{r, fig.width=5.5, fig.height=7}
+fs1 <- plot_grid(fs1a, fs1b + theme(legend.position = 'none'), fs1c, ncol = 1, labels = 'AUTO')
+
+ggsave(plot = fs1, filename = 'figures/FigS1.png', width = 5.5, height = 7)
+ggsave(plot = fs1, filename = 'figures/FigS1.tiff', width = 5.5, height = 7)
 
@@
 
-```{r, fig.width=10, fig.height=8}
-fs2 <- plot_grid(fs2ab, fs2c, labels = c('', 'C'), label_x = -0.1)
-ggsave('figures/FigS2.png', width = 10, height = 8)
-fs2
-```
-
-## Evidence of importation
-
-```{r}
+```{r}
+# identifying phylogenetic transition events
 kpc_bin <- metadat_complete %>% select(isolate_no, kpc) %>% drop_na() 
@@
 keepers <- facil_st %>% 
-  filter(!(duplicated(Patient_ID) & duplicated(ST))) %>% 
+  filter(!(duplicated(Patient_ID) & duplicated(ST) & duplicated(Facility_ID))) %>% 
   pull(Sample_ID)
 
+# phylogenetic clusters
 transitions <- lapply(1:length(tree_list), function(x){
@@
 
+```{r}
+# plasmid info 
+mob_plasmids_long <- mob_plasmids %>% 
+  rownames_to_column(var = 'isolate_no') %>% 
+  pivot_longer(!isolate_no, names_to = 'mash_nearest_neighbor', values_to = 'present') %>% 
+  left_join(mob_results %>% 
+              filter(!grepl('novel', primary_cluster_id)) %>% 
+              select(primary_cluster_id, mash_nearest_neighbor) %>% distinct()) %>% 
+  left_join(mob_results %>% 
+              filter(!grepl('novel', primary_cluster_id)) %>% 
+              select(isolate_no, primary_cluster_id, mash_nearest_neighbor, KPC_present)) %>% 
+  mutate(state = gsub('_.*', '', isolate_no), .before = 2) %>% 
+  group_by(mash_nearest_neighbor) %>% 
+  filter(any(sum(KPC_present, na.rm = TRUE) > 0)) %>% 
+  left_join(read_csv('data/isolate_data.csv') %>% 
+  select(Sample_ID, Species, ST) %>% 
+  rename(isolate_no = Sample_ID)) %>% 
+  mutate(ST_orig = ST,
+         ST = case_when(is.na(ST) ~ 'Other species',
+                        ST %in% dominant_sts ~ ST, 
+                        TRUE ~ 'Other')) %>% 
+  ungroup()
+
+cluster_info <- mob_plasmids_long %>% 
+  filter(KPC_present == 1) %>%
+  rename(plasmid_cluster = primary_cluster_id) %>% 
+  inner_join(transitions %>% 
+               mutate(phylo_cluster = paste0(state, ST, cluster, class2)) %>% 
+               select(id, state, phylo_cluster, class2) %>% 
+               rename(isolate_no = id)) %>% 
+  select(isolate_no, state, ST, class2, phylo_cluster, plasmid_cluster) %>% 
+  group_by(state, ST, phylo_cluster, plasmid_cluster) %>% 
+  mutate(cluster_id = paste0(state, ST, phylo_cluster, plasmid_cluster),
+         cluster_size = n()) %>% 
+  ungroup()
+
+nodes <- cluster_info %>%
+               select(ST, state, class2, phylo_cluster, plasmid_cluster) %>%
+               pivot_longer(!c(ST, class2, state), names_to = 'node_type', values_to = 'clust_id') %>%
+  group_by(node_type, clust_id) %>% 
+  summarize(n_in_clust = n()) %>% 
+  ungroup() %>% 
+  arrange(clust_id) %>% 
+  left_join(cluster_info %>% select(ST, class2, phylo_cluster) %>% rename(clust_id = phylo_cluster) %>% distinct()) %>% 
+  mutate(index = as.numeric(factor(clust_id)),
+         node_type = ifelse(grepl('phylo', node_type), ST, 'Plasmid'),
+         node_import = factor(ifelse(is.na(class2), 'Plasmid', as.character(class2)),
+                              levels = c('Yes', 'No', 'Plasmid')))
+
+edges <- cluster_info %>% 
+  left_join(nodes %>% rename(phylo_cluster = clust_id) %>% rename(from = index) %>% select(-c(node_type, n_in_clust, ST, class2, node_import))) %>%
+  left_join(nodes %>% rename(plasmid_cluster = clust_id) %>% rename(to = index) %>% select(-c(node_type, n_in_clust, ST, class2, node_import))) %>% 
+  group_by(from, to, phylo_cluster, plasmid_cluster, state, ST, class2) %>% 
+  summarize(n_in_grp = n())
+
+clust_graph <- tbl_graph(directed = FALSE, 
+             nodes = nodes,
+             edges = edges)
+```
+
+```{r}
+fs2d <- cluster_info %>%
+  group_by(phylo_cluster) %>%
+  filter(n() > 1) %>%
+  group_by(state, ST, phylo_cluster, plasmid_cluster, cluster_size) %>% 
+  summarize() %>% 
+  ggplot(aes(x = phylo_cluster, y = cluster_size, group = plasmid_cluster, fill = ST)) +
+  facet_grid(~state, scales = 'free', space = 'free') +
+  geom_col(col = 'white') +
+  scale_fill_manual(values = st_cols,
+                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
+                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
+                               'Other'), 
+                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
+                               'KPC-Kp ST258', 'KPC-Kp ST307',
+                               'Other'),
+                    drop = FALSE) +
+  theme(axis.text.x = element_blank()) +
+  labs(x = 'Phylogenetic cluster', y = 'Number of isolates')
+```
+
+```{r, fig.width=8, fig.height=10}
+fs2 <- plot_grid(plot_grid(fs2ab, fs2c, labels = c('', 'C'), label_x = -0.1),
+                  fs2d, ncol = 1, labels = c('', 'D'), rel_heights = c(2, 0.5))
+ggsave('figures/FigS2.png', width = 8, height = 10)
+ggsave('figures/FigS2.tiff', width = 8, height = 10)
+fs2
+```
+
+```{r}
+cluster_info %>%
+  group_by(phylo_cluster) %>%
+  filter(n() > 1) %>%
+  group_by(state, ST, phylo_cluster, plasmid_cluster, cluster_size) %>% 
+  summarize() %>% 
+  group_by(state, phylo_cluster) %>% 
+  tally() %>% 
+  group_by(state, n > 1) %>% 
+  tally() %>% 
+  # group_by(state) %>%
+  # group_by(`n > 1`) %>% 
+  ungroup() %>%
+  mutate(tot = sum(n))
+```
+
+## Evidence of importation
+
 ### Evidence of importation (Figure 1)
@@
 ```{r}
-transitions <- transitions %>% 
-  mutate(ST = ifelse(grepl('ST258', ST_orig), 'K. pneumoniae ST258', as.character(ST))) 
-clust_cols <- st_cols[as.character(transitions$ST)]
-names(clust_cols) <- paste0(transitions$state, transitions$ST, transitions$cluster)
+# cluster_info <- cluster_info %>% 
+  # mutate(ST = ifelse(grepl('ST258', ST_orig), 'K. pneumoniae ST258', as.character(ST))) 
+clust_cols <- st_cols[as.character(cluster_info$ST)]
+names(clust_cols) <- cluster_info$cluster_id #paste0(cluster_info$state, cluster_info$ST, cluster_info$cluster)
 
@@
 
+```{r}
+set.seed(2)
+clust_graph_ct <- clust_graph %>%
+  activate(edges) %>%
+  filter(state == 'CT') %>% 
+  activate(nodes) %>% 
+  filter(centrality_degree() > 0) %>% 
+  mutate(n_in_clust = centrality_degree(weights = n_in_grp)) %>% 
+  ggraph(layout = 'fr') +
+    geom_edge_link(aes(alpha = n_in_grp)) +
+    geom_node_point(aes(col = node_type, size = n_in_clust)) +
+  theme_graph() +
+  scale_color_manual(values = c(st_cols, Plasmid = 'grey20'),
+                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
+                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
+                               'Other', 'Plasmid'), 
+                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
+                               'KPC-Kp ST258', 'KPC-Kp ST307',
+                               'Other', 'Plasmid'),
+                    drop = FALSE) +
+  scale_edge_alpha_continuous(limits = c(0, 20), breaks = seq(5, 20, by = 5)) +
+  scale_size_continuous(limits = c(0, 50), breaks = c(10, 20, 30), labels = c(10, 20, 30)) +
+  labs(col = 'Cluster type', size = 'Number of isolates', edge_alpha = 'Number of\nshared isolates') +
+  theme(title = element_text(size = 10))
+
+clust_graph_mn <- clust_graph %>%
+  activate(edges) %>%
+  filter(state == 'MN') %>% 
+  activate(nodes) %>% 
+  filter(centrality_degree() > 0) %>% 
+  mutate(n_in_clust = centrality_degree(weights = n_in_grp)) %>% 
+  ggraph(layout = 'fr') +
+    geom_edge_link(aes(alpha = n_in_grp)) +
+    geom_node_point(aes(col = node_type, size = n_in_clust)) +
+  theme_graph() +
+  scale_color_manual(values = c(st_cols, Plasmid = 'grey20'),
+                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
+                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
+                               'Other', 'Plasmid'), 
+                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
+                               'KPC-Kp ST258', 'KPC-Kp ST307',
+                               'Other', 'Plasmid'),
+                    drop = FALSE) +
+  scale_edge_alpha_continuous(limits = c(0, 20), breaks = seq(5, 20, by = 5)) +
+  scale_size_continuous(limits = c(0, 50), breaks = c(10, 20, 30), labels = c(10, 20, 30)) +
+  labs(col = 'Cluster type', size = 'Number of isolates', edge_alpha = 'Number of\nshared isolates')
+
+clust_graph_tn <- clust_graph %>%
+  activate(edges) %>%
+  filter(state == 'TN') %>% 
+  activate(nodes) %>% 
+  filter(centrality_degree() > 0) %>%
+  mutate(n_in_clust = centrality_degree(weights = n_in_grp)) %>% 
+  ggraph(layout = 'fr') +
+    geom_edge_link(aes(alpha = n_in_grp)) +
+    geom_node_point(aes(col = node_type, size = n_in_clust)) +
+  theme_graph() +
+  scale_color_manual(values = c(st_cols, Plasmid = 'grey20'),
+                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
+                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
+                               'Other', 'Plasmid'), 
+                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
+                               'KPC-Kp ST258', 'KPC-Kp ST307',
+                               'Other', 'Plasmid'),
+                    drop = FALSE) +
+  scale_edge_alpha_continuous(limits = c(0, 20), breaks = seq(5, 20, by = 5)) +
+  scale_size_continuous(limits = c(0, 50), breaks = c(10, 20, 30), labels = c(10, 20, 30)) +
+  labs(col = 'Cluster type', size = 'Number of isolates', edge_alpha = 'Number of\nshared isolates')
+
+f1a <- plot_grid(plot_grid(clust_graph_ct + 
+                   theme(legend.position = 'none', 
+                         plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
+            labs(title = 'CT'), 
+          clust_graph_mn + 
+            theme(legend.position = 'none', 
+                  plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
+            labs(title = 'MN'), 
+          clust_graph_tn + 
+            theme(legend.position = 'none', 
+                  plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
+            labs(title = 'TN'), 
+          ncol = 1), get_legend(clust_graph_mn + theme(legend.position = 'right')), nrow = 1) %>% 
+  suppressWarnings()
+```
+
 ```{r, fig.width=7, fig.height=4}
-f1 <- transitions %>% 
+f1b <- cluster_info %>% 
+  arrange(ST, -cluster_size) %>% 
   ggplot(aes(x = class2, 
-             fill = factor(paste0(state, ST, cluster), 
-                           levels = unique(paste0(state, ST, cluster))))) + 
+             fill = factor(cluster_id, #paste0(state, ST, cluster), 
+                           levels = unique(cluster_id)))) + #unique(paste0(state, ST, cluster))))) + 
   geom_bar(color = 'white', position = position_stack(reverse = TRUE)) + 
@@
   theme(legend.text = ggtext::element_markdown())
-
-ggsave(plot = f1, filename = 'figures/Fig1.png', width = 7, height = 4)
-
-f1
+```
+
+```{r, fig.width=12, fig.height=6}
+# f1 <- plot_grid(f1a, plot_grid(f1b, NULL, nrow = 1, rel_widths = c(2, 0.5)), 
+#                 ncol = 1, rel_heights = c(2, 1), labels = c('A', 'B'))
+
+networks <- plot_grid(plot_grid(clust_graph_ct + 
+                   theme(legend.position = 'none', 
+                         plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
+            labs(title = 'CT'), 
+          clust_graph_mn + 
+            theme(legend.position = 'none', 
+                  plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
+            labs(title = 'MN'), nrow = 1), 
+          plot_grid(NULL,
+          clust_graph_tn + 
+            theme(legend.position = 'none', 
+                  plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
+            labs(title = 'TN'),
+          NULL, nrow = 1, rel_widths = c(0.5, 1, 0.5)),
+          ncol = 1)
+
+f1 <- plot_grid(networks, 
+                get_legend(clust_graph_mn + theme(legend.position = 'right')),
+                f1b + theme(legend.position = 'none'), labels = c('A', '', 'B'), 
+                nrow = 1, rel_widths = c(2, 0.5, 2)) %>% 
+  suppressWarnings()
+
+ggsave(plot = f1, filename = 'figures/Fig1.png', width = 12, height = 6)
+ggsave(plot = f1, filename = 'figures/Fig1.tiff', width = 12, height = 6)
 ```
@@
 ggsave(plot = fs3, filename = 'figures/FigS3.png', width = 7.5, height = 10.5)
+ggsave(plot = fs3, filename = 'figures/FigS3.tiff', width = 7.5, height = 10.5)
 
@@
 }) %>% bind_rows() %>% suppressMessages() %>% suppressWarnings()
-```
-
-```{r}
-min_dat %>% 
-  filter(!is.na(`Inter-facility pair`) & !is.na(`Intra-facility pair`)) %>% 
-  group_by(ST_orig, state1) %>% 
-  summarize(n = n(), 
-            median_intra = median(`Intra-facility pair`),
-            median_inter = median(`Inter-facility pair`),
-            wilcox_p = wilcox.test(`Intra-facility pair`, `Inter-facility pair`, 
-                                   paired = TRUE)$p.value) %>% 
-  arrange(-n)
 ```
@@
 
-```{r}
-bind_rows(pr_hosp_pairs_ct, pr_hosp_pairs_mn %>% filter(!is.na(state1)), pr_hosp_pairs_tn) %>%
-  left_join(snv_dat %>% filter(loc1 != loc2)) %>% 
-  mutate(st1 = ifelse(st1 %in% c("E. hormaechei ST114", "E. hormaechei ST171", "K. pneumoniae ST258", "K. pneumoniae ST307"), st1, 'Other')) %>% 
-  left_join(pr_hops_pvals) %>% 
-  filter(state1 == 'MN' & st1 == 'E. hormaechei ST171' |
-           state1 == 'TN' & st1 %in% c('K. pneumoniae ST258', 'K. pneumoniae ST307', 'E. hormaechei ST114')) %>% 
-  select(state1, ST, n_shared, n_not_shared) %>% 
-  unique()
-```
-
 ```{r, fig.width=7, fig.height=7}
@@
 ggsave(plot = f2, filename = 'figures/Fig2.png', width = 7, height = 7)
+ggsave(plot = f2, filename = 'figures/Fig2.tiff', width = 7, height = 7)
 
@@
 ggsave(plot = fs4, filename = 'figures/FigS4.png', width = 7, height = 7)
+ggsave(plot = fs4, filename = 'figures/FigS4.tiff', width = 7, height = 7)
 fs4
@@
 mn_snv_dat_orig %>% 
-  filter(ST %in% c('ST171') & pairwise_dist <= 10 &
+  filter(ST %in% c('ST171') & pairwise_dist <= 5 &
            !is.na(loc1) & !is.na(loc2) & loc1 != loc2) %>% 
@@
 mn_snv_dat_orig %>% 
-  filter(ST %in% c('ST171') & pairwise_dist <= 10 &
+  filter(ST %in% c('ST171') & pairwise_dist <= 5 &
            !is.na(loc1) & !is.na(loc2) & loc1 == loc2) %>% 
@@
 f3d <- mn_snv_dat_orig %>% 
-  filter(ST %in% c('ST171') & pairwise_dist <= 10 &
+  filter(ST %in% c('ST171') & pairwise_dist <= 5 &
            !is.na(loc1) & !is.na(loc2) & loc1 != loc2) %>% 
@@
          under_thresh = factor(under_thresh, 
-                               levels = rev(unique(levels(under_thresh)))),
-         loc1 = ifelse(loc1 %in% c('F38', 'F55', 'F34', 'F46'), loc1, 'Other'),
-         loc2 = ifelse(loc2 %in% c('F38', 'F55', 'F34', 'F46'), loc2, 'Other')
+                               levels = rev(unique(levels(under_thresh))))
          ) %>% 
   rowwise() %>% 
-  mutate(loc1_new = paste(sort(c(loc1, loc2)))[1],
-         loc2_new = paste(sort(c(loc1, loc2)))[2]) %>% 
+  mutate(f38_loc = any(c(loc1, loc2) == 'F38'),
+         loc1_new = ifelse(f38_loc, 
+                           c(loc1, loc2)[c(loc1, loc2) == 'F38'],
+                           paste(sort(c(loc1, loc2)))[1]),
+         loc2_new = ifelse(f38_loc, 
+                           c(loc1, loc2)[c(loc1, loc2) != 'F38'],
+                           paste(sort(c(loc1, loc2)))[2]),
+         loc1_new = ifelse(loc1_new %in% c('F38'), loc1_new, 'Other')) %>% #ifelse(loc2_new %in% c('F38', 'F55', 'F34', 'F46'), loc2_new, 'Other')) %>%
+  filter(loc1_new == 'F38') %>% 
   ungroup() %>%
@@
          pct = paste0(round(count/tot*100), '%')) %>% 
+  arrange(count/tot) %>% 
+  mutate(loc2_new = factor(loc2_new, levels = unique(loc2_new))) %>% 
   ggplot(aes(x = loc1_new, y = loc2_new, fill = count, label = pct)) + 
   geom_tile() +
-  geom_text(col = 'white') +
+  geom_text(col = 'white', size = 3) +
   labs(x = 'Location of isolate 1', y = 'Location of isolate 2', 
-       fill = 'Number of pairs\nwithin 10 SNVs')
+       fill = 'Number of pairs\nwithin 5 SNVs')
 ```
@@
     aes(label = 'F38', y = 0), col = 'red', size = 3) +
-  labs(x = 'Degree centrality', y = 'Number of facilities') +
+  labs(x = 'Facility degree centrality based on patient transfers', y = 'Number of facilities') +
   # facet_wrap(~name, scales = 'free') +
@@
   ggplot(aes(x = sum_transfers)) +
-  geom_histogram(bins = 100) +
+  geom_density() +
   geom_point(data = mn_pt_trans_all %>% 
@@
   labs(y = 'Number of facilities', x = 'Number of patient transfers between F38 and another facility')
-
-```
-
-```{r, fig.width=15, fig.height=7}
+```
+
+```{r, fig.width=7, fig.height=8}
 f3 <- plot_grid(f3a + theme(legend.position = 'left'), 
@@
 
-ggsave(plot = f3, filename = 'figures/Fig3.png', width = 15, height = 7)
+f3 <- plot_grid(f3a, 
+          plot_grid(f3b + theme(legend.position = 'none'), 
+                    f3c + theme(legend.position = 'none',
+                                axis.text.x = element_text(size = 5)),
+          nrow = 1, rel_widths = c(0.5, 1), labels = c('B', 'C')),
+          plot_grid(f3d + 
+                      theme(legend.position = 'left'),
+          f3e, 
+          labels = c( 'D', 'E'), rel_widths = c(1, 1, 1), nrow = 1, align = 'h', axis = 'trbl'),
+          ncol = 1, labels = c('A', '', ''), rel_heights = c(1,1,1))
+
+ggsave(plot = f3, filename = 'figures/Fig3.png', width = 7, height = 8)
+ggsave(plot = f3, filename = 'figures/Fig3.tiff', width = 7, height = 8)
 f3
@@
 prior_hosps_tn_counts <- prior_hosps_tn %>% 
+  select(Sample_ID, Fac_ID) %>% 
+  unique() %>% 
+  rename(Facility = Fac_ID) %>% 
+  bind_rows(prior_hosps_tn %>% 
+              select(Sample_ID, Prior_facil) %>% 
+              rename(Facility = Prior_facil)) %>% 
+  left_join(facil_st %>% 
+              mutate(ST = ifelse(ST %in% c('E. hormaechei ST171','E. hormaechei ST114',
+                                 'K. pneumoniae ST258','K. pneumoniae ST307', 
+                                 'E. cloacae ST114', 
+                                 'ST258', 'ST307', 'ST114'),
+                       ST, 'Other'))) %>% 
+  filter(!is.na(ST)) %>% 
+  group_by(ST, Facility) %>% 
+  summarize(count = n()) %>% 
+  arrange(-count)
+
+tn_hm <- prior_hosps_tn_counts %>% 
+  pivot_wider(names_from = Facility, values_from = count, values_fill = 0) %>% 
+  data.frame(row.names = .$ST) %>% 
+  select(-ST) %>% 
+  pheatmap(show_colnames = FALSE, fontsize_col = 8)
+
+hm_dat <- prior_hosps_tn %>% 
   select(Sample_ID, Fac_ID) %>% 
@@
                        ST, 'Other'))) %>% 
-  filter(!is.na(ST)) %>% 
-  group_by(ST, Facility) %>% 
+  left_join(cluster_info %>% rename(Sample_ID = isolate_no) %>% select(-ST)) %>% 
+  filter(!is.na(Facility) & !is.na(plasmid_cluster)) %>% 
+  group_by(Facility, ST) %>% 
+  mutate(facil_st_count = n()) %>%
+  group_by(plasmid_cluster, Facility, ST, facil_st_count) %>% 
   summarize(count = n()) %>% 
-  arrange(-count)
-
-tn_hm <- prior_hosps_tn_counts %>% 
-  pivot_wider(names_from = Facility, values_from = count, values_fill = 0) %>% 
-  data.frame(row.names = .$ST) %>% 
-  select(-ST) %>% 
-  pheatmap(show_colnames = FALSE, fontsize_col = 8)
-
-tn_facil_hm <- prior_hosps_tn_counts %>% 
   mutate(Facility = factor(Facility, levels = rev(tn_hm$tree_col$labels[tn_hm$tree_col$order]) %>% gsub('TDH.', 'TDH-', .) %>% gsub('LONG.', 'LONG-', .) %>% gsub('VA.', 'VA-', .) %>% gsub('\\.', ' ', .)),
@@
          ST = gsub('K. pneumoniae', 'KPC-Kp', ST),
-         ST = factor(ST, levels = rev(c('KPC-Kp ST258', 'KPC-Eh ST171', 'Other', 'KPC-Eh ST114', 'KPC-Kp ST307'))),
+         ST = ifelse(ST == 'KPC-Eh ST171', 'Other', ST),
+         ST = factor(ST, levels = rev(c('KPC-Kp ST258', 'KPC-Eh ST114', 'KPC-Kp ST307', 'Other'))),
 ) %>% 
+  group_by(Facility) %>% 
+  mutate(dominant_st = case_when(is.infinite(max(facil_st_count[ST != 'Other'])) ~ 'Other\nmost common',
+                         max(facil_st_count[ST == 'KPC-Kp ST258']) == max(facil_st_count[ST != 'Other']) ~ 'KPC-Kp ST258\nmost common',
+                         max(facil_st_count[ST == 'KPC-Kp ST307']) == max(facil_st_count[ST != 'Other']) ~ 'KPC-Kp ST307\nmost common',
+                         max(facil_st_count[ST == 'KPC-Eh ST114']) == max(facil_st_count[ST != 'Other']) ~ 'KPC-Eh ST114\nmost common',
+                         TRUE ~ 'Other\nmost common'),
+         dominant_st = factor(dominant_st, levels = c('KPC-Kp ST258\nmost common', 'KPC-Kp ST307\nmost common',
+                                                      'KPC-Eh ST114\nmost common', 'Other\nmost common'))) %>% 
+  # filter(!is.na(plasmid_cluster)) %>% 
+  group_by(plasmid_cluster) %>% 
+  mutate(plasmid_cluster = ifelse(sum(count) >= 10, plasmid_cluster, 'Other\nplasmid'))
+
+tn_facil_hm <- hm_dat %>% 
   ggplot(aes(x = Facility, y = ST, fill = count)) +
+  facet_grid(plasmid_cluster~., scales = 'free_x', space = 'free_x') +
   geom_tile() +
-  geom_rect(aes(xmin = 0+0.5, xmax = 7+0.5, ymin = 0+0.5, ymax = 3+0.5), fill = NA, col = 'black', size = 1, linetype = 'dotted') +
+  geom_vline(xintercept = 7+0.5, col = 'black', size = 1) +
+  # geom_rect(aes(xmin = 0+0.5, xmax = 7+0.5, ymin = 0+0.5, ymax = 4+0.5), fill = NA, col = 'black', size = 1, linetype = 'dotted') +
   scale_fill_viridis_c(direction = -1, breaks = scales::pretty_breaks(), option = 'magma') +
-  theme_classic() +
-  theme(axis.text.x = element_blank(), axis.text.y = element_markdown()) +
+  # theme_classic() +
+  theme(axis.text.x = element_blank(), axis.text.y = element_markdown(),
+        panel.grid = element_blank()) +
   labs(fill = 'Count')
@@
                                  'K. pneumoniae ST258','K. pneumoniae ST307', 
-                                 'E. cloacae ST171', 'E. cloacae ST114', 
-                                 'ST258', 'ST307', 'ST114', 'ST171'),
+                                 'E. cloacae ST114', 
+                                 'ST258', 'ST307', 'ST114'),
                        ST, 'Other')) %>% 
@@
 ```{r}
+tn_facils_st <- facil_st %>% 
+  filter(State == 'TN') %>% 
+              mutate(ST = ifelse(ST %in% c('E. hormaechei ST114',
+                                 'K. pneumoniae ST258','K. pneumoniae ST307', 
+                                 'E. cloacae ST114'),
+                       ST, 'Other')) %>% 
+  left_join(cluster_info %>% rename(Sample_ID = isolate_no) %>% select(-ST)) %>% 
+  filter(!is.na(Facility_ID)) %>% 
+  group_by(Facility_ID, ST) %>% 
+  summarize(facil_st_count = n()) %>%
+  mutate(ST = gsub('E. hormaechei', 'KPC-Eh', ST),
+         ST = gsub('K. pneumoniae', 'KPC-Kp', ST),
+         ST = ifelse(ST == 'KPC-Eh ST171', 'Other', ST),
+         ST = factor(ST, levels = rev(c('KPC-Kp ST258', 'KPC-Eh ST114', 'KPC-Kp ST307', 'Other'))),
+) %>% 
+  group_by(Facility_ID) %>% 
+  mutate(dominant_st = case_when(is.infinite(max(facil_st_count[ST != 'Other'])) ~ 'Other\nmost common',
+                         max(facil_st_count[ST == 'KPC-Kp ST258']) == max(facil_st_count[ST != 'Other']) ~ 'KPC-Kp ST258\nmost common',
+                         max(facil_st_count[ST == 'KPC-Kp ST307']) == max(facil_st_count[ST != 'Other']) ~ 'KPC-Kp ST307\nmost common',
+                         max(facil_st_count[ST == 'KPC-Eh ST114']) == max(facil_st_count[ST != 'Other']) ~ 'KPC-Eh ST114\nmost common',
+                         TRUE ~ 'Other\nmost common'),
+         dominant_st = factor(dominant_st, levels = c('KPC-Kp ST258\nmost common', 'KPC-Kp ST307\nmost common',
+                                                      'KPC-Eh ST114\nmost common', 'Other\nmost common'))) %>% 
+  group_by(Facility_ID, dominant_st) %>% 
+  summarize(n_isolates = sum(facil_st_count)) %>% 
+  ungroup() %>% 
+  arrange(Facility_ID) %>% 
+  mutate(index = as.numeric(factor(Facility_ID)))
+```
+
+```{r}
 graph_tn <- tbl_graph(nodes = tn_facils_st %>% 
-                        mutate(facil_st_type = gsub('K. pneumoniae', 'KPC-Kp', facil_st_type),
-                               facil_st_type = gsub('E. hormaechei', 'KPC-Eh', facil_st_type),
-                               facil_st_type = gsub(' facility cluster', 
-                                                    '\nfacility cluster', facil_st_type)
-                               ), 
+                        mutate(dominant_st = gsub('\nmost common', '', dominant_st)), 
+                      # node_key = 'Facility_ID',
                       edges = tn_trans_net_sub %>% 
-                        filter(!grepl('I', source_facil) & !grepl('I', dest_facil) & n_transfers > 1),
+                        filter(!grepl('I', source_facil) & !grepl('I', dest_facil) & n_transfers > 1) %>% 
+                        left_join(tn_facils_st %>% rename(source_facil = Facility_ID, from = index) %>%
+                                    select(-c(dominant_st, n_isolates))) %>% 
+                        left_join(tn_facils_st %>% rename(dest_facil = Facility_ID, to = index) %>%
+                                    select(-c(dominant_st, n_isolates))),
                       directed = TRUE) %>%
   induced.subgraph(igraph::degree(.) > 0)
+
+st_cols_short <- st_cols
+names(st_cols_short) <- gsub('K. pneumoniae', 'KPC-Kp', names(st_cols_short))
+names(st_cols_short) <- gsub('E. hormaechei', 'KPC-Eh', names(st_cols_short))
 
@@
   geom_edge_link(aes(color = n_transfers, alpha = n_transfers), arrow = arrow(length = unit(3, 'mm'))) + 
-  geom_node_point(aes(color = facil_st_type, size = n_isolates)) + 
+  geom_node_point(aes(color = dominant_st, size = n_isolates)) + 
   # scale_edge_color_viridis(direction = -1, option = 'E') +
   scale_edge_color_continuous(high = '#00204c', low = '#ffe945') +
-  scale_color_manual(values = c(unname(st_cols[1]), 'coral3', 'lightsalmon', 'grey')) +
+  scale_color_manual(values = st_cols_short) +
   guides(edge_alpha = 'none') +
@@
 tn_facil_transfers <- tn_trans_net_sub %>% 
-  left_join(tn_facils_st %>% rename(source_of_interest = of_interest) %>% 
+  left_join(tn_facils_st %>% mutate(source_of_interest = dominant_st %in% c('KPC-Kp ST307\nmost common', 'KPC-Eh ST114\nmost common')) %>% 
               select(Facility_ID, source_of_interest), 
             by = c(source_facil = 'Facility_ID')) %>% 
-  left_join(tn_facils_st %>% rename(dest_of_interest = of_interest) %>% 
+  left_join(tn_facils_st %>% mutate(dest_of_interest = dominant_st %in% c('KPC-Kp ST307\nmost common', 'KPC-Eh ST114\nmost common')) %>% 
               select(Facility_ID, dest_of_interest), 
@@
 tn_trans_net_sub %>% 
-  left_join(tn_facils_st %>% rename(source_of_interest = of_interest) %>% 
+  left_join(tn_facils_st %>% mutate(source_of_interest = dominant_st %in% c('KPC-Kp ST307\nmost common', 'KPC-Eh ST114\nmost common')) %>% 
               select(Facility_ID, source_of_interest), 
             by = c(source_facil = 'Facility_ID')) %>% 
-  left_join(tn_facils_st %>% rename(dest_of_interest = of_interest) %>% 
+  left_join(tn_facils_st %>% mutate(dest_of_interest = dominant_st %in% c('KPC-Kp ST307\nmost common', 'KPC-Eh ST114\nmost common')) %>% 
               select(Facility_ID, dest_of_interest), 
@@
   filter(!grepl('I', source_facil) & !grepl('I', dest_facil) & n_transfers > 1) %>% 
-  filter(source_of_interest | dest_of_interest) %>% 
   ungroup() %>% 
@@
 ggsave(filename = 'figures/Fig4.png', plot = f4, width = 12, height = 5)
+ggsave(filename = 'figures/Fig4.tiff', plot = f4, width = 12, height = 5)
 f4

### 2023-06-15 plasmid_sharing/CRE_EIP_Manuscript_Supplemental_Figure_5_Final_061323.Rmd
Updates based on review responses

---
title: "CRE EIP Manuscript Supplemental Figure 5 Figure"
author: "Kyle Gontjes"
date: "`r Sys.Date()`"
output: html_document
---

# Project: EIP CRE Manuscript

# Date: 06/13/23

# Author: Kyle Gontjes

# Purpose: Generate final figure with KPC-associated plasmid cluster sharing for isolates found in three states: Minnesota, Tennessee, and Connecticut

# Great Lakes Location:

### 1. Project Directory: /nfs/turbo/umms-esnitkin/Project_CRE_EIP/Analysis/Plasmid_analysis

### 2. Final Results Folder: 2023-06-13_CRE_EIP_final_manuscript_figure

# Notes:

### 1. Please note in the manuscript that this analysis evaluated the sharing of KPC-associated plasmid clusters, irrespective of the predicted location of the isolate pairs' KPC-containing contigs.

### 2. This was performed, instead of KPC-containing plasmid cluster sharing, because our use of short-read sequencing hinders our confidence in the specific location of the KPC-containing contig.

### 3. This caveat is important because nearly 60% of isolates had more than one KPC-associated plasmid cluster predicted via their short read assemblies. Therefore, two isolates could be classified as sharing a KPC-associated plasmid cluster, but each isolate's KPC-containing contig could be located on a different KPC-associated plasmid.

### 4. I talked w/ Evan and we agreed that we could keep the KPC-associated plasmid cluster analyses assuming that we are clear about our expectations and the limitations of the data.

```{r setup, include=FALSE}
require(knitr)
opts_knit$set(echo = TRUE,warning = FALSE, message = FALSE) 
opts_knit$set(root.dir='/Users/kyle/Desktop/gl_turbo/Project_CRE_EIP/Analysis/Plasmid_analysis/')
```

# Create Environment

```{r create environment,echo=T,include=T,results=T}
#Packages
packages <- c("tidyverse","cowplot","RColorBrewer","colorRamp2","grid","ComplexHeatmap")

#Load Packages
invisible(lapply(packages,library,character.only=T))
Sys.info()
sessionInfo()
```

# Load CRE EIP Metadata

### 1. I am working with the 2023-05_02_CRE_EIP_QC.csv generated by Ali. This includes Kleborate and Ariba ST data on all of our isolates.

```{r,include=T,echo=T,results=T,message=F,warning=F}
### Load Metadata ###
df <- readRDS("./data/CRE_EIP_metadata.RDS")

# Create State Variable 
df <- df %>% mutate(state = ifelse(grepl("MN",isolate_no),"MN", ifelse(grepl("TN",isolate_no),"TN","CT")))
```

# Species Recoding Notes:

### 1. We are using Kleborate calls for species (found in the variable, Kleborate_species).

### 2. I abbreviated the genus' and collapsed subspecies into their species.

### 3. For K. pneumoniae, we are using the Kleborate sequence type (ST) data (found in the variable, ST).

### 3. For Enterobacter species, we are using the E. cloacae ST data generated using Ariba and found in "Ecloacae.ST" variable.

### 4. Low-confidence ST calls were removed (i.e., "\*", "Novel", "ND", or "1LV")

```{r}
### Recode Species ###
# Abbreviating genus name and collapsing subspecies into their species.
df <- df %>% mutate(Kleborate_species =  dplyr::recode(Kleborate_species,
                                             `Citrobacter freundii"` =  "C. freundii",
                                             `Citrobacter portucalensis` =  "C. portucalensis",
                                             `Citrobacter werkmanii` =  "C. werkmanii",
                                             `Enterobacter asburiae` =  "E. asburiae",
                                             `Enterobacter cloacae` = "E. cloacae",
                                             `Enterobacter hormaechei` =  "E. hormaechei",
                                             `Enterobacter roggenkampii` =  "E. roggenkampii",
                                             `Escherichia coli / Shigella` =  "E. coli / Shigella",
                                             `Klebsiella aerogenes` =  "K. aerogenes",
                                             `Klebsiella michiganensis` =  "K. michiganensis",
                                             `Klebsiella oxytoca` =  "K. oxytoca",
                                             `Klebsiella pneumoniae` =  "K. pneumoniae",
                                             `Klebsiella quasipneumoniae subsp. quasipneumoniae` =  "K. quasipneumoniae",
                                             `Klebsiella quasipneumoniae subsp. similipneumoniae` =  "K. quasipneumoniae",
                                             `Klebsiella variicola subsp. variicola` =  "K. variicola",
                                             `Lelliottia nimipressuralis` =  "L. nimipressuralis",
                                             `Serratia marcescens` =  "S. marcescens"))

### Only keep K. pneumoniae & E. hormaechei isolates ###
df <- df %>% subset(Kleborate_species %in% c("K. pneumoniae","E. hormaechei"))

### Species/ST Recoding ###
# Removed K. pneumoniae isolates with 1LV or novel ST
KPN_isolates_to_remove <- df %>% subset(Kleborate_species == "K. pneumoniae") %>% subset(grepl("\\*",ST) == T | ST == "Novel" | ST =="ND" | grepl("1LV",ST) == T) %>% select(isolate_no) %>% unlist

# Removed E. hormaechei isolates with low confidence Ariba ST call
Enterobacter_isolates_to_remove <- df %>% subset(Kleborate_species == "E. hormaechei") %>%  subset(grepl("\\*",Ecloacae.ST) == T | Ecloacae.ST == "Novel" | Ecloacae.ST =="ND") %>% select(isolate_no) %>% unlist

# Remove the isolates
df <- df %>% subset(!isolate_no %in% c(KPN_isolates_to_remove,Enterobacter_isolates_to_remove))

### Add ST to K. pneumoniae and Enterobacter isolates ###
df <- df %>% mutate(ST_Kpn_Eh = ifelse(Kleborate_species == "K. pneumoniae",paste0("K. pneumoniae ",ST),
                                   ifelse(Kleborate_species %in% c("E. hormaechei"),paste0("E. hormaechei ST ",Ecloacae.ST),Kleborate_species)))
```

# Load Mob-Suite Data.

### 1. Mob_recon was run using default settings on the 2023-05-05 genome assemblies found in the project's Sequence_data/assembly/2023-05-05_assembly directory

### 2. blastn was run to identify KPC alleles found in the CARD database on our isolate assemblies

### 3. KPC-containing plasmids were identified by linking the KPC-containing plasmid to their location (i.e., chromosome or plasmid X)

```{r}
### Load KPC & Mobtyper Data ###
# Matrices
plasmid_cluster <- readRDS("./2023-05-09_mobsuite/results/CRE_EIP_mobtyper_plasmid_cluster_matrix_050923.RDS")

# Mobtyper Contig Data
mobtyper_results_kpc <- readRDS("./2023-05-09_mobsuite/results/CRE_EIP_mobtyper_results_clean_KPC_050923.RDS")

### KPC-Associated Plasmid Clusters ###
KPC_associated_plasmid_clusters <- subset(mobtyper_results_kpc,KPC_present==1) %>% select(primary_cluster_id) %>% unique %>% unlist(use.names = F)

KPC_associated_plasmid_cluster_matrix <- plasmid_cluster %>% select(all_of(KPC_associated_plasmid_clusters))

### Get KPC Containing Isolates ###
KPC_containing_isolates <- KPC_associated_plasmid_cluster_matrix %>% subset(rowSums(.)>0) %>% rownames %>% unlist
df <- df %>% subset(isolate_no %in% KPC_containing_isolates)
```

# Generate State-Specific Data

```{r,include=T,echo=F,results=T,message=F,warning=F}
### Get State-Specific Isolates ###
MN_isolates <- df %>% subset(state == "MN") %>% select(isolate_no) %>% unlist
TN_isolates <- df %>% subset(state == "TN") %>% select(isolate_no) %>% unlist
CT_isolates <- df %>% subset(state == "CT") %>% select(isolate_no) %>% unlist

### Get State-specific dataframe ###
MN_df <-  df %>% subset(state == "MN")
TN_df <-  df %>% subset(state == "TN")
CT_df <- df %>% subset(state == "CT")

### State specific KPC-associated plasmid cluster matrices ###
MN_KPC_plasmid_cluster <- KPC_associated_plasmid_cluster_matrix %>%  subset(rownames(.) %in% MN_isolates) %>% select(any_of(KPC_associated_plasmid_clusters)) %>% select_if(colSums(.)>0)

TN_KPC_plasmid_cluster <- KPC_associated_plasmid_cluster_matrix %>% subset(rownames(.) %in% TN_isolates) %>% select(any_of(KPC_associated_plasmid_clusters)) %>% select_if(colSums(.)>0)

CT_KPC_plasmid_cluster <- KPC_associated_plasmid_cluster_matrix %>% subset(rownames(.) %in% CT_isolates) %>% select(any_of(KPC_associated_plasmid_clusters)) %>% select_if(colSums(.)>0)
```

# Algorithm to Characterize Plasmid Sharing Across Species/ST Boundaries

## Step 1. Create "Other" Sequence Type Categories

### 1. All sequence types with \<5 isolates were grouped into "other" categories

```{r}
### Create "other" sequence type categories for Kpn and Ehc ###
collapse_STs_under_5 <- function(df){

  Kpn_names_5 <- df %>% subset(Kleborate_species == "K. pneumoniae") %>% select(ST_Kpn_Eh) %>% table %>% ifelse(. >4,.,"Other K. pneumoniae STs") %>% ifelse(. != "Other K. pneumoniae STs",names(.),.)
  Eh_names_5 <- df %>% subset(Kleborate_species %in% c("E. hormaechei")) %>% select(ST_Kpn_Eh) %>% table %>%  ifelse(. >4,.,"Other E. hormaechei STs") %>% ifelse(. != "Other E. hormaechei STs",names(.),.)
  Renames_5 <- c(Kpn_names_5,Eh_names_5)

  updated_df <- df %>% mutate(Species_Kpn_Eh_ST_5 = dplyr::recode(ST_Kpn_Eh,!!!Renames_5))

  return(updated_df)
}
```

## Step 2. Calculate the Proportion of KPC-Associated Plasmid Sharing

### 1. The proportion of plasmid sharing across species/ST categories was calculated by dividing the number of isolate pairs across the two categories that share AT LEAST ONE KPC-associated plasmid cluster by the total number of isolate pairs across the two categories.

### 2. This is calculated irrespective of whether the isolate pairs both have KPC identified on the shared KPC-associated plasmid cluster(s).

```{r}
### Plasmid sharing algorithm ###
plasmid_sharing_algorithm <- function(dataset,species_variable,plasmid_matrix){

  ## Get total number of isolates per species
  isolates_per_species <- table(dataset[,species_variable]) %>% as.vector %>% `names<-`(names(table(dataset[,species_variable])))

  ## Total comparison count for species/ST boundary comparisons##
  total_comparison_count <- c()

  for(i in 1:length(isolates_per_species)){
   total_comparison_count[[i]] <- isolates_per_species[i]*isolates_per_species
  } %>% `names<-`(isolates_per_species)

  total_comparison_count <- do.call(cbind,total_comparison_count) %>% as.data.frame %>% `colnames<-`(names(isolates_per_species))

  ## Isolates and their species/ST data
  isolate_STs <- dataset %>% select(isolate_no,paste(species_variable))

  ## Function to get the number of plasmids shared
  get_number_of_plasmids_shared <- function(ST1,ST2,isolate_ST_dataset,plasmid_matrix){

  # ST1 Isolates Data
  ST1_isolates <- isolate_ST_dataset[isolate_ST_dataset[,paste(species_variable)] == ST1,]  %>% select(isolate_no) %>% unlist %>% `names<-`(NULL)
  ST1_plasmid_matrix <- plasmid_matrix %>% subset(rownames(.) %in% ST1_isolates) %>% subset(rowSums(.)>0) %>%  as.matrix

  # ST2 Isolates Data
  ST2_isolates <- isolate_ST_dataset[isolate_ST_dataset[,paste(species_variable)] == ST2,]  %>% select(isolate_no) %>% unlist %>% `names<-`(NULL)
  ST2_plasmid_matrix <- plasmid_matrix %>% subset(rownames(.) %in% ST2_isolates) %>% subset(rowSums(.)>0) %>%  as.matrix

  # Compare plasmid sharing of isolates  
  compare_ST1_to_ST2 <- apply(ST1_plasmid_matrix,MARGIN = 1,FUN=function(x){
      isolate_1 <- x %>% subset(.==1) %>% names(.) %>% unlist

      sharing_sum_for_isolate_1 <- apply(ST2_plasmid_matrix,MARGIN = 1,FUN=function(y){
        isolate_2 <- y %>% subset(.==1) %>% names(.) %>% unlist 
        count <- ifelse(sum(isolate_1 %in% isolate_2)>0,1,0)
        return(count)
      }) %>% sum()
      return(sharing_sum_for_isolate_1)
  })

   pairs_sharing_plasmids_ST1_to_ST2 <- sum(compare_ST1_to_ST2)
   return(pairs_sharing_plasmids_ST1_to_ST2)
  }

  ## Calculate plasmid sharing
  plasmid_sharing <- c()

  for(i in 1:length(isolates_per_species)){
    plasmid_sharing[[i]] <- lapply(colnames(total_comparison_count),FUN=get_number_of_plasmids_shared,ST2=colnames(total_comparison_count)[i],isolate_ST_dataset = isolate_STs,plasmid_matrix = plasmid_matrix)
} %>% `names<-`(isolates_per_species)

  plasmid_sharing_matrix <- do.call(cbind,plasmid_sharing) %>% as.data.frame %>% `colnames<-`(names(isolates_per_species)) %>% `rownames<-`(names(isolates_per_species))

  ## Divide the plasmid sharing data by the total number of possible pairs
  proportion_plasmid_sharing_matrix <- c()
  for(i in 1:length(isolates_per_species)){
   proportion_plasmid_sharing_matrix[[i]] <- plasmid_sharing_matrix[i,]%>% as.vector %>% unlist / total_comparison_count[i,]%>% as.vector%>% unlist
  }
  proportion_plasmid_sharing_matrix <- do.call(cbind,proportion_plasmid_sharing_matrix) %>% as.data.frame %>% `colnames<-`(paste0(names(isolates_per_species)," (n=",isolates_per_species,")")) %>% `rownames<-`(paste0(names(isolates_per_species)," (n=",isolates_per_species,")"))

  return(proportion_plasmid_sharing_matrix)
}
```

## Step 3. Visualize the data using ComplexHeatmap

### 1. I initially tried to work with the pheatmap package. Unfortunately, it is not very customizable.

### 2. In response, I implemented the heatmap using ComplexHeatmap

-   ComplexHeatmap implemented pheatmap function in their package alongside many useful add-ons.

-   Thus, I kept the formatting provided by pheatmap, but altered the legend

### 3. Citation information for ComplexHeatmap:

-   Doi: <https://academic.oup.com/bioinformatics/article/32/18/2847/1743594>

-   Google Scholar Citation: Gu, Z., Eils, R. and Schlesner, M., 2016. Complex heatmaps reveal patterns and correlations in multidimensional genomic data. *Bioinformatics*, *32*(18), pp.2847-2849.

```{r}
### Create Heatmap for Plasmid Sharing ###
pheatmap_plasmid_sharing <- function(proportion_plasmid_sharing_matrix,title){
  figure <- ComplexHeatmap::pheatmap(proportion_plasmid_sharing_matrix %>% as.matrix,border_color = "black",name ="KPC-Associated Plasmid Sharing",legend=F,main =title,fontsize = 26,
    row_names_max_width = max_text_width(rownames(proportion_plasmid_sharing_matrix), gp = gpar(fontsize = 26)),
    column_names_max_height = max_text_width(rownames(proportion_plasmid_sharing_matrix), gp = gpar(fontsize = 26)),
    legend_breaks = c(0, .25, .5, .75, 1),legend_labels = c("0%", "25%", "50%", "75%", "100%"))
  return(figure)
}
```

# Create Individual Heatmaps for Each State

## A. Tennessee (TN)

```{r,fig.height=15.5,fig.width=14.33}
TN_df <- collapse_STs_under_5(TN_df)

plasmid_sharing_5_isolate_cutoff_TN <- plasmid_sharing_algorithm(TN_df,"Species_Kpn_Eh_ST_5",TN_KPC_plasmid_cluster)

plasmid_sharing_5_isolate_cutoff_TN_heatmap <- pheatmap_plasmid_sharing(plasmid_sharing_5_isolate_cutoff_TN,title="Tennessee")

TN_heatmap_grob <- grid.grabExpr(draw(plasmid_sharing_5_isolate_cutoff_TN_heatmap,heatmap_legend_side = "bottom"))
```

## B. Minnesota (MN)

```{r,fig.height=15.5,fig.width=14.33}
MN_df <- collapse_STs_under_5(MN_df)

plasmid_sharing_5_isolate_cutoff_MN <- plasmid_sharing_algorithm(MN_df,"Species_Kpn_Eh_ST_5",MN_KPC_plasmid_cluster)

plasmid_sharing_5_isolate_cutoff_MN_heatmap <- pheatmap_plasmid_sharing(plasmid_sharing_5_isolate_cutoff_MN,title="Minnesota")

MN_heatmap_grob <- grid.grabExpr(draw(plasmid_sharing_5_isolate_cutoff_MN_heatmap,heatmap_legend_side = "bottom"))
```

## C. Connecticut (CT)

```{r,fig.height=15.5,fig.width=14.33}
CT_df <- collapse_STs_under_5(CT_df)

plasmid_sharing_5_isolate_cutoff_CT <- plasmid_sharing_algorithm(CT_df,"Species_Kpn_Eh_ST_5",CT_KPC_plasmid_cluster)

plasmid_sharing_5_isolate_cutoff_CT_heatmap <- pheatmap_plasmid_sharing(plasmid_sharing_5_isolate_cutoff_CT,title="Connecticut")

CT_heatmap_grob <- grid.grabExpr(draw(plasmid_sharing_5_isolate_cutoff_CT_heatmap,heatmap_legend_side = "bottom"))
```

# Plot & Save Figure With All Heatmaps for Publication

```{r,fig.height=18.6,fig.width=46.5}
### Plot Heatmaps Together ###
master_heatmap <- plot_grid(CT_heatmap_grob,MN_heatmap_grob,TN_heatmap_grob,ncol=3, axis="l", align="hv")

### Create Legend ###
one_hundred_breaks <- c()
for(i in 1:100){
  one_hundred_breaks[i] <- 0.01*i
}

fig_legend <- Legend(at = c(0, .25, .5, .75, 1),labels = c("0%", "25%", "50%", "75%", "100%"),title = "KPC-Associated Plasmid Sharing",direction = "horizontal",col_fun = colorRamp2(breaks = one_hundred_breaks,colors = colorRampPalette(rev(brewer.pal(n = 7, name =  "RdYlBu")))(100)), title_gp = gpar(fontsize = 32, fontface = "bold"),labels_gp=gpar(fontsize=32),legend_width = unit(18, "cm"))

fig_legend_image <- grid.grabExpr(draw(fig_legend))

### Plot & Save Heatmap + Legend Plots ###
plasmid_sharing_heatmap_final <- plot_grid(master_heatmap,NULL,fig_legend_image,NULL,ncol=1,axis="l", align="hv",rel_heights = c(1,-.085,.25,-.085))

plasmid_sharing_heatmap_final

# PDF (Normal Size, w = 46.5; h = 18.6. This converts to w = 6.5; height = 2.6 in word)
ggsave("~/Desktop/gl_turbo/Project_CRE_EIP/Analysis/Plasmid_analysis/2023-06-13_CRE_EIP_final_manuscript_figure/results/CRE_EIP_manuscript_supplemental_figure_5_final_061323.pdf",width = 46.5,height = 18.6,dpi = 600)

# EPS
ggsave("~/Desktop/gl_turbo/Project_CRE_EIP/Analysis/Plasmid_analysis/2023-06-13_CRE_EIP_final_manuscript_figure/results/CRE_EIP_manuscript_supplemental_figure_5_final_061323.eps",width = 46.5,height = 18.6,dpi = 600)
```
